### 2023-03-27 paper_2023/paper.tex
initial version

\documentclass{article}

\usepackage{arxiv}
\usepackage[utf8]{inputenc} % allow utf-8 input
\usepackage[T1]{fontenc}    % use 8-bit T1 fonts
\usepackage{hyperref}       % hyperlinks
\usepackage{url}            % simple URL typesetting
\usepackage{booktabs}       % professional-quality tables
\usepackage{amsfonts}       % blackboard math symbols
\usepackage{nicefrac}       % compact symbols for 1/2, etc.
\usepackage{microtype}      % microtypography
\usepackage{cleveref}       % smart cross-referencing
\usepackage{lipsum}         % Can be removed after putting your text content
\usepackage{graphicx}
\usepackage{natbib}
\usepackage{doi}
\usepackage{listings}

\title{Colonies - Compute Continuums across Platforms}

\author{{\hspace{1mm}Johan Kristiansson} \\
	Department of Computer Science \\
	RISE Research Institutes of Sweden \\
	Luleå, Sweden \\
	\texttt{johan.kristiansson@ri.se} \\
	\And
	{\hspace{1mm}Thomas Ohlson Timoudas} \\
	Department of Computer Science \\
	RISE Research Institutes of Sweden \\
	Luleå, Sweden \\
	\texttt{thomas.ohlson.timoudas@ri.se} \\
	\And
	{\hspace{1mm}Henrik Forsgren} \\
	Department of Computer Science \\
	RISE Research Institutes of Sweden \\
	Luleå, Sweden \\
	\texttt{thomas.ohlson.timoudas@ri.se} \\
}

% Uncomment to override  the `A preprint' in the header
%\renewcommand{\headeright}{Technical Report}
%\renewcommand{\undertitle}{Technical Report}
\renewcommand{\shorttitle}{\textit{arXiv} Template}

%%% Add PDF metadata to help others organize their library
%%% Once the PDF is generated, you can check the metadata with
%%% $ pdfinfo template.pdf
\hypersetup{
pdftitle={A template for the arxiv style},
pdfsubject={q-bio.NC, q-bio.QM},
pdfauthor={David S.~Hippocampus, Elias D.~Striatum},
pdfkeywords={First keyword, Second keyword, More},
}

\begin{document}
\maketitle

\begin{abstract}
This paper presents a novel framework for building 
\end{abstract}

% keywords can be removed
\keywords{Serverless computing \and Parallel computing \and Workflow orchestration}

\section{Introduction}
TODO
\section{The Colonies framework}
\label{sec:headings}

\subsection{Architecture}
TODO
\begin{figure}[h]
	\centering
    \includegraphics[scale=0.4]{arch.png}
	\caption{cron management}
	\label{fig:fig1}
\end{figure}

\begin{figure}[h]
	\centering
    \includegraphics[scale=0.4]{raft.png}
	\caption{cron management}
	\label{fig:fig1}
\end{figure}

\subsubsection{Workflows}
TODO
\begin{figure}[h]
	\centering
    \includegraphics[scale=0.32]{workflow.png}
	\caption{cron management}
	\label{fig:fig1}
\end{figure}

\begin{table}[h]
	\caption{Function Specifications}
	\centering
	\begin{tabular}{llllll}
		\toprule
		\cmidrule(r){1-2}
        Function Spec & Function        & Executor Type & Priority & Max Exec Time & Max Retries \\
		\midrule
        $F_{1}$       & gen\_nums()     & Edge          & 1        & 200 s         & 5 \\
        $F_{2}$       & square()        & Cloud         & 1        & 200 s         & 5 \\
        $F_{3}$       & square()        & Cloud         & 1        & 200 s         & 5 \\
        $F_{4}$       & sum()           & Browser       & 1        & 200 s         & 5 \\
		\bottomrule
	\end{tabular}
	\label{tab:table}
\end{table}

\begin{table}[h]
	\caption{Snapshot of Process Table as in Step 2}
	\centering
	\begin{tabular}{llllll}
		\toprule
		\cmidrule(r){1-2}
        Process Id & Function Spec & Wait for Parents & Assigned Executor Id & State      & Priority Time \\
		\midrule
        $P_{1}$    & $F_{1}$       & $False$          & $E_{1}$              & Successful & 1679906715352024000 \\
        $P_{2}$    & $F_{2}$       & $False$          & $E_{1}$              & Running    & 1679906715353453000 \\
        $P_{3}$    & $F_{3}$       & $False$          & $E_{2}$              & Running    & 1679906715354286000 \\
        $P_{4}$    & $F_{4}$       & $True$           & -                    & Waiting    & 1679906715355188000 \\
		\bottomrule
	\end{tabular}
	\label{tab:table}
\end{table}

dt = -1000000000 * 60 * 60 * 24
process.PriorityTime = int64(process.FunctionSpec.Priority)*dt + submissionTime.UnixNano()

\begin{table}[h]
	\caption{Dependency Table}
	\centering
	\begin{tabular}{lll}
		\toprule
		\cmidrule(r){1-2}
        Process Id & Name       & Dependencies           \\
		\midrule
        $P_{1}$    & $Task_{1}$ & -                      \\
        $P_{2}$    & $Task_{2}$ & $Task_{1}$             \\
        $P_{3}$    & $Task_{3}$ & $Task_{1}$             \\
        $P_{4}$    & $Task_{4}$ & $Task_{2}$, $Task_{3}$ \\
		\bottomrule
	\end{tabular}
	\label{tab:table}
\end{table}

\begin{table}[h]
	\caption{Input/Output Table}
	\centering
	\begin{tabular}{lll}
		\toprule
		\cmidrule(r){1-2}
        Process Id & Input & Output \\
		\midrule
        $P_{1}$    & & [2,3] \\
        $P_{2}$    & 2 & 4 \\
        $P_{3}$    & 3 & 9 \\
        $P_{4}$    & [4,9] & 13 \\
		\bottomrule
	\end{tabular}
	\label{tab:table}
\end{table}

\subsubsection{Cron}
TODO
\begin{figure}[h]
	\centering
    \includegraphics[scale=0.4]{cron.png}
	\caption{Sample figure caption.}
	\label{fig:fig1}
\end{figure}

\subsubsection{Generators}
TODO

\subsubsection{Zero-trust security}
TODO

\section{Evaluation}
\subsection{Implementation}
\begin{lstlisting}[language=c]
gen_nums = Function(gen_data, colonyid, executortype="edge")
square1 = Function(square, colonyid, executortype="cloud")
square2 = Function(square, colonyid, executortype="cloud")
sum = Function(square, colonyid, executortype="browser")

wf = ColoniesWorkflow("localhost", 50080, colonyid, executor_prvkey)
wf >> gennums
gennums >> square1
gennums >> square2
[square1, square2] >> sum
res = wf.execute()
\end{lstlisting}

\subsection{References}
TODO
\bibliographystyle{unsrtnat}
\bibliography{references}  %%% Uncomment this line and comment out the ``thebibliography'' section below to use the external .bib file (using bibtex) .

%%% Uncomment this section and comment out the \bibliography{references} line above to use inline references.
% \begin{thebibliography}{1}

% 	\bibitem{kour2014real}
% 	George Kour and Raid Saabne.
% 	\newblock Real-time segmentation of on-line handwritten arabic script.
% 	\newblock In {\em Frontiers in Handwriting Recognition (ICFHR), 2014 14th
% 			International Conference on}, pages 417--422. IEEE, 2014.

% 	\bibitem{kour2014fast}
% 	George Kour and Raid Saabne.
% 	\newblock Fast classification of handwritten on-line arabic characters.
% 	\newblock In {\em Soft Computing and Pattern Recognition (SoCPaR), 2014 6th
% 			International Conference of}, pages 312--318. IEEE, 2014.

% 	\bibitem{hadash2018estimate}
% 	Guy Hadash, Einat Kermany, Boaz Carmeli, Ofer Lavi, George Kour, and Alon
% 	Jacovi.
% 	\newblock Estimate and replace: A novel approach to integrating deep neural
% 	networks with existing applications.
% 	\newblock {\em arXiv preprint arXiv:1804.09028}, 2018.

% \end{thebibliography}

\end{document}

### 2023-03-28 paper_2023/paper.tex
wip

@@
 \begin{abstract}
-This paper presents a novel framework for building 
+Running AI/ML models in production is becoming more widespread, and maintaining these systems is more complex
+
+    This paper presents a novel framework for managing computational workload across heterogeneous platforms. Colonies is based on  
+a loosly coupled microservice architecture where complex workflows are broken down in composable functions that are executed by independently deployable executors.
+Using a HTTP protocol, workflows can be expressed declarative in any computer language and be executed across platforms by so-called executors, independently deployed in the cloud, edge, devices, or even in web browser creating compute continuums across platforms.
+Colonies supports both real-time processing and batch jobs while at the same time offer full tracability and zero-trust security.
+
+The paper
+
 \end{abstract}
@@
 TODO
+https://modelserving.com/blog/why-do-people-say-its-so-hard-to-deploy-a-ml-model-to-production
 \section{The Colonies framework}

### 2023-03-28 paper_2023/paper.tex
wip

--- 2023-03-27 paper_2023/paper.tex
@@
 \begin{abstract}
-This paper presents a novel framework for building 
+Running AI/ML models in production is becoming widespread.
+At the same time, developing and maintaining AI workloads are becoming more difficult.
+In particular, most workloads are not portable and cannot easily be moved from one provider to another.
+Creating and operating fully automated end-to-end workflows across devices, edge, cloud adds even more complexity.
+
+This paper presents a novel framework for running computational workload across heterogeneous platforms.
+Colonies is based on a loosely coupled microservice architecture where complex workflows are broken down in composable functions that are executed by independently deployable executors.
+Using a HTTP protocol, functions can be composed into declarative workflows in any computer language.
+The workflows are then executed across platforms by independently deployed executors running in the cloud, edge, devices, or even in web browser, creating compute continuums across platforms.
+Colonies support both real-time processing and batch jobs while at the same time offer full traceability and zero-trust security.
+
+The paper also describes how Colonies can be used to build a scalable remote sensing platform on Kubernetes, how it can be used as a building block for edge computing, and how it can be integrated with HPC platforms.
+Finally, the paper presents a performance investigation as well as a scalability and robustness evaluation.
 \end{abstract}
@@
 \section{Introduction}
-TODO
+Building robust and scalable AI systems is challenging and requires deep understanding in various fields.
+First an AI model must be trained, requiring technical skills in advanced statistics or machine learning, but also access to training and validation data.
+Usually, the data need to be pre-processed in several steps before it can be used, or a simulator needs to be developed to either generate syntentic data or play back historical data.
+While it may be resonable for small scale projects to run a whole environment on a local development computer, training large AI models usually requires access to powerful compute clusters or even HPC systems.
+Manually utilizing such infrastructure is cumbersome and time consuming.
+Being able to automate the training processes makes it possible to more quickly iterate and find useful models.
+
+Going forward and taking an AI model into production requires significant software engineering skills.
+In constract to traditional IT workloads both the data and the model itself need to be managed.
+As most models need to be re-trained or re-calibrated regularly, it muste be possible to seamlessly update deployed model and the software without loosing information or introducing delays.
+In many cases, there is a constant flow of data that is ingested into the system that needs to be managed while parts of the system is not working correctly.
+This becomes even more challenging when nodes parts of the underlying infrastructure crashes or becomes unavailable due to maintenance such as software updates or misconfiguration errors.
+
+Sometimes there is also need to scale the system to increase capacity or scale down to save resources.
+This is particularly important when using expensive cloud resources.
+Scaling the system means that underlying infastructure may change anytime causing additional failures.
+It must therefor be possible to detected failed computations and re-process failed tasks part of a larger workflow.
+If failed computation cannot gracefully be recovered, there must be a away for engineers perform root cause analysis and manually recover failures.
+
+Taking an AI model intro production,
+
+Development
+ - Data science
+ - Production system
+ - Integration, data ingestion, pre-processing
+ Workflows, real-time processing, back processing.
+Operation
+ - Computer faults
+ - Updates, DevOps
+Scability
+HPC
+ Requires a whole different set of toolchains.
+
+https://modelserving.com/blog/why-do-people-say-its-so-hard-to-deploy-a-ml-model-to-production
 \section{The Colonies framework}

### 2023-03-29 paper_2023/paper.tex
wip

@@
 
-The paper also describes how Colonies can be used to build a scalable remote sensing platform on Kubernetes, how it can be used as a building block for edge computing, and how it can be integrated with HPC platforms.
+In additional to a technical description, the paper also describes how Colonies can be used to build a scalable remote sensing platform on Kubernetes, how it can be used as a building block for edge computing, and how it can be integrated with HPC platforms.
 Finally, the paper presents a performance investigation as well as a scalability and robustness evaluation.
@@
 \section{Introduction}
-Building robust and scalable AI systems is challenging and requires deep understanding in various fields.
-First an AI model must be trained, requiring technical skills in advanced statistics or machine learning, but also access to training and validation data.
-Usually, the data need to be pre-processed in several steps before it can be used, or a simulator needs to be developed to either generate syntentic data or play back historical data.
-While it may be resonable for small scale projects to run a whole environment on a local development computer, training large AI models usually requires access to powerful compute clusters or even HPC systems.
-Manually utilizing such infrastructure is cumbersome and time consuming.
-Being able to automate the training processes makes it possible to more quickly iterate and find useful models.
+Building robust and scalable AI systems is challenging and requires a deep understanding of various fields.
+Firstly, an AI model must be trained, which requires technical skills in advanced statistics or machine learning, as well as access to training and validation data.
+Usually, the data needs to be pre-processed in several steps before it can be used, or a simulator needs to be developed to generate synthetic data or play back historical data.
+While it may be reasonable for small-scale projects to run the whole environment on local development computers, training larger AI models usually requires access to powerful compute clusters or even HPC systems.
+Manually utilizing such infrastructure is cumbersome and time-consuming.
+Automating the training processes makes it possible to iterate more quickly and find useful models.
 
 Going forward and taking an AI model into production requires significant software engineering skills.
-In constract to traditional IT workloads both the data and the model itself need to be managed.
-As most models need to be re-trained or re-calibrated regularly, it muste be possible to seamlessly update deployed model and the software without loosing information or introducing delays.
-In many cases, there is a constant flow of data that is ingested into the system that needs to be managed while parts of the system is not working correctly.
-This becomes even more challenging when nodes parts of the underlying infrastructure crashes or becomes unavailable due to maintenance such as software updates or misconfiguration errors.
+In contrast to traditional IT workloads, both the data and the model itself need to be managed.
+As most models need to be re-trained or re-calibrated regularly, it must be possible to seamlessly update the deployed model and the software without losing information or introducing delays.
+In many cases, there is a constant flow of data that is ingested into the system that needs to be managed while parts of the system are not working correctly.
+This becomes even more challenging when nodes or parts of the underlying infrastructure crash or become unavailable due to maintenance, such as software updates or misconfiguration errors.
 
@@
 This is particularly important when using expensive cloud resources.
-Scaling the system means that underlying infastructure may change anytime causing additional failures.
+Scaling the system means that underlying infastructure may change anytime causing instability problems for running applications.
 It must therefor be possible to detected failed computations and re-process failed tasks part of a larger workflow.
-If failed computation cannot gracefully be recovered, there must be a away for engineers perform root cause analysis and manually recover failures.
+Running workflows must therefor be designed to handle an ever changing infrastructure, and if a failed computation cannot gracefully be restored, there must be a away for engineers perform root cause analysis and manually recover failures.
+
+In reality, several systems must be integrated to build fully working AI applications.
+For example, data might must be captured from an IoT system or pulled from third-party services running on different domains than the compute cluster.
+With the introduction of edge computing, parts of the data pipeline might also execute on edge servers, or even in on devices to bring computation closer where the data is produced.
+Configuring and setting up such pipelines adds further complexity.
+
+In addition, many compute clusters run on-prem and somtimes there is a need to temporary increase the capacity by mixing on-prem and cloud resources, for example to handle peek load or re-process historical data.
+This also requires integration of various security protocols to provision and .
+Creating hybrid workflows where some jobs run on HPC systems requires even more software development efforts and is out of the scope for many users, preventing them from utilizing powerful hardware.
+Clearly, there is a need for a framework that can consolidate various platforms to simplify development and allowing computations to seamlessly run across platforms.
+
+This paper presents a loosly coupled microservice inspired architecuture that seperates workflow definitions from implementation and deployment.
+The ultimate goal is to create an architecuture where monolitic workflows are broken down into independent and seperated compute units, which can be dynamically added or removed to running workflows.
+These compute units could then be implemented in any computer language and be deployed anywhere on the Internet.
+The paper describes a framework called Colonies implementing such an architecutre.
+The paper describes several use cases
 

### 2023-03-29 paper_2023/paper.tex
wip

--- 2023-03-28 paper_2023/paper.tex
@@
 	\texttt{thomas.ohlson.timoudas@ri.se} \\
+	\And
+	{\hspace{1mm}Erik Källman} \\
+	Department of Computer Science \\
+	RISE Research Institutes of Sweden \\
+	Luleå, Sweden \\
+	\texttt{thomas.ohlson.timoudas@ri.se} \\
 }
@@
 \begin{abstract}
-Running AI/ML models in production is becoming more widespread, and maintaining these systems is more complex
+Artificial intelligence and machine learning has gained significant traction in recent years.
+At the same time, development and operation of AI workloads has become increasingly challenging.
+One difficulty is the lack of portability, making it cumbersome to move from one platform or provider to another.
+Creating and operating fully automated end-to-end workflows across devices, edge, and cloud platforms is even more challenging.
 
-    This paper presents a novel framework for managing computational workload across heterogeneous platforms. Colonies is based on  
-a loosly coupled microservice architecture where complex workflows are broken down in composable functions that are executed by independently deployable executors.
-Using a HTTP protocol, workflows can be expressed declarative in any computer language and be executed across platforms by so-called executors, independently deployed in the cloud, edge, devices, or even in web browser creating compute continuums across platforms.
-Colonies supports both real-time processing and batch jobs while at the same time offer full tracability and zero-trust security.
+To address these issues, the paper presents a novel framework called Colonies for running computational workloads across heterogeneous platforms.
+Colonies is built upon a loosely coupled microservice architecture that breaks down complex workflows into composable functions, which can be executed by independently deployed executors.
+With the use of an HTTP protocol, the functions can be composed into declarative workflows in any computer language, and be executed across platforms by executors running in the cloud, edge, devices, or even in web browsers, creating seamless compute continuums across platforms.
 
-The paper
+In addition to a technical description of the Colonies framework, the paper also describes some potential use cases.
+The paper describe how Colonies can be leveraged to build a scalable remote sensing platform on Kubernetes, serve as a building block for edge computing, and how it can be integrated with HPC platforms.
+Finally, the paper presents a performance investigation, as well as a scalability and robustness evaluation.
 
+In summary, Colonies is a highly versatile and scalable framework that can streamline the development and deployment of computational workloads across heterogeneous platforms while also ensuring full traceability and zero-trust security.
 \end{abstract}
@@
 \section{Introduction}
-TODO
-https://modelserving.com/blog/why-do-people-say-its-so-hard-to-deploy-a-ml-model-to-production
+Developing robust and scalable AI systems is a challenging task that requires deep understanding in several fields.
+To begin with, an AI model must be trained which requires knowledge in advanced statistics or machine learning, as well as access to training and validation data.
+Typically, this data must be pre-processed through various stages before it can be utilized.
+Although it may be practical for small-scale projects to run the entire training processes on local development computers, larger AI models typically require access to powerful compute clusters or even HPC systems.
+Manual use of such infrastructure can be laborious and time-consuming.
+Automating the training process enables faster iterations and quicker discovery of useful models.
+
+Taking an AI model into production requires substantial software engineering expertise.
+In contrast to traditional IT workloads, both the data and the model must be managed in addition to the software itself.
+As most models require regular re-training or re-calibration, it must be possible to update deployed models and software seamlessly without losing information or breaking the system.
+In many cases, there is a constant flow of data ingested into the system which must be managed even if components are malfunctioning.
+This becomes even more challenging when nodes or parts of the underlying infrastructure become unavailable due to maintenance or failaure such as software updates, hardware replacements or simply missconfiguration.
+
+In some cases, it may be necessary to scale the system to increase or reduce the capacity.
+This is especially critical when using expensive cloud resources.
+Scaling the system means that the underlying infrastructure may change at any time, causing instability issues for running services or workflows.
+Therefore, it must be possible to detect failed computations and reprocess failed tasks part of a larger workflow.
+Workflows must hence be designed to handle an ever-changing infrastructure, and if a failed computation cannot be restored gracefully, engineers must be able to quickly perform root cause analysis to manually recover the system.
+
+In reality, AI system requires integration of multiple systems.
+For instance, data may need to be captured from an IoT system or pulled from third-party database running on different domains than the compute cluster.
+With the emergence of edge computing, parts of a data pipeline may also run on edge servers to bring computations closer to data sources.
+Configuring and setting up such pipelines add even more complexity.
+
+Additionally, many compute clusters operate on-premises installations.
+Sometimes it is necessary to temporarily increase the capacity of on-prem clusters by combining resources from multiple providers, for example, adding cloud compute resources to handle peak loads or utilize HPC resources to quickly reprocess historical data.
+Developing hybrid workflows where some jobs run in the cloud and others run on HPC systems requires even more software development efforts and is beyond the scope of many users, preventing them from utilizing powerful hardware.
+Clearly, there is a need for a framework that can consolidate various platforms to simplify development and enable seamless execution across platforms.
+
+This paper presents a framework called Colonies, which is built around a loosely coupled microservices architecture that separates workflow definitions from implementation and deployment.
+The main objective is to create a tool where monolithic workflows can be broken down into independent and separated compute units that can be dynamically added or removed while executing workflows.
+These compute units can be implemented in any computer language and be deployed anywhere on the Internet.
+Ultimately, the Colonies framework facilitates coordination among diverse software systems operating on different platforms in executing unified workflows.
+
+The next section presents related work. Section ....
+
+\section{Related work}
+A workflow is a series of tasks that need to be completed in a specific order.
+Workflows are often represented as directed acyclic graphs (DAGs), where nodes represent tasks and edges represent dependencies or data flow between tasks.
+
+TODO:
+
 \section{The Colonies framework}
 \label{sec:headings}
+Microservices is an architectural design pattern in which an application is structured as a collection of small, independently deployable, and loosely coupled services that communicate with other microservices through well-defined APIs.
+By dividing the application into smaller, focused microservices, applications become easier to understand, maintain, and develop.
+Each microservice can be scaled independently, making it easier to handle increased demand for a certain service.
+Additionally, different microservices can be developed by using diverse technologies, frameworks, or programming languages, enabling developers to select the most suitable tools for each specific problem.
+By assigning ownership of specific microservices to individual teams, it also becomes easier to coordinate work and maintain a consistent development process.
 
-\subsection{Architecture}
-TODO
+Currently, microservices are primarily used to implement large-scale web applications or Internet applications requiring high-availability.
+It has not yet become a prevalent design principle for workload management or implementation of HPC applications.
+Instead, simple job scripts are commonly used.
+In some cases, message brokers (e.g., RabbitMQ) are used to build worker queues to distribute tasks among multiple workers.
+Although this approach may be effective for simpler applications, creating dependencies between tasks, such as controlling the order of execution or passing information between tasks is not straightforward.
+
+Colonies is based on a microservice architecture where workflows are decomposed into a set of functions.
+An example of a function could for example be to train a neural network, prepare a batch of data, or upload inferred results to a third-party database.
+An executor is responsible for executing one or several functions, making it very simular to a microservice.
+All coordination is managed by a cluster of Colonies servers, allowing complex workflows to be broken down into independent functions that can be developed and tested separately.
+Analogous to traditional microservices, scalability can be achieved simply by adding more executors.
+If an executor crashes during task execution, the task is automatically reassigned to another executor.
+
+\subsection{The role of queues for seperation of concerns}
+Separation of concerns (SoC) is a fundamental design principle in computer science that aims to break down a complex software system into smaller, more manageable parts.
+For example, HTTP APIs can be used to abstract away implementation detail and provide a clear and simple interface for interacting with a particular service.
+However, HTTP protocols alone are insufficient for handling dynamic environments where components frequently fails or the underlying infrastructures is constantly changing.
+
+To address such environments, an alternative mechanism is necessary.
+Queues enable different executors to communicate indirectly by acting as a buffer between them.
+Queues makes it possible to decouple each executor to operate independently, e.g.
+a service can be updated or replaced without affecting other executors.
+Queues also allow for asynchronous communication between executors, enabling them to process tasks at their own pace.
+This ensures that slower executors do not bottleneck faster ones, leading to a more efficient and scalable system.
+Most importantely, queues enable load balancing by distributing tasks among multiple executors, thus making it possible to parallelize workflow execution.
+
+Queues can be implemented in various ways.
+As previously mentioned, a common solution involves using a message broker.
+However, Colonies adopts an alternative approach by utilizing a standard database and querying it to assign tasks to different executors.
+The advantage of employing a database lies in the fine-grained control it offers for assigning task to executors.
+For instance, it allows a specific tasks be assigned to particular executor type, thus enabling a workflow to be executed across diverse platforms.
+
 \begin{figure}[h]

### 2023-03-29 paper_2023/paper.tex
wip

@@
 These compute units can be implemented in any computer language and be deployed anywhere on the Internet.
-Ultimately, the Colonies framework facilitates coordination among diverse software systems operating on different platforms in executing unified workflows.
+Ultimately, the Colonies framework facilitates coordination among diverse software components operating on different platforms in executing unified workflows.
 

### 2023-03-31 paper_2023/paper.tex
wip

--- 2023-03-29 paper_2023/paper.tex
@@
 \usepackage{listings}
+\usepackage{caption}
+\usepackage{subcaption}
 
@@
 	Luleå, Sweden \\
-	\texttt{thomas.ohlson.timoudas@ri.se} \\
+	\texttt{erik.kallman@ri.se} \\
 }
@@
 At the same time, development and operation of AI workloads has become increasingly challenging.
-One difficulty is the lack of portability, making it cumbersome to move from one platform or provider to another.
+One difficulty is the lack of portability, making it cumbersome to move from one platform to another.
 Creating and operating fully automated end-to-end workflows across devices, edge, and cloud platforms is even more challenging.
 
-To address these issues, the paper presents a novel framework called Colonies for running computational workloads across heterogeneous platforms.
-Colonies is built upon a loosely coupled microservice architecture that breaks down complex workflows into composable functions, which can be executed by independently deployed executors.
-With the use of an HTTP protocol, the functions can be composed into declarative workflows in any computer language, and be executed across platforms by executors running in the cloud, edge, devices, or even in web browsers, creating seamless compute continuums across platforms.
+To address the aforementioned challenges, the paper introduces a framework termed Colonies designed to facilitate the execution of computational workloads across a diverse range of platforms.
+Colonies is founded upon a loosely-coupled microservice architecture that beaks down complex workflows into composable functions.
+With the use of an HTTP protocol, these composable functions can be implemented in any computer language and be executed by independent executors deployed across various systems, e.g.
+cloud, edge, devices, or even in web browsers.
+Colonies orchestrates the execution and by using a zereo-trust security protocol, a collection of distributed executors can function as a single cohesive unit, thereby establishing a seamless compute continuum across multiple platforms.
 
@@
 As most models require regular re-training or re-calibration, it must be possible to update deployed models and software seamlessly without losing information or breaking the system.
-In many cases, there is a constant flow of data ingested into the system which must be managed even if components are malfunctioning.
+In many cases, there is a constant flow of data ingested into the system which must be managed even if some components are malfunctioning.
 This becomes even more challenging when nodes or parts of the underlying infrastructure become unavailable due to maintenance or failaure such as software updates, hardware replacements or simply missconfiguration.
@@
 Sometimes it is necessary to temporarily increase the capacity of on-prem clusters by combining resources from multiple providers, for example, adding cloud compute resources to handle peak loads or utilize HPC resources to quickly reprocess historical data.
-Developing hybrid workflows where some jobs run in the cloud and others run on HPC systems requires even more software development efforts and is beyond the scope of many users, preventing them from utilizing powerful hardware.
+Developing hybrid workflows where some jobs run in the cloud and others run on HPC systems requires even more software development efforts and is beyond the scope of most users, preventing them from utilizing powerful hardware.
 Clearly, there is a need for a framework that can consolidate various platforms to simplify development and enable seamless execution across platforms.
 
-This paper presents a framework called Colonies, which is built around a loosely coupled microservices architecture that separates workflow definitions from implementation and deployment.
+This paper presents a framework called Colonies, which is built around a loosely-coupled microservices architecture that separates workflow definitions from implementation and deployment.
 The main objective is to create a tool where monolithic workflows can be broken down into independent and separated compute units that can be dynamically added or removed while executing workflows.
 These compute units can be implemented in any computer language and be deployed anywhere on the Internet.
-Ultimately, the Colonies framework facilitates coordination among diverse software systems operating on different platforms in executing unified workflows.
 
-The next section presents related work. Section ....
+The next section presents related work. The next section describes related work. Section X ...
 
@@
 Workflows are often represented as directed acyclic graphs (DAGs), where nodes represent tasks and edges represent dependencies or data flow between tasks.
+Workflow management has been extensively studied in both academic and industrial settings, with numerous approaches proposed to address the challenges in this field.
+
+Another popular approach is the use of workflow engines, which provide a framework for creating and executing workflows.
+Workflow engines can be integrated with other tools and systems to automate complex business processes.
+Examples of workflow engines include Apache Airflow, Camunda, and Zeebe.
+
+Apache Airflow is a popular open-source workflow management system used to programmatically author, schedule, and monitor workflows.
+Luigi is also an open-source workflow management system developed by Spotify. It is designed to handle long-running batch jobs and data pipelines.
+
+Pinball is a workflow management system developed by Pinterest that allows users to define and execute workflows using Python scripts.
+It has a simple architecture, with support for DAGs and task dependencies.
+However, it lacks some of the advanced features of Airflow, such as task retries and monitoring.
 
@@
 \label{sec:headings}
-Microservices is an architectural design pattern in which an application is structured as a collection of small, independently deployable, and loosely coupled services that communicate with other microservices through well-defined APIs.
+Microservices is an architectural design pattern in which an application is structured as a collection of small, independently deployable, and loosely coupled services that communicate with other microservices through a well-defined API.
 By dividing the application into smaller, focused microservices, applications become easier to understand, maintain, and develop.
@@
 
-Colonies is based on a microservice architecture where workflows are decomposed into a set of functions.
-An example of a function could for example be to train a neural network, prepare a batch of data, or upload inferred results to a third-party database.
+Colonies is based on a microservice architecture where workflows are decomposed into a set of independent functions.
+An example of a function could for example be a function that trains a neural network, prepare a batch of data, or upload inferred results to a third-party database etc.
 An executor is responsible for executing one or several functions, making it very simular to a microservice.
 All coordination is managed by a cluster of Colonies servers, allowing complex workflows to be broken down into independent functions that can be developed and tested separately.
-Analogous to traditional microservices, scalability can be achieved simply by adding more executors.
+Analogous to traditional microservices, scalability can be achieved simply by adding more executors implementing the same function specification.
 If an executor crashes during task execution, the task is automatically reassigned to another executor.
 
-\subsection{The role of queues for seperation of concerns}
+\subsection{The role of queues as seperation of concerns}
 Separation of concerns (SoC) is a fundamental design principle in computer science that aims to break down a complex software system into smaller, more manageable parts.
@@
 However, HTTP protocols alone are insufficient for handling dynamic environments where components frequently fails or the underlying infrastructures is constantly changing.
+To address such environments, an alternative mechanism is necessary.
 
-To address such environments, an alternative mechanism is necessary.
-Queues enable different executors to communicate indirectly by acting as a buffer between them.
-Queues makes it possible to decouple each executor to operate independently, e.g.
-a service can be updated or replaced without affecting other executors.
+Queues enable different software components to communicate indirectly by acting as a buffer between them.
+Queues makes it possible to decouple each executor and make them operate independently, e.g.
+an executor can be updated or replaced without affecting other executors.
 Queues also allow for asynchronous communication between executors, enabling them to process tasks at their own pace.
@@
 
-Queues can be implemented in various ways.
-As previously mentioned, a common solution involves using a message broker.
-However, Colonies adopts an alternative approach by utilizing a standard database and querying it to assign tasks to different executors.
-The advantage of employing a database lies in the fine-grained control it offers for assigning task to executors.
-For instance, it allows a specific tasks be assigned to particular executor type, thus enabling a workflow to be executed across diverse platforms.
+Queues can be implemented in different ways, and while message brokers are a common solution, Colonies has adopted an alternative strategy by leveraging a standard database and querying it for tasks to assign to different executors.
+One key advantage of this approach is that it enables fine-grained task assignments, making it possible to assign specific tasks to particular executors.
+For instance, an executor of the browser type can be limited to only executing tasks of the browser type.
+This level of granularity cannot easily be implemented using message brokers which generally do not offer introspection of queue, or provide the ability to pull specific messages out of the queue.
+Generarally, the only way to retrieve a specific message is to pull all messages from the queue, obtain the message, and then place all remaining messages back into the queue in the same order.
+In contrast, using a database a query can match any columns and be ordered using a timestamp to make it function as a queue.
+
+\begin{figure}
+     \centering
+     \begin{subfigure}[b]{0.3\textwidth}
+         \centering
+         \includegraphics[scale=0.39]{arch.png}
+         \caption{Process assignment steps.}
+         \label{fig:y equals x}
+     \end{subfigure}
+     \hfill
+     \begin{subfigure}[b]{0.3\textwidth}
+         \centering
+         \includegraphics[scale=0.39]{process.png}
+         \caption{Manipulation of a process.}
+         \label{fig:three sin x}
+     \end{subfigure}
+     \caption{Process management via Colonies HTTP API. Note that only the assigned executor has write access to the process database entry.}
+     \label{fig:three graphs}
+\end{figure}
+
+\subsection{Process tables}
+Colonies provides a platform for executors to interact with each other by enabling them to submit function specifications to a Colonies server.
+Once submitted, other executors can connect to the server to receive execution assignments.
+When a function specification is submitted to the Colonies server, it is store as a process entry in a dedicated process table database.
+In this case, a process is analogous to a task, but contains additional contextual information about the execution, for example, execution status (waiting, running, successful, failed), assigned executor Id, submission time, execution time, deadlines, priority etc.
+It also contains input and argument values as well as return values from the function innovation.
+
+The Colonies server acts as a job broker for executors, almost like an employment agency.
+When an executor connects to the Colonies server, the server hang the incomming HTTP connection\footnote{An alternative protocol is to use WebSockets or gPRC to communicate with the Colonies server.} until the executor is assigned a process, or until a connection timer expires.
+Note that the Colonies server does not connect to any executor.
+Rather, it is the responsibility of the executors to connect to the Colonies server, enabling them to be deployed anywhere on the Internet, even behind firewalls or commercial telco networks.
+
+Figure \ref{fig:fig1} illustrates an executor submitting a function specification that is later assigned to another executor.
+Executors have to specify to the Colonies server which functions they are capable of executing.
+The Colonies server makes sure that the conditions (or requirmenets) of a function specification matches the capability of the executors.
 
@@
 	\centering
-    \includegraphics[scale=0.4]{arch.png}
-	\caption{cron management}
-	\label{fig:fig1}
+    \includegraphics[scale=0.35]{function_spec.png}
+	\caption{Example of a function spec.}
+	\label{fig:fig2}
 \end{figure}
+
+Figure \ref{fig:fig2} shows an example of a function specification.
+The \emph{maxexectime} attribute specifies the maximum time an executor may run a process (in this case, 100 seconds).
+Before a process is assigned to an executor, the Colonies server updates the process entry in the process table database and calculates a deadline for when the process must be finished.
+The server then regularly checks for any running processes that have exceeded their deadlines.
+If such process is detected, it is reset, allowing it to be re-assigned to other executors.
+This scheme is completely stateless, so if the Colonies server crashes, another Colonies server can take over and look for processes that have exceeded their deadline.
+Moreover, if an executor assigned to a process crashes, the process is simply re-executed by another executor.
+
+To scale up a system, more executors can be deployed.
+Scaling down, however, can be more challenging.
+One solution is to select a set of executors to be removed and then starv them out by denying them new process assignments.
+Another, simpler solution, is to immediately destroy the executors, or if possible notify them in advance so that they can gracefully unregister.
+
+The \emph{maxexectime} failsafe mechanism ensures that processes will eventually be executed even in the case of failures.
+This mechanism also relieves the burden of user to check if a process has been executed or not, as they can simply look up the process in the database to get its current status.
+It is also possible to subscribe on process changhes using an event-driven protocol, but this may be impractical if a process runs for a very long time (e.g.
+days or weeks).
 
@@
 	\caption{cron management}
-	\label{fig:fig1}
+	\label{fig:fig2}
 \end{figure}
+
+\subsubsection{Concurreny and synchronization}
+Synchronization is esstential to prevent data inconsistency and race conditions when accessing shared resources concurrently with multiple threads.
+However, synchronization also slows down execution as only one thread can access critical sections at a time.
+By carefully designing multithreaded applications and employing the right synchronization techniques, it possible to minimize the performance impact while still ensuring data consistency and correctness.
+
+The purpose of all Colonies APIs is to either modify the database or read information from the database. A request
+
+Colonies is designed to be stateless, meaning that the Colonies server does not keep any information about between API requests.
+In other words, each request is handled independently, without relying on any information from previous requests.
+Once the database has been updated,
+
+if an instance fails or becomes unavailable, subsequent requests can be easily redirected to another
+
+. After a The Colonies server ensure that only the assigned executor has write access to the process entry in the database.
+
+Colonies ensures that exactly one executor is assigned to a specific process, so only the assigned executor has permission to set output values and close the process as successful.
+However, there is a window of time during which an executor may execute a process.
 
@@
 	\caption{cron management}
-	\label{fig:fig1}
+	\label{fig:fig3}
 \end{figure}

### 2023-04-01 paper_2023/paper.tex
wip

@@
     In addition to a technical description of the Colonies framework, the paper also describes some potential use cases.
-The paper describe how Colonies can be leveraged to build a scalable remote sensing platform on Kubernetes, serve as a building block for edge computing, and how it can be integrated with HPC platforms.
-Finally, the paper presents a performance investigation, as well as a scalability and robustness evaluation.
+The paper describe how Colonies can be leveraged to build a remote sensing platform on Kubernetes, serve as a building block for edge computing, implement a serverless FaaS (Function-as-a-Service), and how it can be integrated with HPC platforms.
+Finally, the paper presents a performance investigation, as well as  scalability and robustness evaluation.
 
@@
 These compute units can be implemented in any computer language and be deployed anywhere on the Internet.
-
-The next section presents related work. The next section describes related work. Section X ...
 
@@
 
+\begin{equation}
+    priority_{time}=priority \cdot -10^9 \cdot 60 \cdot 60 \cdot 24 + submission_{time}  
+\end{equation}
+
 \begin{figure}
@@
          \centering
-         \includegraphics[scale=0.39]{arch.png}
+         \includegraphics[scale=0.37]{arch.png}
          \caption{Process assignment steps.}
-         \label{fig:y equals x}
      \end{subfigure}
@@
          \centering
-         \includegraphics[scale=0.39]{process.png}
+         \includegraphics[scale=0.37]{process.png}
          \caption{Manipulation of a process.}
-         \label{fig:three sin x}
      \end{subfigure}
      \caption{Process management via Colonies HTTP API. Note that only the assigned executor has write access to the process database entry.}
-     \label{fig:three graphs}
+     \label{fig:process_assignment}
 \end{figure}
@@
 
-The Colonies server acts as a job broker for executors, almost like an employment agency.
+The Colonies server acts as a job broker for executors, almost like an employment agency for people.
 When an executor connects to the Colonies server, the server hang the incomming HTTP connection\footnote{An alternative protocol is to use WebSockets or gPRC to communicate with the Colonies server.} until the executor is assigned a process, or until a connection timer expires.
 Note that the Colonies server does not connect to any executor.
-Rather, it is the responsibility of the executors to connect to the Colonies server, enabling them to be deployed anywhere on the Internet, even behind firewalls or commercial telco networks.
+Rather, it is the responsibility of the executors to connect to the Colonies server.
+This enables executors to be deployed anywhere on the Internet, even behind firewalls, commercial telco networks, or in application running in web browsers.
 
-Figure \ref{fig:fig1} illustrates an executor submitting a function specification that is later assigned to another executor.
+Figure \ref{fig:process_assignment} illustrates an executor submitting a function specification that is later assigned to another executor.
 Executors have to specify to the Colonies server which functions they are capable of executing.
-The Colonies server makes sure that the conditions (or requirmenets) of a function specification matches the capability of the executors.
+The Colonies server makes sure that the conditions (i.e requirmenets) of a function specification matches the capability of an executor.
 
@@
 	\caption{Example of a function spec.}
-	\label{fig:fig2}
+	\label{fig:function_spec}
 \end{figure}
 
-Figure \ref{fig:fig2} shows an example of a function specification.
+Figure \ref{fig:function_spec} shows an example of a function specification.
 The \emph{maxexectime} attribute specifies the maximum time an executor may run a process (in this case, 100 seconds).
@@
 
-To scale up a system, more executors can be deployed.
+Making it possible to specify maximum execution time is a simple but powerful mechanism.
+To scale up a system, more executors can simply be deployed.
 Scaling down, however, can be more challenging.
@@
 This mechanism also relieves the burden of user to check if a process has been executed or not, as they can simply look up the process in the database to get its current status.
-It is also possible to subscribe on process changhes using an event-driven protocol, but this may be impractical if a process runs for a very long time (e.g.
+The Colonies framework also supports subscribing on process changes using an event-driven protocol, but this may be impractical if a process runs for a very long time (e.g.
 days or weeks).
+
+Utilizing the \emph{maxexectime} failsafe mechanism not only enhances system reliability, but also provides an opportunity to apply of Chaos engineering \cite{chaos_engineering}.
+For example, deliberately terminating executors using a Chaos monkey.
+If executors are deployed on Kubernetes, Kubernetes will automatically redeploy terminated executors.
+The constant replacement of executor instances ensures that the system is capable of gracefully tolerating failures.
 
@@
 	\centering
-    \includegraphics[scale=0.4]{raft.png}
-	\caption{cron management}
-	\label{fig:fig2}
+    \includegraphics[scale=0.5]{raft.png}
+	\caption{High-availability deployment of Colonies servcer.}
+	\label{fig:ha_deployment}
 \end{figure}
@@
 Synchronization is esstential to prevent data inconsistency and race conditions when accessing shared resources concurrently with multiple threads.
-However, synchronization also slows down execution as only one thread can access critical sections at a time.
+However, synchronization can also slows down execution as only one thread can access critical sections at a time.
 By carefully designing multithreaded applications and employing the right synchronization techniques, it possible to minimize the performance impact while still ensuring data consistency and correctness.
 
-The purpose of all Colonies APIs is to either modify the database or read information from the database. A request
+The primary objective of all Colonies API requests is to either alter a state stored in the database or retrieve information from it.
+The Colonies framework is designed to be stateless, meaning that the Colonies server does not keep any information between requests.
+In other words, each request is handled independently, without relying on any information from previous requests.
 
-Colonies is designed to be stateless, meaning that the Colonies server does not keep any information about between API requests.
-In other words, each request is handled independently, without relying on any information from previous requests.
-Once the database has been updated,
+One example of a request is the assign request, which allocates a waiting process from the database to an executor.
+Given the multi-threaded nature of the Colonies server, it is essential that the assign request is synchronized to ensure that only one thread at a time can modify the database and update the process table, thus preventing multiple executors from being assigned to the same process.
+As only the assigned executor can modify a running process and only one executor must be assigned to a process, synchronization is required for the assign request.
+It is worth noting that synchronization is not necessary for other requests, such as the submit request or the close request.
+As the submit request only add new entries to the process table, there is no race conditions.
+The close request set the output of the function innovation and updates the process state to either successful or failed in the process table.
+Since there exists only one assigned executor there is no need for synchronization.
 
-if an instance fails or becomes unavailable, subsequent requests can be easily redirected to another
+To minimize downtime in the event of hardware or software failures, there is a need for high-availability and make use of a cluster of Colonies servers.
+If one Coloniser server crashes, an executor simply need to re-send a failed request, which will then be served by another Colonies server replica.
+However, by introducing multiple Colonies servers, there is again a risk of race conditions when assigning processes to executors.
+This means that the Colonies server replicas must coordintate which replica server incomming assign requests so that precisely one executor is assigned to a process.
 
-. After a The Colonies server ensure that only the assigned executor has write access to the process entry in the database.
+Raft \cite{raft} is a consensus algorithm specifically designed to manage a replicated log within a distributed system.
+It functions within a cluster of servers, where a single server takes on the role of leader while the remaining servers act as followers.
+The leader is responsible for managing the replicated log, processing client requests, and replicating entries to the followers.
+Followers passively replicate the leader's log and participate in leader elections.
+The leadership can change over time due to elections triggered by timeouts or other factors.
 
-Colonies ensures that exactly one executor is assigned to a specific process, so only the assigned executor has permission to set output values and close the process as successful.
-However, there is a window of time during which an executor may execute a process.
+Incorporating Raft into the Colonies framework allows incoming assign requests to be directed towards the leading Colonies server, ensuring that only one Colonies server replica handles these requests.
+Figure \ref{fig:ha_deployment} shows an overview of a high-availability Colonies deployment.
+A new leader is elected in the event that a Colonies server replica fails.
+The Raft protocol also enables seamless updates to the Colonies server software by making it possible to upgrade each replica individually.
+As a result, Colonies becomes well-suited for Kubernetes deployments.
 
@@
 	\centering
-    \includegraphics[scale=0.32]{workflow.png}
+    \includegraphics[scale=0.30]{workflow.png}
 	\caption{cron management}
@@
 \end{table}
-
-dt = -1000000000 * 60 * 60 * 24
-process.PriorityTime = int64(process.FunctionSpec.Priority)*dt + submissionTime.UnixNano()
 
@@
 \subsection{Implementation}
-\begin{lstlisting}[language=c]
-gen_nums = Function(gen_data, colonyid, executortype="edge")
-square1 = Function(square, colonyid, executortype="cloud")
-square2 = Function(square, colonyid, executortype="cloud")
-sum = Function(square, colonyid, executortype="browser")
 
-wf = ColoniesWorkflow("localhost", 50080, colonyid, executor_prvkey)
-wf >> gennums
-gennums >> square1
-gennums >> square2
-[square1, square2] >> sum
-res = wf.execute()
-\end{lstlisting}
-
-\subsection{References}
-TODO
-\bibliographystyle{unsrtnat}
-\bibliography{references}  %%% Uncomment this line and comment out the ``thebibliography'' section below to use the external .bib file (using bibtex) .
-
-%%% Uncomment this section and comment out the \bibliography{references} line above to use inline references.
-% \begin{thebibliography}{1}
-
-% 	\bibitem{kour2014real}
-% 	George Kour and Raid Saabne.
-% 	\newblock Real-time segmentation of on-line handwritten arabic script.
-% 	\newblock In {\em Frontiers in Handwriting Recognition (ICFHR), 2014 14th
-% 			International Conference on}, pages 417--422. IEEE, 2014.
-
-% 	\bibitem{kour2014fast}
-% 	George Kour and Raid Saabne.
-% 	\newblock Fast classification of handwritten on-line arabic characters.
-% 	\newblock In {\em Soft Computing and Pattern Recognition (SoCPaR), 2014 6th
-% 			International Conference of}, pages 312--318. IEEE, 2014.
-
-% 	\bibitem{hadash2018estimate}
-% 	Guy Hadash, Einat Kermany, Boaz Carmeli, Ofer Lavi, George Kour, and Alon
-% 	Jacovi.
-% 	\newblock Estimate and replace: A novel approach to integrating deep neural
-% 	networks with existing applications.
-% 	\newblock {\em arXiv preprint arXiv:1804.09028}, 2018.
-
-% \end{thebibliography}
+\bibliographystyle{unsrt}
+\bibliography{references}
 

### 2023-04-02 paper_2023/paper.tex
wip

@@
 
-To address the aforementioned challenges, the paper introduces a framework termed Colonies designed to facilitate the execution of computational workloads across a diverse range of platforms.
+To address the aforementioned challenges, the paper introduces a framework termed Colonies which is available as open-source\footnote{https://github.com/colonyos/colonies} and designed to facilitate execution of computational workloads across a diverse range of platforms.
 Colonies is founded upon a loosely-coupled microservice architecture that beaks down complex workflows into composable functions.
@@
 cloud, edge, devices, or even in web browsers.
-Colonies orchestrates the execution and by using a zereo-trust security protocol, a collection of distributed executors can function as a single cohesive unit, thereby establishing a seamless compute continuum across multiple platforms.
+Colonies orchestrates the execution and by using a zereo-trust security protocol, a collection of distributed executors can function as a single cohesive unit, thereby establishing seamless compute continuums across multiple platforms.
 
@@
 As most models require regular re-training or re-calibration, it must be possible to update deployed models and software seamlessly without losing information or breaking the system.
-In many cases, there is a constant flow of data ingested into the system which must be managed even if some components are malfunctioning.
-This becomes even more challenging when nodes or parts of the underlying infrastructure become unavailable due to maintenance or failaure such as software updates, hardware replacements or simply missconfiguration.
+In many cases, there is a constant flow of data ingested into the system which must be managed even in case of failures.
+This becomes even more challenging when nodes or parts of the underlying infrastructure become unavailable due to maintenance such as software updates, hardware replacements and sometimes missconfiguration failures.
 
@@
 In reality, AI system requires integration of multiple systems.
-For instance, data may need to be captured from an IoT system or pulled from third-party database running on different domains than the compute cluster.
+For instance, data need to be captured from an IoT system or pulled from third-party database running on different domains than the compute cluster.
 With the emergence of edge computing, parts of a data pipeline may also run on edge servers to bring computations closer to data sources.
@@
 These compute units can be implemented in any computer language and be deployed anywhere on the Internet.
+The remainder of the paper describes the Colonies framework and how it can be used to create robust and scalable AI systems.
 
 \section{Related work}
-A workflow is a series of tasks that need to be completed in a specific order.
-Workflows are often represented as directed acyclic graphs (DAGs), where nodes represent tasks and edges represent dependencies or data flow between tasks.
-Workflow management has been extensively studied in both academic and industrial settings, with numerous approaches proposed to address the challenges in this field.
+Workflow management has been extensively studied in both academic and industrial settings \cite{apache_airflow}, with numerous approaches proposed to address the challenges in this field.
+For instance, Apache Airflow \cite{apache_airflow} is a popular open-source workflow management system for handling data engineering pipelines.
+Like Colonies, Apache Airflow enables developers to create custom operators and executors that can be integrated with various systems.
+Additionally, Apache Airflow offers an HTTP API that allows for the development of clients and software development kits (SDKs) in multiple programming languages.
 
-Another popular approach is the use of workflow engines, which provide a framework for creating and executing workflows.
-Workflow engines can be integrated with other tools and systems to automate complex business processes.
-Examples of workflow engines include Apache Airflow, Camunda, and Zeebe.
+However, Apache Airflow is not designed around a queuing system.
+Instead, it must be integrated with a message broker, such as RabbitMQ \cite{rabbitmq} or Kafka \cite{apache_kafka}, to implement task queues, resulting in a more complex architecture than Colonies.
+Furthermore, Colonies is based on a distributed microservice architecture that makes it more suitable for DevOps software development.
+As Colonies is loosely coupled, executors can be implemented and dynamically deployed without reconfiguring the workflow engine, effectively transforming it into a grid computing platform.
 
-Apache Airflow is a popular open-source workflow management system used to programmatically author, schedule, and monitor workflows.
-Luigi is also an open-source workflow management system developed by Spotify. It is designed to handle long-running batch jobs and data pipelines.
-
-Pinball is a workflow management system developed by Pinterest that allows users to define and execute workflows using Python scripts.
-It has a simple architecture, with support for DAGs and task dependencies.
-However, it lacks some of the advanced features of Airflow, such as task retries and monitoring.
-
-TODO:
+When it comes to academic works, \cite{GarciaRepresa1740746} t
 
@@
 
-Queues enable different software components to communicate indirectly by acting as a buffer between them.
+Queues enable software services to communicate indirectly by acting as a buffer between them.
 Queues makes it possible to decouple each executor and make them operate independently, e.g.
@@
 Generarally, the only way to retrieve a specific message is to pull all messages from the queue, obtain the message, and then place all remaining messages back into the queue in the same order.
-In contrast, using a database a query can match any columns and be ordered using a timestamp to make it function as a queue.
+In contrast, a database can also function as a queue and a query can match any columns thus making it possible to assign specific executors to processes.
 
 \begin{equation}
+    \label{eq:pt}
     priority_{time}=priority \cdot -10^9 \cdot 60 \cdot 60 \cdot 24 + submission_{time}  
 \end{equation}
+
+Implementing queue functionality using a database can be achieved in SQL by utilizing the \emph{order by timestamp} and \emph{limit 1} clauses to select tasks according to their submission times.
+Upon submission of a task, a \emph{priority time} value is computed and stored in the database.
+The primary objective of the priority time is to modify the submission timestamp by subtracting a delta time calculated as a function a priority value.
+This enables prioritization of more critical tasks.
+Equation \ref{eq:pt} shows how the priority time is calculated.
 
@@
 Once submitted, other executors can connect to the server to receive execution assignments.
-When a function specification is submitted to the Colonies server, it is store as a process entry in a dedicated process table database.
-In this case, a process is analogous to a task, but contains additional contextual information about the execution, for example, execution status (waiting, running, successful, failed), assigned executor Id, submission time, execution time, deadlines, priority etc.
+When a function specification is submitted to the Colonies server, it is store as a process entry in a process table database.
+In this case, a process is similar to a task or a job, but also contains contextual information about the execution, for example, execution status (waiting, running, successful, failed), assigned executor ID, submission time, execution time, deadlines, priority etc.
 It also contains input and argument values as well as return values from the function innovation.
+In the remainder of the paper, we are going to use the term process instead of task or job.
 
@@
 Rather, it is the responsibility of the executors to connect to the Colonies server.
-This enables executors to be deployed anywhere on the Internet, even behind firewalls, commercial telco networks, or in application running in web browsers.
+This enables executors to be deployed anywhere on the Internet, behind firewalls, commercial telco networks, or even in web browser-based applications.
 
 Figure \ref{fig:process_assignment} illustrates an executor submitting a function specification that is later assigned to another executor.
-Executors have to specify to the Colonies server which functions they are capable of executing.
+When registering, executors have to specify to the Colonies server which functions they are capable of executing.
 The Colonies server makes sure that the conditions (i.e requirmenets) of a function specification matches the capability of an executor.
-
 \begin{figure}[h]
@@
 
+\subsection{A stateless failsafe mechanism}
+The primary objective of all Colonies API requests is to alter a state stored in the database or retrieve information from the database.
+The Colonies framework is designed to be stateless, meaning that the Colonies server does not keep any information between requests.
+In other words, each request is handled independently, without relying on any information from previous requests.
+
 Figure \ref{fig:function_spec} shows an example of a function specification.
@@
 If such process is detected, it is reset, allowing it to be re-assigned to other executors.
-This scheme is completely stateless, so if the Colonies server crashes, another Colonies server can take over and look for processes that have exceeded their deadline.
-Moreover, if an executor assigned to a process crashes, the process is simply re-executed by another executor.
 
@@
 One solution is to select a set of executors to be removed and then starv them out by denying them new process assignments.
-Another, simpler solution, is to immediately destroy the executors, or if possible notify them in advance so that they can gracefully unregister.
+Another, simpler solution, is to immediately destroy the executors and use the \emph{maxexectime} failsafe mechanism move back failed processes to the queue.
 
@@
 
-Utilizing the \emph{maxexectime} failsafe mechanism not only enhances system reliability, but also provides an opportunity to apply of Chaos engineering \cite{chaos_engineering}.
-For example, deliberately terminating executors using a Chaos monkey.
-If executors are deployed on Kubernetes, Kubernetes will automatically redeploy terminated executors.
+Utilizing the \emph{maxexectime} failsafe mechanism not only enhances system reliability, but also provides an opportunity to apply Chaos engineering \cite{chaos_engineering}.
+For example, a Chaos monkey can be used to deliberately terminate executors.
+If executors are deployed on Kubernetes, Kubernetes will then automatically redeploy terminated executors.
 The constant replacement of executor instances ensures that the system is capable of gracefully tolerating failures.
@@
 
-The primary objective of all Colonies API requests is to either alter a state stored in the database or retrieve information from it.
-The Colonies framework is designed to be stateless, meaning that the Colonies server does not keep any information between requests.
-In other words, each request is handled independently, without relying on any information from previous requests.
-
-One example of a request is the assign request, which allocates a waiting process from the database to an executor.
+The assign request allocates binds a process from the database to an executor.
 Given the multi-threaded nature of the Colonies server, it is essential that the assign request is synchronized to ensure that only one thread at a time can modify the database and update the process table, thus preventing multiple executors from being assigned to the same process.
 As only the assigned executor can modify a running process and only one executor must be assigned to a process, synchronization is required for the assign request.
-It is worth noting that synchronization is not necessary for other requests, such as the submit request or the close request.
-As the submit request only add new entries to the process table, there is no race conditions.
+It is worth noting that synchronization is not necessary for other requests.
+For example, as the submit request only add new entries to the process table, there is no race conditions.
 The close request set the output of the function innovation and updates the process state to either successful or failed in the process table.
-Since there exists only one assigned executor there is no need for synchronization.
+Since there is only one assigned executor to a process there is no race conditions and consequently no need for synchronization.
 
@@
 \subsubsection{Workflows}
-TODO
+A workflow is a series of tasks that need to be completed in a specific order.
+Workflows are often represented as directed acyclic graphs (DAGs), where nodes represent tasks and edges represent dependencies or data flow between tasks.
+
 \begin{figure}[h]

### 2023-04-02 paper_2023/paper.tex
wip

@@
 \section{Related work}
-Workflow management has been extensively studied in both academic and industrial settings \cite{apache_airflow}, with numerous approaches proposed to address the challenges in this field.
-For instance, Apache Airflow \cite{apache_airflow} is a popular open-source workflow management system for handling data engineering pipelines.
+Workflow management has been extensively studied in both academic and industrial settings with numerous approaches proposed to address the challenges in this field.
+For example, Apache Airflow \cite{apache_airflow} is a popular open-source workflow management system for handling data engineering pipelines.
 Like Colonies, Apache Airflow enables developers to create custom operators and executors that can be integrated with various systems.
-Additionally, Apache Airflow offers an HTTP API that allows for the development of clients and software development kits (SDKs) in multiple programming languages.
-
-However, Apache Airflow is not designed around a queuing system.
+Additionally, Apache Airflow offers an HTTP API that makes it possible to develop software development kits (SDKs) in various programming languages.
+However, Apache Airflow does not realy on a queuing system.
 Instead, it must be integrated with a message broker, such as RabbitMQ \cite{rabbitmq} or Kafka \cite{apache_kafka}, to implement task queues, resulting in a more complex architecture than Colonies.
 Furthermore, Colonies is based on a distributed microservice architecture that makes it more suitable for DevOps software development.
-As Colonies is loosely coupled, executors can be implemented and dynamically deployed without reconfiguring the workflow engine, effectively transforming it into a grid computing platform.
+As Colonies is loosely coupled, executors can be implemented and dynamically deployed without reconfiguring the workflow engine, effectively making it work like a grid computing platform.
 
-When it comes to academic works, \cite{GarciaRepresa1740746} t
+Argo is a ...
+
+OpenWolf \cite{openwolf} is a serverless workflow engine designed to utilize the Function-as-a-Service (FaaS) paradigm for composing complex scientific workflows.
+It is based on OpenFaaS \cite{openfaas}, which allows functions to run on Kubernetes clusters.
+In contrast, Colonies does not require Kubernetes, but provides a zero-trust security protocol to allow functions to be executed securely by distributed executor workers deployed anywhere on the Internet.
+
+The serverless workflow project \cite{serverlessworkflow} aims at providing a vendor-neutral workflow DSL.
+Synapse is a workflow management system implementing the serverless workflow specification.
+It is important to point out that Colonies does not provide an infrastructure for function execution.
+Instead, Colonies primary role is to serve as a platform for coordinating function executions, which are carried out by distributed executor services.
+
+Colonies is based on a microservice architecture.
+J.
+Represa et al.
+\cite{GarciaRepresa1740746} discusses various challenges developing microservice-based workflow management for industrial automation.
+They concludes that microservice-based workflow technologies is feasible for industrial usage, especially considering its flexibility.
+The main contribution of this paper is a technica
+
+SCAFE: A Service-Centered Cloud-Native Workflow Engine Architecture
+
+https://www.alibabacloud.com/product/serverless-workflow
+https://serverlessworkflow.io/
+
+https://github.com/serverlessworkflow/synapse
+
+AWS Step Functions to design and run a serverless workflow that coordinates multiple AWS Lambda functions.
 
@@
 
-Queues can be implemented in different ways, and while message brokers are a common solution, Colonies has adopted an alternative strategy by leveraging a standard database and querying it for tasks to assign to different executors.
+Queues can be implemented in different ways, and while message brokers are a common solution, Colonies adopts an alternative strategy and leverage a standard database and querying it for tasks to assign to different executors.
 One key advantage of this approach is that it enables fine-grained task assignments, making it possible to assign specific tasks to particular executors.
@@
 
-Incorporating Raft into the Colonies framework allows incoming assign requests to be directed towards the leading Colonies server, ensuring that only one Colonies server replica handles these requests.
+Integrating Raft with the Colonies framework allows incoming assign requests to be directed towards the leading Colonies server, thereby ensuring that only one Colonies server replica handles such requests.
 Figure \ref{fig:ha_deployment} shows an overview of a high-availability Colonies deployment.

### 2023-04-02 paper_2023/paper.tex
wip

@@
     \label{eq:pt}
-    priority_{time}=priority \cdot -10^9 \cdot 60 \cdot 60 \cdot 24 + submission_{time}  
+    priority_{time}=submission_{time} - priority \cdot 10^9 \cdot 60 \cdot 60 \cdot 24
 \end{equation}
@@
 This enables prioritization of more critical tasks.
-Equation \ref{eq:pt} shows how the priority time is calculated.
+Equation \ref{eq:pt} shows how the priority time is calculated for a nanosecond timestamp.
 
@@
 The Colonies server makes sure that the conditions (i.e requirmenets) of a function specification matches the capability of an executor.
+Note that the executors are responsibility of interpreting the assigned process and function specification to execute relevant code.
+
 \begin{figure}[h]

### 2023-04-03 paper_2023/paper.tex
wip

@@
 % Uncomment to override  the `A preprint' in the header
-%\renewcommand{\headeright}{Technical Report}
-%\renewcommand{\undertitle}{Technical Report}
-\renewcommand{\shorttitle}{\textit{arXiv} Template}
+\renewcommand{\headeright}{Technical Report}
+\renewcommand{\undertitle}{Technical Report}
+%\renewcommand{\shorttitle}{\textit{arXiv} Template}
 
@@
 
-Argo is a ...
+Argo \cite{argowf} is an open source container-native workflow engine for orchestrating parallel jobs on Kubernetes.
+It is can be used for running CI/CD pipelines or compute intensive machine learning or data processing tasks where each job runs as a container.
+In contrast, Colonies offers a more versatile approach, allowing jobs to be launched within a container.
+As launching new containers on Kubernetes can occasionally be time-consuming, Colonies can delivers higher throughput as the costs of starting new jobs are minimal.
+This is particular useful when launching large container images, or workloads (e.g.
+Julia scripts) or that takes relative long time to start.
 
@@
 
-The serverless workflow project \cite{serverlessworkflow} aims at providing a vendor-neutral workflow DSL.
-Synapse is a workflow management system implementing the serverless workflow specification.
+The serverless workflow project \cite{serverlessworkflows} aims at providing a vendor-neutral workflow DSL.
+Synapse \cite{synapse} is a workflow management system similar to Argo, but that implements the serverless workflow specification.
+A similar engine is proposed in \cite{scafe}.
 It is important to point out that Colonies does not provide an infrastructure for function execution.
-Instead, Colonies primary role is to serve as a platform for coordinating function executions, which are carried out by distributed executor services.
+Instead, Colonies primary role is to serve as a platform for coordinating function executions, which are carried out by distributed executor.
 
-Colonies is based on a microservice architecture.
 J.
 Represa et al.
-\cite{GarciaRepresa1740746} discusses various challenges developing microservice-based workflow management for industrial automation.
-They concludes that microservice-based workflow technologies is feasible for industrial usage, especially considering its flexibility.
-The main contribution of this paper is a technica
-
-SCAFE: A Service-Centered Cloud-Native Workflow Engine Architecture
-
-https://www.alibabacloud.com/product/serverless-workflow
-https://serverlessworkflow.io/
-
-https://github.com/serverlessworkflow/synapse
-
-AWS Step Functions to design and run a serverless workflow that coordinates multiple AWS Lambda functions.
+\cite{GarciaRepresa1740746} explore various challenges associated with developing microservice-based workflow management for industrial automation within the context of the Arrowhead project.
+The authors conclude that microservice-based workflow technologies is viable for industrial applications, particularly due to their inherent flexibility.
+The primary contribution of this paper is a comprehensive technical description of how to implement a distributed workflow engine based on microservice principles, extending beyond orchestrating microservices to execute automation tasks.
 
@@
 
-Integrating Raft with the Colonies framework allows incoming assign requests to be directed towards the leading Colonies server, thereby ensuring that only one Colonies server replica handles such requests.
+Incorporting Raft with the Colonies framework allows incoming assign requests to be directed towards the leading Colonies server, thereby ensuring that only one Colonies server replica handles such requests.
 Figure \ref{fig:ha_deployment} shows an overview of a high-availability Colonies deployment.

### 2023-04-04 paper_2023/paper.tex
wip

@@
 To address the aforementioned challenges, the paper introduces a framework termed Colonies which is available as open-source\footnote{https://github.com/colonyos/colonies} and designed to facilitate execution of computational workloads across a diverse range of platforms.
-Colonies is founded upon a loosely-coupled microservice architecture that beaks down complex workflows into composable functions.
+It can be used to implement grid-alike distributed virtual computer systems, and is founded upon a loosely-coupled microservice architecture that beaks down complex workflows into composable functions.
 With the use of an HTTP protocol, these composable functions can be implemented in any computer language and be executed by independent executors deployed across various systems, e.g.
@@
 
-In summary, Colonies is a highly versatile and scalable framework that can streamline the development and deployment of computational workloads across heterogeneous platforms while also ensuring full traceability and zero-trust security.
+In summary, Colonies is a highly versatile and scalable framework that can streamline development and deployment of computational workloads across heterogeneous platforms while ensuring full traceability and zero-trust security.
 \end{abstract}
@@
 In many cases, there is a constant flow of data ingested into the system which must be managed even in case of failures.
-This becomes even more challenging when nodes or parts of the underlying infrastructure become unavailable due to maintenance such as software updates, hardware replacements and sometimes missconfiguration failures.
+This becomes even more challenging when nodes or parts of the underlying infrastructure become unavailable due to maintenance such as software updates, hardware replacements and sometimes missconfiguration problems.
 
@@
 
-This paper presents a framework called Colonies, which is built around a loosely-coupled microservices architecture that separates workflow definitions from implementation and deployment.
+This paper presents a framework called Colonies, specifically designed for implementing distributed virtual computer systems that facilitate integration with other systems.
+Colonies is built around a loosely-coupled microservices architecture that separates workflow definitions from implementation and deployment, offering a modern approach to grid computing.
 The main objective is to create a tool where monolithic workflows can be broken down into independent and separated compute units that can be dynamically added or removed while executing workflows.
-These compute units can be implemented in any computer language and be deployed anywhere on the Internet.
+These compute units can be implemented in any computer language and easily integrated with other systems, and then be deployed anywhere on the Internet.
 The remainder of the paper describes the Colonies framework and how it can be used to create robust and scalable AI systems.
@@
 Furthermore, Colonies is based on a distributed microservice architecture that makes it more suitable for DevOps software development.
-As Colonies is loosely coupled, executors can be implemented and dynamically deployed without reconfiguring the workflow engine, effectively making it work like a grid computing platform.
+As Colonies is loosely coupled, executors can be implemented and dynamically deployed without reconfiguring the workflow engine, effectively turning it to a grid computing platform.
 
@@
 The primary contribution of this paper is a comprehensive technical description of how to implement a distributed workflow engine based on microservice principles, extending beyond orchestrating microservices to execute automation tasks.
+
+Grid computing is a distributed computing model that allows multiple computers, sometimes geographically distributed, to work together as a cohesive system to solve large-scale computational problems.
+It can be used to harness resources from numerous computers to solve complex tasks.
 

### 2023-04-04 paper_2023/paper.tex
wip

@@
 
-To address the aforementioned challenges, the paper introduces a framework termed Colonies which is available as open-source\footnote{https://github.com/colonyos/colonies} and designed to facilitate execution of computational workloads across a diverse range of platforms.
-It can be used to implement grid-alike distributed virtual computer systems, and is founded upon a loosely-coupled microservice architecture that beaks down complex workflows into composable functions.
-With the use of an HTTP protocol, these composable functions can be implemented in any computer language and be executed by independent executors deployed across various systems, e.g.
-cloud, edge, devices, or even in web browsers.
-Colonies orchestrates the execution and by using a zereo-trust security protocol, a collection of distributed executors can function as a single cohesive unit, thereby establishing seamless compute continuums across multiple platforms.
+To address the aforementioned challenges, the paper introduces an open-source framework called Colonies\footnote{https://github.com/colonyos/colonies}, which is designed to facilitate execution of computational workloads across a diverse range of platforms.
+Colonies enables the development of a \emph{grid-like distributed virtual computer}, referred to as a \emph{colony}, which can effortlessly integrate with any third-party application or other workflow management systems.
+
+Built on a modular microservice architecture, Colonies effectively decomposes complex workflows into highly composable functions.
+With the use of an HTTP protocol, these composable functions can be implemented in any computer language, and be executed by independent executors deployed across various systems, e.g.
+cloud, edge, high-performance computing (HPC) environments, devices, or even web browsers.
+By using a zero-trust security protocol, a collection of distributed executors can function as a single cohesive unit, thereby establishing seamless compute continuums across multiple platforms.
 
 In addition to a technical description of the Colonies framework, the paper also describes some potential use cases.
-The paper describe how Colonies can be leveraged to build a remote sensing platform on Kubernetes, serve as a building block for edge computing, implement a serverless FaaS (Function-as-a-Service), and how it can be integrated with HPC platforms.
+The paper describe how Colonies can be leveraged to build a remote sensing platform on Kubernetes, serve as a building block for edge computing, implement a serverless FaaS (Function-as-a-Service), and how it can be integrated with the Slurm workload manager.
 Finally, the paper presents a performance investigation, as well as scalability and robustness evaluation.
@@
 Sometimes it is necessary to temporarily increase the capacity of on-prem clusters by combining resources from multiple providers, for example, adding cloud compute resources to handle peak loads or utilize HPC resources to quickly reprocess historical data.
-Developing hybrid workflows where some jobs run in the cloud and others run on HPC systems requires even more software development efforts and is beyond the scope of most users, preventing them from utilizing powerful hardware.
-Clearly, there is a need for a framework that can consolidate various platforms to simplify development and enable seamless execution across platforms.
+Developing hybrid workflows where some jobs run in the cloud and others run on HPC systems requires even more software development efforts \cite{wf_challenges} and is beyond the scope of many users, preventing them from utilizing powerful hardware.
+Clearly, there is a need for a framework that can consolidate various workflow management platforms to simplify development and enable seamless execution across platforms.
 
-This paper presents a framework called Colonies, specifically designed for implementing distributed virtual computer systems that facilitate integration with other systems.
-Colonies is built around a loosely-coupled microservices architecture that separates workflow definitions from implementation and deployment, offering a modern approach to grid computing.
-The main objective is to create a tool where monolithic workflows can be broken down into independent and separated compute units that can be dynamically added or removed while executing workflows.
-These compute units can be implemented in any computer language and easily integrated with other systems, and then be deployed anywhere on the Internet.
+This paper presents a framework called Colonies, specifically designed to implement a distributed virtual computer system.
+Colonies is built around a loosely-coupled microservices architecture that separates workflow definitions from implementation and deployment.
+The primary objective is to establish a platform where monolithic workflows can be decomposed into independent computational units that can easily integrate with other systems or applications.
+These compute units can be implemented in any computer language and be deployed anywhere on the Internet.
 The remainder of the paper describes the Colonies framework and how it can be used to create robust and scalable AI systems.
@@
 \section{Related work}
-Workflow management has been extensively studied in both academic and industrial settings with numerous approaches proposed to address the challenges in this field.
-For example, Apache Airflow \cite{apache_airflow} is a popular open-source workflow management system for handling data engineering pipelines.
+Workflow management has been extensively studied in both academic and industrial settings with numerous approaches \cite{service_wfs, schmitt2022workflow, GarciaRepresa1740746, Ouyang2010, NIKOLOV2021100440, workflow_in_bigdata} proposed to address the challenges in this field.
+Recently, Apache Airflow \cite{apache_airflow} has become a popular open-source workflow management system for handling data engineering pipelines.
 Like Colonies, Apache Airflow enables developers to create custom operators and executors that can be integrated with various systems.
@@
 It is can be used for running CI/CD pipelines or compute intensive machine learning or data processing tasks where each job runs as a container.
-In contrast, Colonies offers a more versatile approach, allowing jobs to be launched within a container.
+In contrast, Colonies offers a more versatile approach, allowing jobs to be launched within an already started container.
 As launching new containers on Kubernetes can occasionally be time-consuming, Colonies can delivers higher throughput as the costs of starting new jobs are minimal.
@@
 The authors conclude that microservice-based workflow technologies is viable for industrial applications, particularly due to their inherent flexibility.
-The primary contribution of this paper is a comprehensive technical description of how to implement a distributed workflow engine based on microservice principles, extending beyond orchestrating microservices to execute automation tasks.
+The primary contribution of this paper is a comprehensive technical description of how to implement a distributed workflow engine based on microservice principles, extending beyond orchestrating microservices to execute automation tasks within one platform.
+The vision is to create a distributed architecuture, where microservices can reside anywhere on the Internet and still function as a cohesive unit to execute cross-platform workflows.
 
-Grid computing is a distributed computing model that allows multiple computers, sometimes geographically distributed, to work together as a cohesive system to solve large-scale computational problems.
-It can be used to harness resources from numerous computers to solve complex tasks.
+Grid computing is a distributed computing model that allows multiple computers, sometimes geographically distributed, to work together to solve large-scale computational problems.
+It can be used to harness compute resources from numerous computers to solve complex tasks.
+To the authors best knowledge, no one has ever used to grid platform as an integration platform to orchestrate AI workload across platforms.
+The main contribution of this paper is a technical description of the Colonies framework and a description of potential use cases.
 

### 2023-04-05 paper_2023/paper.tex
wip

@@
 
-Built on a modular microservice architecture, Colonies effectively decomposes complex workflows into highly composable functions.
-With the use of an HTTP protocol, these composable functions can be implemented in any computer language, and be executed by independent executors deployed across various systems, e.g.
-cloud, edge, high-performance computing (HPC) environments, devices, or even web browsers.
-By using a zero-trust security protocol, a collection of distributed executors can function as a single cohesive unit, thereby establishing seamless compute continuums across multiple platforms.
+Built upon a modular microservice architecture, Colonies enables complex workflows to be broken down into composable functions.
+With the use of an HTTP protocol, these composable functions can be implemented in any programming language and be executed by independent executors deployed across various platforms on the Internet, such as cloud, edge, high-performance computing (HPC) environments, devices, or even web browsers.
+With the integration of a zero-trust security protocol, a collection of distributed executors can operate as a unified entity, thus establishing seamless computational continuums across multiple platforms.
 
@@
 
-This paper presents a framework called Colonies, specifically designed to implement a distributed virtual computer system.
+This paper presents a framework called Colonies, specifically designed to implement a \emph{distributed virtual computer system} that can integrate any third-party system.
 Colonies is built around a loosely-coupled microservices architecture that separates workflow definitions from implementation and deployment.
@@
 Furthermore, Colonies is based on a distributed microservice architecture that makes it more suitable for DevOps software development.
-As Colonies is loosely coupled, executors can be implemented and dynamically deployed without reconfiguring the workflow engine, effectively turning it to a grid computing platform.
+As Colonies is loosely coupled, executors can be implemented and dynamically deployed without reconfiguring the workflow engine.
 
@@
 Represa et al.
-\cite{GarciaRepresa1740746} explore various challenges associated with developing microservice-based workflow management for industrial automation within the context of the Arrowhead project.
+\cite{GarciaRepresa1740746, GarciaRepresa1640771} explore various challenges associated with developing microservice-based workflow management for industrial automation within the context of the Arrowhead project \cite{delsing2017iot}.
 The authors conclude that microservice-based workflow technologies is viable for industrial applications, particularly due to their inherent flexibility.
@@
 
-Grid computing is a distributed computing model that allows multiple computers, sometimes geographically distributed, to work together to solve large-scale computational problems.
-It can be used to harness compute resources from numerous computers to solve complex tasks.
-To the authors best knowledge, no one has ever used to grid platform as an integration platform to orchestrate AI workload across platforms.
-The main contribution of this paper is a technical description of the Colonies framework and a description of potential use cases.
+Grid computing \cite{grid_computing} is a distributed computing model that allows multiple computers, which may be geographically dispersed, to collaborate in addressing large-scale computational challenges.
+The Colonies framework is founded upon a grid computing model, with the primary purpose of assigning tasks to distributed executors.
+To the best of the authors' knowledge, no previous work has integrated a microservice-oriented architecture with a grid computing model to serve as an integration point for coordinating artificial intelligence workloads across a diverse range of platforms.
+Additionally, Colonies is designed to function as a ledger, offering complete transparency and execution history, which is essential for implementing zero-trust security.
+
+\begin{figure}[h]
+	\centering
+    \includegraphics[scale=0.35]{overview.png}
+	\caption{Overview of ....}
+	\label{fig:function_spec}
+\end{figure}
 

### 2023-04-05 paper_2023/paper.tex
wip

@@
 To address the aforementioned challenges, the paper introduces an open-source framework called Colonies\footnote{https://github.com/colonyos/colonies}, which is designed to facilitate execution of computational workloads across a diverse range of platforms.
-Colonies enables the development of a \emph{grid-like distributed virtual computer}, referred to as a \emph{colony}, which can effortlessly integrate with any third-party application or other workflow management systems.
+Colonies enables the development of a \emph{grid-like distributed virtual computer}, called a \emph{Colony}, which can effortlessly integrate with any third-party application or other workflow management systems.
 
@@
 Developing robust and scalable AI systems is a challenging task that requires deep understanding in several fields.
-To begin with, an AI model must be trained which requires knowledge in advanced statistics or machine learning, as well as access to training and validation data.
-Typically, this data must be pre-processed through various stages before it can be utilized.
-Although it may be practical for small-scale projects to run the entire training processes on local development computers, larger AI models typically require access to powerful compute clusters or even HPC systems.
+To begin with, an AI model must be trained which requires knowledge in advanced statistics or machine learning.
+Typically, training and validation data must be pre-processed through various stages before it can be utilized.
+Although it may be practical for small-scale projects to run the entire training processes on local development computers, larger AI models typically require access to powerful compute clusters or even  high-performance computing (HPC) systems.
 Manual use of such infrastructure can be laborious and time-consuming.
@@
 
-This paper presents a framework called Colonies, specifically designed to implement a \emph{distributed virtual computer system} that can integrate any third-party system.
+This paper presents a framework called Colonies, specifically designed to implement a \emph{distributed virtual computer system} that can easily integrate wth any third-party system.
 Colonies is built around a loosely-coupled microservices architecture that separates workflow definitions from implementation and deployment.
@@
 Furthermore, Colonies is based on a distributed microservice architecture that makes it more suitable for DevOps software development.
-As Colonies is loosely coupled, executors can be implemented and dynamically deployed without reconfiguring the workflow engine.
+As Colonies is loosely coupled, executors can be implemented and dynamically deployed on the Internet without reconfiguring the workflow engine.
 
@@
 This is particular useful when launching large container images, or workloads (e.g.
-Julia scripts) or that takes relative long time to start.
+Julia scripts) that takes relative long time to start.
 
@@
 It is important to point out that Colonies does not provide an infrastructure for function execution.
-Instead, Colonies primary role is to serve as a platform for coordinating function executions, which are carried out by distributed executor.
+Instead, Colonies primary role is to serve as a platform for coordinating function executions which are carried out by distributed executor.
 
@@
 The primary contribution of this paper is a comprehensive technical description of how to implement a distributed workflow engine based on microservice principles, extending beyond orchestrating microservices to execute automation tasks within one platform.
-The vision is to create a distributed architecuture, where microservices can reside anywhere on the Internet and still function as a cohesive unit to execute cross-platform workflows.
+The vision is to create a distributed architecuture where microservices can reside anywhere on the Internet and still function as a cohesive unit to execute cross-platform workflows.
 
 Grid computing \cite{grid_computing} is a distributed computing model that allows multiple computers, which may be geographically dispersed, to collaborate in addressing large-scale computational challenges.
-The Colonies framework is founded upon a grid computing model, with the primary purpose of assigning tasks to distributed executors.
+The Colonies framework is founded upon a grid computing model with the primary purpose of assigning tasks to distributed executors.
 To the best of the authors' knowledge, no previous work has integrated a microservice-oriented architecture with a grid computing model to serve as an integration point for coordinating artificial intelligence workloads across a diverse range of platforms.
@@
     \includegraphics[scale=0.35]{overview.png}
-	\caption{Overview of ....}
-	\label{fig:function_spec}
+	\caption{Overview of the Colonies framework. Executors may be deployed anywhere on the Internet.}
+	\label{fig:overview}
 \end{figure}
@@
 \label{sec:headings}
+A core concept of the Colonies framework is the notion of \emph{processes}.
+A process contains meta-information about computations executed by a remote computer programs, referred to as a \emph{distributed executors}.
+Specifically, a process contains a definition of a function to be executed, as well as contextual information such as execution status, including the result of the computation.
+One can think of a process as a digital twin of a computing operation.
+It is essential to note that a process does not necessarily have to be an operating system process; it can represent any type of computation, e.g.
+a remote function call executed by any kind of software.
+
+Figure 1 depicts an overview of the components part of the Colonies framework.
+The \emph{Colonies servers} form the backbone of the framework, functioning as a centralized control system for task submission and assignment.
+All execution information and history are stored in a database maintained by the Colonies servers.
+Upon submission, a process is stored in the database.
+When it assigned to an executor its execution status is changed from \emph{waiting} to \emph{running}.
+Consequently, it is possible to submit a process that should run in the future, even if no executors can run the process at the moment.
+
+\emph{Distributed executors} are responsible for executing processes assigned by the Colonies Servers.
+Several distributed executors form a \emph{Colony}, which is a collection of executors operating as a cohesive unit, where each executor is responsible for executing a specific type of process.
+One could view a Colony as an organization of distributed computer software acting a single virtual computer system.
+To interact with other executors, executors must prove their Colony membership using a cryptographic protocol that follows the zero-trust security principle \emph{never trust, always verify}.
+This security protocol ensures that users can keep control even when executors are distributed across plattforms, e.g.
+some  XXX
+
+The remainder of this section outlines the underlying design principles and offers a deeper technical description of the framework. The next section describes how
+
+\subsection{Microservices}
 Microservices is an architectural design pattern in which an application is structured as a collection of small, independently deployable, and loosely coupled services that communicate with other microservices through a well-defined API.
@@
 Generarally, the only way to retrieve a specific message is to pull all messages from the queue, obtain the message, and then place all remaining messages back into the queue in the same order.
-In contrast, a database can also function as a queue and a query can match any columns thus making it possible to assign specific executors to processes.
+In contrast, a database can function as a queue and a query can match any columns thus making it possible to assign specific executors to specific processes.
 
@@
 The Raft protocol also enables seamless updates to the Colonies server software by making it possible to upgrade each replica individually.
-As a result, Colonies becomes well-suited for Kubernetes deployments.
+Consequently, Colonies is well-suited for Kubernetes deployments, ensuring high levels of availability and fault tolerance.
 

### 2023-04-07 paper_2023/paper.tex
wip

@@
 Built upon a modular microservice architecture, Colonies enables complex workflows to be broken down into composable functions.
-With the use of an HTTP protocol, these composable functions can be implemented in any programming language and be executed by independent executors deployed across various platforms on the Internet, such as cloud, edge, high-performance computing (HPC) environments, devices, or even web browsers.
-With the integration of a zero-trust security protocol, a collection of distributed executors can operate as a unified entity, thus establishing seamless computational continuums across multiple platforms.
+With the use of an HTTP protocol, these composable functions can be implemented in any programming language and be executed by independent executors deployed across platforms, anywhere on the Internet, including cloud, edge, high-performance computing (HPC) environments, devices, or even web browsers.
+A zero-trust security protocol enables a collection of distributed executors to operate as a unified entity, thus establishing seamless computational continuums across multiple platforms.
 
@@
 % keywords can be removed
-\keywords{Serverless computing \and Parallel computing \and Workflow orchestration}
+\keywords{Distributed computing, serverless computing \and parallel computing \and Workflow orchestration}
 
@@
 
-Taking an AI model into production requires substantial software engineering expertise.
+Taking an AI model into production requires substantial software engineering expertise and collaboration across teams.
 In contrast to traditional IT workloads, both the data and the model must be managed in addition to the software itself.
@@
 In reality, AI system requires integration of multiple systems.
-For instance, data need to be captured from an IoT system or pulled from third-party database running on different domains than the compute cluster.
+For instance, data need to be captured from an IoT system or pulled from third-party database running on different domains than the compute cluster itself.
 With the emergence of edge computing, parts of a data pipeline may also run on edge servers to bring computations closer to data sources.
@@
 
-This paper presents a framework called Colonies, specifically designed to implement a \emph{distributed virtual computer system} that can easily integrate wth any third-party system.
+This paper presents a framework called Colonies, specifically designed to implement a \emph{distributed virtual computer system} that can easily integrate with any third-party system.
 Colonies is built around a loosely-coupled microservices architecture that separates workflow definitions from implementation and deployment.
-The primary objective is to establish a platform where monolithic workflows can be decomposed into independent computational units that can easily integrate with other systems or applications.
-These compute units can be implemented in any computer language and be deployed anywhere on the Internet.
-The remainder of the paper describes the Colonies framework and how it can be used to create robust and scalable AI systems.
+The primary objective is to establish a platform where monolithic workflows can be decomposed into independent microservices that can easily be integrated with other systems or applications running anywhere on the Internet.
+The remainder of the paper describes the Colonies framework and how it can be used to create robust and scalable AI systems across plattforms.
 
@@
 Additionally, Apache Airflow offers an HTTP API that makes it possible to develop software development kits (SDKs) in various programming languages.
-However, Apache Airflow does not realy on a queuing system.
+However, Apache Airflow does not rely on a queuing system.
 Instead, it must be integrated with a message broker, such as RabbitMQ \cite{rabbitmq} or Kafka \cite{apache_kafka}, to implement task queues, resulting in a more complex architecture than Colonies.
@@
 This is particular useful when launching large container images, or workloads (e.g.
-Julia scripts) that takes relative long time to start.
+Julia scripts) that can take long time to start.
 
-OpenWolf \cite{openwolf} is a serverless workflow engine designed to utilize the Function-as-a-Service (FaaS) paradigm for composing complex scientific workflows.
+Today, utilization of serverless computing is experiencing a significant growth \cite{cognit}, primarily due to its potential to liberate developers from the burden of managing underlying cloud infrastructures.
+Attempt are currently being made to implement serverless workflow management systems.
+For example, OpenWolf \cite{openwolf} is a serverless workflow engine designed to utilize the Function-as-a-Service (FaaS) paradigm for composing complex scientific workflows.
 It is based on OpenFaaS \cite{openfaas}, which allows functions to run on Kubernetes clusters.
-In contrast, Colonies does not require Kubernetes, but provides a zero-trust security protocol to allow functions to be executed securely by distributed executor workers deployed anywhere on the Internet.
-
 The serverless workflow project \cite{serverlessworkflows} aims at providing a vendor-neutral workflow DSL.
@@
 A similar engine is proposed in \cite{scafe}.
+Colonies can also be used to implement serverless workflow management systems.
+This will be further discussed in Section \ref{faas}.
+However, in contrast to previous work, Colonies is plattform independent and does not require Kubernetes.
+By using a zero-trust security protocol, functions can be securely executed by distributed executor deployed anywhere on the Internet.
 It is important to point out that Colonies does not provide an infrastructure for function execution.
@@
 
+Currently, microservices are primarily used to implement large-scale web applications or Internet applications requiring high-availability.
+It has not yet become a prevalent design principle for workload management or implementation of HPC applications.
+Instead, simple job scripts are commonly used.
 J.
@@
 A core concept of the Colonies framework is the notion of \emph{processes}.
-A process contains meta-information about computations executed by a remote computer programs, referred to as a \emph{distributed executors}.
+A process contains meta-information about computations executed by a remote computer programs, which are referred to as \emph{Distributed executors}.
 Specifically, a process contains a definition of a function to be executed, as well as contextual information such as execution status, including the result of the computation.
-One can think of a process as a digital twin of a computing operation.
+One can think of a process as a digital twin of a real computing operation process.
 It is essential to note that a process does not necessarily have to be an operating system process; it can represent any type of computation, e.g.
-a remote function call executed by any kind of software.
+a remote procedure call executed by any kind of software.
 
@@
 The \emph{Colonies servers} form the backbone of the framework, functioning as a centralized control system for task submission and assignment.
+The Colonies server acts as a job broker for executors, almost like an employment agency for people.
 All execution information and history are stored in a database maintained by the Colonies servers.
-Upon submission, a process is stored in the database.
-When it assigned to an executor its execution status is changed from \emph{waiting} to \emph{running}.
-Consequently, it is possible to submit a process that should run in the future, even if no executors can run the process at the moment.
+Upon submission, a process is stored in the database, serving as a queueing system.
+When a process is assigned to an executor its execution status is changed from \emph{waiting} to \emph{running}.
+Consequently, it is possible to submit a process that should run in the future even if no executors can run the process at the moment.
+In this way, Colonies supports both batch and real-time processing.
 
 \emph{Distributed executors} are responsible for executing processes assigned by the Colonies Servers.
-Several distributed executors form a \emph{Colony}, which is a collection of executors operating as a cohesive unit, where each executor is responsible for executing a specific type of process.
-One could view a Colony as an organization of distributed computer software acting a single virtual computer system.
+Several distributed executors form a \emph{Colony}, which is a collection of executors operating as a cohesive unit where each executor is responsible for executing specific type of processes.
+One could view a Colony as an organization or a society of distributed computer softwares acting a single virtual computer system.
 To interact with other executors, executors must prove their Colony membership using a cryptographic protocol that follows the zero-trust security principle \emph{never trust, always verify}.
-This security protocol ensures that users can keep control even when executors are distributed across plattforms, e.g.
-some  XXX
-
-The remainder of this section outlines the underlying design principles and offers a deeper technical description of the framework. The next section describes how
+This security protocol ensures that users can keep control even when executors are distributed across plattforms.
+Zero-trust security is fundamental to the Colonies framework and will be further discussed in Section \ref{zerotrustsecurity}.
+The following sections outlines the underlying design principles and describes the Colonies framework in more details.
 
@@
 By dividing the application into smaller, focused microservices, applications become easier to understand, maintain, and develop.
-Each microservice can be scaled independently, making it easier to handle increased demand for a certain service.
-Additionally, different microservices can be developed by using diverse technologies, frameworks, or programming languages, enabling developers to select the most suitable tools for each specific problem.
-By assigning ownership of specific microservices to individual teams, it also becomes easier to coordinate work and maintain a consistent development process.
+In the Colonies framework, executors are microservices having the following characteristics:
 
-Currently, microservices are primarily used to implement large-scale web applications or Internet applications requiring high-availability.
-It has not yet become a prevalent design principle for workload management or implementation of HPC applications.
-Instead, simple job scripts are commonly used.
-In some cases, message brokers (e.g., RabbitMQ) are used to build worker queues to distribute tasks among multiple workers.
-Although this approach may be effective for simpler applications, creating dependencies between tasks, such as controlling the order of execution or passing information between tasks is not straightforward.
+\begin{itemize}
+\item \textbf{Single responsibility:} Each executor is only accountable for executing specific functions. This makes the system easier to understand, develop, test and maintain.  
+\item \textbf{Loosely coupled:} Executors are designed with minimal dependencies on other executors, enabling various software development teams to work independently.
+For instance, a data engineering team may handle the implementation of pre-processing functions, while a data science team oversees machine learning functions, and another team manages visualization or customer integration etc.
+\item \textbf{Scalability:} Executors can be deployed independently, which enables horizontal scaling. This allows for better resource utilization, parallelism, and improved performance. 
+\item \textbf{Resilience:} The failure of a single executor does not compromise the entire application or workflow.
+Executors' isolation from one another contributes to a more resilient and fault-tolerant system.
+If an executor crashes during execution, the process is automatically reassigned to another executor.
+\item \textbf{DevOps and Continuous Integration:} Executors' inherent fault-tolerant design permits changes to individual executors without impacting the entire system.
+This makes it possible to seamlessly update the system.
+\item \textbf{Technology agnostic:} Executors can be implemented in any programming language, facilitating seamless integration with other applications and systems.
+\item \textbf{Decentralized governance:} As the Colonies framework is technology-agnostic, different software development teams can make independent technology and design decisions when developing executors, promoting greater flexibility and adaptability.
+For example, some executors may be implemented in Rust, while others may use Python to leverage state-of-the-art machine learning frameworks.
+\end{itemize}
 
-Colonies is based on a microservice architecture where workflows are decomposed into a set of independent functions.
-An example of a function could for example be a function that trains a neural network, prepare a batch of data, or upload inferred results to a third-party database etc.
-An executor is responsible for executing one or several functions, making it very simular to a microservice.
-All coordination is managed by a cluster of Colonies servers, allowing complex workflows to be broken down into independent functions that can be developed and tested separately.
-Analogous to traditional microservices, scalability can be achieved simply by adding more executors implementing the same function specification.
-If an executor crashes during task execution, the task is automatically reassigned to another executor.
+Although microservices has attractive properties and simplifes both development and deployment of executors, they also introduce complexity to the Colonies servers.
+To implement a workflow management framework supporting distributed microservices, the following challenges must be addressed:
+
+\begin{itemize}
+\item \textbf{Process management:} Colonies must be able to distribute processes among available executors based on their capabilities and current workload.
+This involves assigning processes to the most appropriate executor and then monitor process progress.
+
+\item \textbf{Fault tolerance and recovery:} In the event of executor failures, the Colonies framework must be able to re-assign processes to other executors, manage executor restarts, or trigger recovery mechanisms to maintain system reliability and resilience.
+In addition, the Colonies framework must be able to handle restarts or crashes of Colonies servers, thereby preventing any internal states in Colonies from becoming corrupted.
+
+\item \textbf{Load balancing:} The Colonies framework must manage load balancing among executors to optimize performance and avoid overloading any single executor.
+
+\item \textbf{Service discovery:} The Colonies framework must enable executors to dynamically register and deregister to Colonies servers.
+It must be possible to deploy executors anywhere on the Internet, even behind firewalls.
+
+\item \textbf{Workflow orchestration:} The Colonies framework must coordinate and orchestrate complex multi-step workflows executed by several executors, sometimes in parallel where executors run on different plattforms.
+Colonies must define the sequence in which processes should be executed, manages dependencies among processes, and ensure that workflows execute successfully to completion.
+
+\item \textbf{Monitoring and debugging:} The Colonies framework must monitor the overall system, allowing administrators to track the health, performance, and resource utilization of the executors and the system as a whole.
+\end{itemize}
+
+The Colonies framework uses a combination of technologies to address the aforementioned challenge.
+A fundamental design principle is stateless web serices, which makes Colonies simpler, more reliable and scalable, and easier to implement.
+To ensure reliability and data consistency, distributed consensus algorithms are needed to offer high-availability and handle server crashes.
+The remainder of this section describes how the Colonies framework is implemented.
+The next discusses the role of queues to support batch jobs and realtime processing, but also as a mechanism to enable loose coupling of system components.
 
 \subsection{The role of queues as seperation of concerns}
-Separation of concerns (SoC) is a fundamental design principle in computer science that aims to break down a complex software system into smaller, more manageable parts.
+Separation of concerns (SoC) is a design principle to break down a complex software system into smaller, more manageable parts.
 For example, HTTP APIs can be used to abstract away implementation detail and provide a clear and simple interface for interacting with a particular service.
@@
 Queues can be implemented in different ways, and while message brokers are a common solution, Colonies adopts an alternative strategy and leverage a standard database and querying it for tasks to assign to different executors.
-One key advantage of this approach is that it enables fine-grained task assignments, making it possible to assign specific tasks to particular executors.
-For instance, an executor of the browser type can be limited to only executing tasks of the browser type.
+One key advantage of this approach is that it enables fine-grained process assignments, making it possible to assign specific processes to particular executors.
+For instance, an executor of the browser type can be limited to only executing processes in web applications.
 This level of granularity cannot easily be implemented using message brokers which generally do not offer introspection of queue, or provide the ability to pull specific messages out of the queue.
@@
 
-Implementing queue functionality using a database can be achieved in SQL by utilizing the \emph{order by timestamp} and \emph{limit 1} clauses to select tasks according to their submission times.
-Upon submission of a task, a \emph{priority time} value is computed and stored in the database.
-The primary objective of the priority time is to modify the submission timestamp by subtracting a delta time calculated as a function a priority value.
+Implementing a queue using a database can be achieved in SQL by utilizing the \emph{order by timestamp} and \emph{limit 1} clauses to select tasks according to their submission times.
+To make it possible to handle priorites, adjust the submission time so that higher priority processes are processed before lower priority processes.
+When a process is submitted, a \emph{priority time} value is computed and stored in the database.
+The primary is a modified submission timestamp calculated by subtracting a delta time calculated as a function of a priority value.
 This enables prioritization of more critical tasks.
@@
 \subsection{Process tables}
-Colonies provides a platform for executors to interact with each other by enabling them to submit function specifications to a Colonies server.
-Once submitted, other executors can connect to the server to receive execution assignments.
+Colonies enables executors to interact with each other by submitting function specifications to the Colonies servers.
+Once submitted, other executors can connect to the server to receive process execution assignments.
 When a function specification is submitted to the Colonies server, it is store as a process entry in a process table database.
-In this case, a process is similar to a task or a job, but also contains contextual information about the execution, for example, execution status (waiting, running, successful, failed), assigned executor ID, submission time, execution time, deadlines, priority etc.
-It also contains input and argument values as well as return values from the function innovation.
-In the remainder of the paper, we are going to use the term process instead of task or job.
 
-The Colonies server acts as a job broker for executors, almost like an employment agency for people.
 When an executor connects to the Colonies server, the server hang the incomming HTTP connection\footnote{An alternative protocol is to use WebSockets or gPRC to communicate with the Colonies server.} until the executor is assigned a process, or until a connection timer expires.
@@
 The Colonies server makes sure that the conditions (i.e requirmenets) of a function specification matches the capability of an executor.
-Note that the executors are responsibility of interpreting the assigned process and function specification to execute relevant code.
 
@@
 \subsection{A stateless failsafe mechanism}
-The primary objective of all Colonies API requests is to alter a state stored in the database or retrieve information from the database.
+The primary objective of all Colonies API requests is to alter some state stored in the database or retrieve information from the database.
 The Colonies framework is designed to be stateless, meaning that the Colonies server does not keep any information between requests.
@@
 The \emph{maxexectime} attribute specifies the maximum time an executor may run a process (in this case, 100 seconds).
-Before a process is assigned to an executor, the Colonies server updates the process entry in the process table database and calculates a deadline for when the process must be finished.
+Before a process is assigned to an executor, the Colonies server updates the process entry in the process table database and calculates a deadline when the process must be finished.
 The server then regularly checks for any running processes that have exceeded their deadlines.
@@
 One solution is to select a set of executors to be removed and then starv them out by denying them new process assignments.
-Another, simpler solution, is to immediately destroy the executors and use the \emph{maxexectime} failsafe mechanism move back failed processes to the queue.
-
+Another, simpler solution, is to immediately destroy the executors and use the \emph{maxexectime} failsafe mechanism move back processes from defunct executors to the queue.
 The \emph{maxexectime} failsafe mechanism ensures that processes will eventually be executed even in the case of failures.
 This mechanism also relieves the burden of user to check if a process has been executed or not, as they can simply look up the process in the database to get its current status.
-The Colonies framework also supports subscribing on process changes using an event-driven protocol, but this may be impractical if a process runs for a very long time (e.g.
-days or weeks).
 
@@
 
-\subsubsection{Concurreny and synchronization}
+\subsubsection{Data consistency and distributed consensus}
 Synchronization is esstential to prevent data inconsistency and race conditions when accessing shared resources concurrently with multiple threads.
@@
 
-The assign request allocates binds a process from the database to an executor.
+The assign API request binds a process from the database to an executor.
 Given the multi-threaded nature of the Colonies server, it is essential that the assign request is synchronized to ensure that only one thread at a time can modify the database and update the process table, thus preventing multiple executors from being assigned to the same process.
-As only the assigned executor can modify a running process and only one executor must be assigned to a process, synchronization is required for the assign request.
+To ensure that only one executor can be assigned to a process, the assign request must be synchronization.
 It is worth noting that synchronization is not necessary for other requests.
@@
 The close request set the output of the function innovation and updates the process state to either successful or failed in the process table.
-Since there is only one assigned executor to a process there is no race conditions and consequently no need for synchronization.
+Since there can only be one executor assigned to a process there is no race conditions and consequently no need for synchronization.
 
-To minimize downtime in the event of hardware or software failures, there is a need for high-availability and make use of a cluster of Colonies servers.
-If one Coloniser server crashes, an executor simply need to re-send a failed request, which will then be served by another Colonies server replica.
+To minimize downtime, Colonies supports high-availability deployments.
+If one Coloniser server crashes, an executor simply need to re-send the failed request, which will then be served by another Colonies server replica.
 However, by introducing multiple Colonies servers, there is again a risk of race conditions when assigning processes to executors.
@@
 \subsubsection{Zero-trust security}
+\label{zerotrustsecurity}
 TODO
+
+\section{Use cases}
+\subsection{Serverless computing}
+\label{faas}
 

### 2023-04-08 paper_2023/paper.tex
wip

@@
 To address the aforementioned challenges, the paper introduces an open-source framework called Colonies\footnote{https://github.com/colonyos/colonies}, which is designed to facilitate execution of computational workloads across a diverse range of platforms.
-Colonies enables the development of a \emph{grid-like distributed virtual computer}, called a \emph{Colony}, which can effortlessly integrate with any third-party application or other workflow management systems.
+Colonies enables development of a \emph{grid-like distributed virtual computer}, called a \emph{Colony}, which can effortlessly integrate with any third-party application or other workflow management systems.
 
-Built upon a modular microservice architecture, Colonies enables complex workflows to be broken down into composable functions.
+Built upon a stateless microservice architecture, Colonies enables complex workflows to be broken down into composable functions.
 With the use of an HTTP protocol, these composable functions can be implemented in any programming language and be executed by independent executors deployed across platforms, anywhere on the Internet, including cloud, edge, high-performance computing (HPC) environments, devices, or even web browsers.
@@
 
-\begin{figure}[h]
+\begin{figure}[t]
 	\centering
-    \includegraphics[scale=0.35]{overview.png}
+    \includegraphics[scale=0.45]{overview.png}
 	\caption{Overview of the Colonies framework. Executors may be deployed anywhere on the Internet.}
@@
 
-The Colonies framework uses a combination of technologies to address the aforementioned challenge.
+The Colonies framework uses a combination of technologies to address the aforementioned challenges.
 A fundamental design principle is stateless web serices, which makes Colonies simpler, more reliable and scalable, and easier to implement.
@@
 The remainder of this section describes how the Colonies framework is implemented.
-The next discusses the role of queues to support batch jobs and realtime processing, but also as a mechanism to enable loose coupling of system components.
+The next discusses the role of queues to support batch jobs and realtime processing, but also how queues can be used to enable loose coupling of system components.
+
+\begin{figure}
+     \centering
+     \begin{subfigure}[b]{0.3\textwidth}
+         \centering
+         \includegraphics[scale=0.37]{arch.png}
+         \caption{Process assignment steps.}
+     \end{subfigure}
+     \hfill
+     \begin{subfigure}[b]{0.3\textwidth}
+         \centering
+         \includegraphics[scale=0.37]{process.png}
+         \caption{Manipulation of a process.}
+     \end{subfigure}
+     \caption{Process management via Colonies HTTP API. Note that only the assigned executor has write access to the process database entry.}
+     \label{fig:process_assignment}
+\end{figure}
 
@@
 
-\begin{equation}
-    \label{eq:pt}
-    priority_{time}=submission_{time} - priority \cdot 10^9 \cdot 60 \cdot 60 \cdot 24
-\end{equation}
-
-Implementing a queue using a database can be achieved in SQL by utilizing the \emph{order by timestamp} and \emph{limit 1} clauses to select tasks according to their submission times.
-To make it possible to handle priorites, adjust the submission time so that higher priority processes are processed before lower priority processes.
-When a process is submitted, a \emph{priority time} value is computed and stored in the database.
-The primary is a modified submission timestamp calculated by subtracting a delta time calculated as a function of a priority value.
-This enables prioritization of more critical tasks.
-Equation \ref{eq:pt} shows how the priority time is calculated for a nanosecond timestamp.
-
-\begin{figure}
-     \centering
-     \begin{subfigure}[b]{0.3\textwidth}
-         \centering
-         \includegraphics[scale=0.37]{arch.png}
-         \caption{Process assignment steps.}
-     \end{subfigure}
-     \hfill
-     \begin{subfigure}[b]{0.3\textwidth}
-         \centering
-         \includegraphics[scale=0.37]{process.png}
-         \caption{Manipulation of a process.}
-     \end{subfigure}
-     \caption{Process management via Colonies HTTP API. Note that only the assigned executor has write access to the process database entry.}
-     \label{fig:process_assignment}
-\end{figure}
-
 \subsection{Process tables}
 Colonies enables executors to interact with each other by submitting function specifications to the Colonies servers.
-Once submitted, other executors can connect to the server to receive process execution assignments.
-When a function specification is submitted to the Colonies server, it is store as a process entry in a process table database.
+Once submitted, other executors can connect to the servers to receive process execution assignments.
+A new process entry is added to the process table database when a function specification is submitted to the Colonies server.
+
+\begin{table}[t]
+	\caption{Process Table}
+	\centering
+	\begin{tabular}{llllll}
+		\toprule
+		\cmidrule(r){1-2}
+        Process Id & Function Spec & Wait for Parents & Assigned Executor Id & State      & Priority Time \\
+		\midrule
+        $P_{1}$    & $F_{1}$       & $False$          & $E_{1}$              & Successful & 1679906715352024000 \\
+        $P_{2}$    & $F_{2}$       & $False$          & $E_{1}$              & Running    & 1679906715353453000 \\
+        $P_{3}$    & $F_{3}$       & $False$          & $E_{2}$              & Running    & 1679906715354286000 \\
+        $P_{4}$    & $F_{4}$       & $True$           & -                    & Waiting    & 1679906715355188000 \\
+		\bottomrule
+	\end{tabular}
+	\label{proctable}
+\end{table}
+
+\begin{table}[t]
+	\caption{Function Specifications}
+	\centering
+	\begin{tabular}{llllll}
+		\toprule
+		\cmidrule(r){1-2}
+        Function Spec & Function        & Executor Type & Priority & Max Exec Time & Max Retries \\
+		\midrule
+        $F_{1}$       & gen\_nums()     & Edge          & 1        & 200 s         & 5 \\
+        $F_{2}$       & square()        & Cloud         & 1        & 200 s         & 5 \\
+        $F_{3}$       & square()        & Cloud         & 1        & 200 s         & 5 \\
+        $F_{4}$       & sum()           & Browser       & 1        & 200 s         & 5 \\
+		\bottomrule
+	\end{tabular}
+	\label{functable}
+\end{table}
 
 When an executor connects to the Colonies server, the server hang the incomming HTTP connection\footnote{An alternative protocol is to use WebSockets or gPRC to communicate with the Colonies server.} until the executor is assigned a process, or until a connection timer expires.
-Note that the Colonies server does not connect to any executor.
+Note that the Colonies server does not connect to the executors.
 Rather, it is the responsibility of the executors to connect to the Colonies server.
 This enables executors to be deployed anywhere on the Internet, behind firewalls, commercial telco networks, or even in web browser-based applications.
-
-Figure \ref{fig:process_assignment} illustrates an executor submitting a function specification that is later assigned to another executor.
-When registering, executors have to specify to the Colonies server which functions they are capable of executing.
-The Colonies server makes sure that the conditions (i.e requirmenets) of a function specification matches the capability of an executor.
 
@@
 \end{figure}
+
+Figure \ref{fig:process_assignment} depicts an executor submitting a function specification that is later assigned to another executor.
+When registering, executors have to specify to the Colonies server which functions they are capable of executing.
+The Colonies server then makes sure that the conditions (i.e requirmenets) part of the function specification matches the capability of an executor.
+
+\begin{equation}
+    \label{eq:pt}
+    priority_{time}=submission_{time} - priority \cdot 10^9 \cdot 60 \cdot 60 \cdot 24
+\end{equation}
+
+Table \ref{proctable} shows an example of a process table.
+The process table also contains a reference to a function specification depicted in Table \ref{functable}.
+To assign a process to an executor, the Colonies server searches in the process table to find a process matching a waiting executor.
+By using the \emph{priority time} column to sort the processes, the process table serves as a queue.
+This can be done in SQL by utilizing the \emph{order by} to sort processes according their submission time.
+To make it possible to handle priorites, the submission time is adjusted so that higher priority processes are processed before lower priority processes.
+When a process is submitted, a \emph{priority time} value is calculated and stored in the process table.
+Equation \ref{eq:pt} shows how the priority time is calculated for a nanosecond timestamp.
 
@@
 If executors are deployed on Kubernetes, Kubernetes will then automatically redeploy terminated executors.
-The constant replacement of executor instances ensures that the system is capable of gracefully tolerating failures.
+The constant flux of executor replacements ensures that the system is capable of gracefully tolerating failures.
 
@@
 \subsubsection{Workflows}
-A workflow is a series of tasks that need to be completed in a specific order.
-Workflows are often represented as directed acyclic graphs (DAGs), where nodes represent tasks and edges represent dependencies or data flow between tasks.
-
-\begin{figure}[h]
-	\centering
-    \includegraphics[scale=0.30]{workflow.png}
-	\caption{cron management}
-	\label{fig:fig3}
-\end{figure}
-
-\begin{table}[h]
-	\caption{Function Specifications}
-	\centering
-	\begin{tabular}{llllll}
-		\toprule
-		\cmidrule(r){1-2}
-        Function Spec & Function        & Executor Type & Priority & Max Exec Time & Max Retries \\
-		\midrule
-        $F_{1}$       & gen\_nums()     & Edge          & 1        & 200 s         & 5 \\
-        $F_{2}$       & square()        & Cloud         & 1        & 200 s         & 5 \\
-        $F_{3}$       & square()        & Cloud         & 1        & 200 s         & 5 \\
-        $F_{4}$       & sum()           & Browser       & 1        & 200 s         & 5 \\
-		\bottomrule
-	\end{tabular}
-	\label{tab:table}
-\end{table}
-
-\begin{table}[h]
-	\caption{Snapshot of Process Table as in Step 2}
-	\centering
-	\begin{tabular}{llllll}
-		\toprule
-		\cmidrule(r){1-2}
-        Process Id & Function Spec & Wait for Parents & Assigned Executor Id & State      & Priority Time \\
-		\midrule
-        $P_{1}$    & $F_{1}$       & $False$          & $E_{1}$              & Successful & 1679906715352024000 \\
-        $P_{2}$    & $F_{2}$       & $False$          & $E_{1}$              & Running    & 1679906715353453000 \\
-        $P_{3}$    & $F_{3}$       & $False$          & $E_{2}$              & Running    & 1679906715354286000 \\
-        $P_{4}$    & $F_{4}$       & $True$           & -                    & Waiting    & 1679906715355188000 \\
-		\bottomrule
-	\end{tabular}
-	\label{tab:table}
-\end{table}
+A workflow is a series of processes that need to be completed in a specific order.
+In Colonies, workflows are represented as directed acyclic graphs (DAGs) where nodes represent processes and edges represent dependencies and data flow between processes.
+Similar to processes, workflows are managed completely statelessly.
+When a workflow is submitted, all processes part of the workflow are submitted and added to the process table similar to how ordinary processes are handled.
+To control the order processes can be assigned, Colonies sets a flag, \emph{wait for parents} to prevent processes to be assigned to an executor before their dependencies have been satisfied.
+Note that processes may run in parallel if sufficient number of executors are available.
 
@@
 	\end{tabular}
-	\label{tab:table}
+	\label{deptable}
 \end{table}
 
+\begin{figure}[h]
+	\centering
+    \includegraphics[scale=0.30]{workflow.png}
+	\caption{Workflow execution timeline.}
+	\label{fig:workflowexec}
+\end{figure}
+
+When a process is closed, the Colonies server checks its child processes to see if all their parent processes have also completed.
+If all parents of a child process have finished, the server updates the process table in the database by setting the flag \emph{wait for parents} to false, which allows child processes to be assigned to executors.
+This operation is also done statelessly when processing the close request.
+
+Figure \ref{fig:workflowexec} portrays an execution timeline for a workflow.
+Upon submission, process \(P_{1}\) is assigned to Executor \(E_{1}\).
+After being closed, processes \(P_{2}\) and \(P_{3}\) are simultaneously assigned to \(E_{1}\) and \(E_{2}\), allowing for concurrent execution.
+Lastly, process \(P_{4}\) is assigned to \(E_{2}\).
+
+To be able to generate a DAG, an additional database table is required.
+Table \ref{deptable} shows an example of a dependency table stroing relationship and dependencies between processes.
+The table is updated when a workflow is submitted to a Colonies server.
+As mentioned before, when an executor is assigned a process it obtains exclusive rights to that process within the process table.
+This exclusivity also means that is possible for an exector to dynamically add new children processes the workflow DAG on the fly.
+As a result, it becomes possible to implement MapReduce \cite{mapreduce} data processing patterns similar to Hadoop \cite{hadoop}.
+
 \begin{table}[h]
-	\caption{Input/Output Table}
+	\caption{Input/Output}
 	\centering
@@
 	\end{tabular}
-	\label{tab:table}
+	\label{inouttable}
 \end{table}
+
+To manage data flow between processes that are part of a workflow, it is necessary to store the input and output values resulting from function invocations in a database.
+An example of such a database is presented in Table \ref{inouttable}.
+An executor can subsequently query the output from a parent process and use that as arguments while invoking a function.
 

### 2023-04-08 paper_2023/paper.tex
wip

@@
 The Colonies framework uses a combination of technologies to address the aforementioned challenges.
-A fundamental design principle is stateless web serices, which makes Colonies simpler, more reliable and scalable, and easier to implement.
+A fundamental design principle is statelessness, which makes Colonies simpler, more reliable and scalable, and easier to implement.
 To ensure reliability and data consistency, distributed consensus algorithms are needed to offer high-availability and handle server crashes.
@@
 The primary objective of all Colonies API requests is to alter some state stored in the database or retrieve information from the database.
-The Colonies framework is designed to be stateless, meaning that the Colonies server does not keep any information between requests.
-In other words, each request is handled independently, without relying on any information from previous requests.
+The Colonies framework is designed to be stateless, meaning that the Colonies server does not keep any information in memory between requests.
+Each request is handled independently without relying on information from previous requests.
 
@@
     \includegraphics[scale=0.5]{raft.png}
-	\caption{High-availability deployment of Colonies servcer.}
+	\caption{High-availability deployment of Colonies.}
 	\label{fig:ha_deployment}
@@
 \subsubsection{Cron}
-TODO
-\begin{figure}[h]
-	\centering
-    \includegraphics[scale=0.4]{cron.png}
-	\caption{Sample figure caption.}
-	\label{fig:fig1}
-\end{figure}
+The Colonies framework is based on a stateless architecture where each request is treated as an independent transaction and is processed without any information from prior requests.
+Since there is no session data stored in memory, Colonies can easily switch between servers without disruptions or data losses.
+In a stateful architecture on the other hand, the server must maintain a record of connected clients' state and session data in memory, which can result in complications if a server fails or crashes.
+
+Cron is a time-based job scheduler designed to trigger workflows at pre-defined intervals.
+It can for example be used to automate tasks such as routinely fetching data from external systems.
+To make Colonies robust and easy to implement it is essential to make Cron workflows stateless.
+This can be achieved by introducing an additional database table to store information about Cron workflow, and let the elected Colonies server leader assumes the responsibility of managing Cron workflows through a two-step process.
+
+Firstly, the leading Colonies server calculates a future deadline, indicating when a specific Cron workflow is scheduled to be executed.
+The leader server then conducts periodic scans of the Cron table.
+When the current time surpasses the calculated deadline for a particular workflow, a new workflow is automatically submitted.
+Note that this protocol maintains complete statefullness as no session information is stored in memory between cron scans.
+In the event that the leader Colonies server experiences a crash, a new leader is simply selected and assumes the responsibility of managing Cron workflows according to the two-step protocol.
 
 \subsubsection{Generators}
-TODO
+Another features of the Colonies framework are so-called generators.
+Generators are workflow templates that are triggered automatically when specific conditions are met.
+Generators can for example be used accumulate input data in bulks to improve performance or parallelism.
+This is achieved by sending a special \emph{pack request} to a Colonies server, targeting a certain generator.
+Generators facilities integration with third-party systems as it allows for stateless implementation where interactions with Colonies can occur using a fire-and-forget style.
+Colonies guarantees that all pack request submitted to a generator is evenutally processed, even if a Coloniers server is restarted or fails.
+
+Without generators, third-party software would have to use an alternative method such as a message broker or a database to enqueue input data and periodically empty the queues to generate workflows themselves.
+This adds complexity to the integration process and may requires additional infrastructure components.
+Generators are a powerful feature of the Colonies framework that can help simplify integration and implement robust data processing pipelines.
+
+Generators can be implemented using a similar method to how Cron workflow are implemented.
+The elected leader Colonies server is responsible for scanning a special generator table database to decide if a workflow should be triggered.
+This approach ensures that there is a single point of control for generating workflows and prevents generation of duplicate workflows.
+
+If the leader server crashes, a new leader Colonies server is automatically elected, and it takes over the responsibility for triggering the generators.
+This ensures that the generation of workflows is not impacted by server failures, and the system can continue to operate smoothly even in case of failures.
+It can be worth to point out that since the pack request only adds information to the database and does not manipulate any states, it can be processed by any Colonies server without synchronization precautious.
+This means that any server in the cluster can handle incoming pack requests, which increases the scalability and fault-tolerance of the system.
 
@@
 \label{zerotrustsecurity}
-TODO
+Zero-trust security is a fundamental to the Colonies framework.
+
+\begin{table}[h]
+	\caption{Colony Id: 4787a5071856a4acf702b2ffcea422e3237a679c681314113d86139461290cf4  }
+	\centering
+	\begin{tabular}{ll}
+		\toprule
+		\cmidrule(r){1-2}
+        Executor Id & Executor Name \\
+		\midrule
+        8a491bcac0be623a54411dd0934fdcdc9c844de5700527a7dbc9da08a6a8310d & $E_{1}$ \\
+        6a65f40343999415f47e739f09b625ab437b069b5f591f9844c234533f505bee & $E_{2}$ \\
+        41e4c1b10f92a53b7a5a86620ed366635901a1a96c48312a0ef566a13217fe03 & $E_{3}$ \\
+		\bottomrule
+	\end{tabular}
+	\label{deptable}
+\end{table}
 

### 2023-04-08 paper_2023/paper.tex
wip

@@
 Generators are workflow templates that are triggered automatically when specific conditions are met.
-Generators can for example be used accumulate input data in bulks to improve performance or parallelism.
-This is achieved by sending a special \emph{pack request} to a Colonies server, targeting a certain generator.
-Generators facilities integration with third-party systems as it allows for stateless implementation where interactions with Colonies can occur using a fire-and-forget style.
-Colonies guarantees that all pack request submitted to a generator is evenutally processed, even if a Coloniers server is restarted or fails.
+Generators can be used accumulate input data in bulks to improve performance or parallelism.
+For example, an external service downloads satellite images and a workflow should automatically be triggered when 10 images are collected.
+This is achieved by sending a special \emph{pack request} containing input data to a Colonies server, targeting a certain generator.
 
+Generators facilities integration with third-party systems as it allows for stateless implementation where interactions with Colonies can be done in a fire-and-forget style.
+Colonies guarantees that all pack requests submitted to a generator is processed correctly, even if a Colonies server is restarted or fails.
 Without generators, third-party software would have to use an alternative method such as a message broker or a database to enqueue input data and periodically empty the queues to generate workflows themselves.
@@
 Generators can be implemented using a similar method to how Cron workflow are implemented.
-The elected leader Colonies server is responsible for scanning a special generator table database to decide if a workflow should be triggered.
+The elected leader Colonies server is responsible for scanning a special generator table to decide if a workflow should be triggered.
 This approach ensures that there is a single point of control for generating workflows and prevents generation of duplicate workflows.
-
 If the leader server crashes, a new leader Colonies server is automatically elected, and it takes over the responsibility for triggering the generators.
 This ensures that the generation of workflows is not impacted by server failures, and the system can continue to operate smoothly even in case of failures.
-It can be worth to point out that since the pack request only adds information to the database and does not manipulate any states, it can be processed by any Colonies server without synchronization precautious.
-This means that any server in the cluster can handle incoming pack requests, which increases the scalability and fault-tolerance of the system.
+It can be worth to point out that since the pack request only adds information to the generator table and does not manipulate any states, it can be processed by any Colonies server without synchronization precautious.
+This means that any Colonies server in the cluster can handle incoming pack requests, which increases the scalability and fault-tolerance of the system as a whole.
 

### 2023-04-08 paper_2023/paper.tex
wip

@@
 % Uncomment to override  the `A preprint' in the header
-\renewcommand{\headeright}{Technical Report}
-\renewcommand{\undertitle}{Technical Report}
+\renewcommand{\headeright}{DRAFT Technical Report}
+\renewcommand{\undertitle}{DRAFT Technical Report}
 %\renewcommand{\shorttitle}{\textit{arXiv} Template}
-
-%%% Add PDF metadata to help others organize their library
-%%% Once the PDF is generated, you can check the metadata with
-%%% $ pdfinfo template.pdf
-\hypersetup{
-pdftitle={A template for the arxiv style},
-pdfsubject={q-bio.NC, q-bio.QM},
-pdfauthor={David S.~Hippocampus, Elias D.~Striatum},
-pdfkeywords={First keyword, Second keyword, More},
-}
 
@@
 Built upon a stateless microservice architecture, Colonies enables complex workflows to be broken down into composable functions.
-With the use of an HTTP protocol, these composable functions can be implemented in any programming language and be executed by independent executors deployed across platforms, anywhere on the Internet, including cloud, edge, high-performance computing (HPC) environments, devices, or even web browsers.
+With the use of an HTTP protocol, these composable functions can be implemented in any programming language and be executed by independent executors deployed across platforms, anywhere on the Internet, including cloud, edge, high-performance computing (HPC) environments.
 A zero-trust security protocol enables a collection of distributed executors to operate as a unified entity, thus establishing seamless computational continuums across multiple platforms.
@@
 In addition to a technical description of the Colonies framework, the paper also describes some potential use cases.
-The paper describe how Colonies can be leveraged to build a remote sensing platform on Kubernetes, serve as a building block for edge computing, implement a serverless FaaS (Function-as-a-Service), and how it can be integrated with the Slurm workload manager.
-Finally, the paper presents a performance investigation, as well as scalability and robustness evaluation.
+The paper describe how Colonies can be leveraged to build a remote sensing platform on Kubernetes, serve as a building block for edge computing, implement a serverless FaaS (Function-as-a-Service) platform, and be integrated with the Slurm workload manager.
+Finally, the paper presents a performance and robustness investigation.
 
@@
 % keywords can be removed
-\keywords{Distributed computing, serverless computing \and parallel computing \and Workflow orchestration}
+\keywords{Distributed computing, Serverless computing \and Parallel computing \and Workflow orchestration}
 
@@
 \subsection{Microservices}
-Microservices is an architectural design pattern in which an application is structured as a collection of small, independently deployable, and loosely coupled services that communicate with other microservices through a well-defined API.
+Microservices \cite{microservices} is an architectural design pattern in which an application is structured as a collection of small, independently deployable, and loosely coupled services that communicate with other microservices through a well-defined API.
 By dividing the application into smaller, focused microservices, applications become easier to understand, maintain, and develop.
@@
 Another features of the Colonies framework are so-called generators.
-Generators are workflow templates that are triggered automatically when specific conditions are met.
-Generators can be used accumulate input data in bulks to improve performance or parallelism.
-For example, an external service downloads satellite images and a workflow should automatically be triggered when 10 images are collected.
+Generators are workflow templates that are automatically triggered when specific conditions are met.
+Generators can be used accumulate input data into bulks to improve performance or parallelism.
+For example, an external service downloads satellite images and use a generator to automatically be triggered a workflow when 10 images are collected.
 This is achieved by sending a special \emph{pack request} containing input data to a Colonies server, targeting a certain generator.
@@
 Colonies guarantees that all pack requests submitted to a generator is processed correctly, even if a Colonies server is restarted or fails.
-Without generators, third-party software would have to use an alternative method such as a message broker or a database to enqueue input data and periodically empty the queues to generate workflows themselves.
+Without generators, third-party integration software would have to use an alternative method such as a message broker or a database to enqueue input data and periodically empty the queues to generate workflows themselves.
 This adds complexity to the integration process and may requires additional infrastructure components.
-Generators are a powerful feature of the Colonies framework that can help simplify integration and implement robust data processing pipelines.
+Generators are a powerful feature of the Colonies framework that can help simplify integration and implementing robust data processing pipelines.
 
@@
 This approach ensures that there is a single point of control for generating workflows and prevents generation of duplicate workflows.
-If the leader server crashes, a new leader Colonies server is automatically elected, and it takes over the responsibility for triggering the generators.
+If the leader server crashes, a new leader Colonies server is automatically elected, which takes over the responsibility for triggering generators.
 This ensures that the generation of workflows is not impacted by server failures, and the system can continue to operate smoothly even in case of failures.
@@
 \label{zerotrustsecurity}
-Zero-trust security is a fundamental to the Colonies framework.
+As executors may be deployed across multiple platforms, there is a need for a unified security model that can be applied on top of any platform.
+Zero-trust security \cite{zerotrust} is a security model that assumes that any device or user is a potential threat, even if they are located within a secure network perimeter.
+Therefore, zero-trust security requires that every interaction between clients and servers is verified and authenticated before being processed, making it ideal for heterogeneous platform deployments.
+However, there must be a way to verify the identities of the clients to determine whether they should be granted access.
+
+As previously stated, a colony is a collection of distributed executors.
+Executors part of the same colony trust each other and can thus submit function specs and get process execution assignments.
+To implement such a scheme, the Colonies server must to be able to verify the identity of the executors and then check their colony membership.
+This can be accomplished in various ways.
+One solution is to use public key encryption and assign each executor a pair of keys - a public key and a private key.
+The public key is openly available and uploaded to the Colonies server, whereas the private key is kept secret and is only used by the executor to sign API request messages.
+The Colony server can then verify the signature of incoming request messages and use the public keys to determine whether an executor is a member of the colony it is attempting to interact with.
+
+ECDSA (Elliptic Curve Digital Signature Algorithm) \cite{ecdsa} is a digital public key encryption signature algorithm.
+It is a widely used in blockchains such as Bitcoin and Ethereum to verify transactions.
+One of the advantages of ECDSA is its ability to recover public keys from received messages and signatures without explicitly transmitting the public keys.
+This is particularly useful as the identity of an executor can be calculated simply as the SHA-3 hash of the recovered signature, saving space and eliminating the need for storing public keys.
+
+In Colonies, there exist three distinct roles, each with its own set of responsibilities and authority levels.
+The Colonies server owner holds the highest level of authority and has permission to add new colonies to the Colonies server.
+The colony owner is responsible for the management of a given colony and has permission to register or unregister executors within that colony.
+Executors, on the other hand, have the lowest level of authority and are only authorized to manage processes within their appointed colony.
+This functionality can be implemented by introducing a new table to the Colonies database, the colony table.
 
 \begin{table}[h]
-	\caption{Colony Id: 4787a5071856a4acf702b2ffcea422e3237a679c681314113d86139461290cf4  }
+    \caption{Colony Id: \(4787a5071856a4acf702b2ffcea422e3237a679c681314113d86139461290cf4\)}
 	\centering
@@
 		\midrule
-        8a491bcac0be623a54411dd0934fdcdc9c844de5700527a7dbc9da08a6a8310d & $E_{1}$ \\
-        6a65f40343999415f47e739f09b625ab437b069b5f591f9844c234533f505bee & $E_{2}$ \\
-        41e4c1b10f92a53b7a5a86620ed366635901a1a96c48312a0ef566a13217fe03 & $E_{3}$ \\
+        \(8a491bcac0be623a54411dd0934fdcdc9c844de5700527a7dbc9da08a6a8310d\) & $E_{1}$ \\
+        \(6a65f40343999415f47e739f09b625ab437b069b5f591f9844c234533f505bee\) & $E_{2}$ \\
+        \(41e4c1b10f92a53b7a5a86620ed366635901a1a96c48312a0ef566a13217fe03\) & $E_{3}$ \\
 		\bottomrule
 	\end{tabular}
-	\label{deptable}
+	\label{coltable}
 \end{table}
 
+Table \ref{coltable} shows an example of a colony table consisting of three executors.
+In this case, to register a new executor, the SHA-3 hash of the recovered signature needs to match the idendity \(4787a5071856a4acf702b2ffcea422e3237a679c681314113d86139461290cf4\).
+This operation can only be performed by the colony owner possessing the corresponding private key.
+When an executor connects to the Colonies server to either submit a function specification or get a process assignment, it needs to sign the API message with its private key.
+The Colonies server will then recovery the identity of the executor and look up the colony table to check if the executor is a rightful colony member.
+
+\section{Implementation}
+
 \section{Use cases}
+\subsection{Remote sensing}
+\subsection{Edge computing}
 \subsection{Serverless computing}
@@
 \section{Evaluation}
-\subsection{Implementation}
+\section{Discussion}
 

### 2023-04-09 paper_2023/paper.tex
wip

@@
 With the use of an HTTP protocol, these composable functions can be implemented in any programming language and be executed by independent executors deployed across platforms, anywhere on the Internet, including cloud, edge, high-performance computing (HPC) environments.
-A zero-trust security protocol enables a collection of distributed executors to operate as a unified entity, thus establishing seamless computational continuums across multiple platforms.
+A zero-trust security protocol enables a collection of distributed executors to operate as a unified entity, thus establishing seamless compute continuums across multiple platforms.
 
@@
 Table \ref{coltable} shows an example of a colony table consisting of three executors.
-In this case, to register a new executor, the SHA-3 hash of the recovered signature needs to match the idendity \(4787a5071856a4acf702b2ffcea422e3237a679c681314113d86139461290cf4\).
+In this case, to register a new executor, the SHA-3 hash of the recovered signature needs to match the idendity \(4787a5071856a4acf702b2ffcea422e3237a679c681314113d86139461290cf4\) stored in the colony table.
 This operation can only be performed by the colony owner possessing the corresponding private key.
-When an executor connects to the Colonies server to either submit a function specification or get a process assignment, it needs to sign the API message with its private key.
+Similarly, when an executor connects to the Colonies server to either submit a function specification or get a process assignment, it needs to sign the API message with its private key.
 The Colonies server will then recovery the identity of the executor and look up the colony table to check if the executor is a rightful colony member.
+
+An advantage of using a stateless architecture for workflow managament as proposed in the paper is the ability to monitor executor performance to detect any unusual or suspicious behaviors.
+Because all transactions are recorded in a database, it is possible to track the activity of individual executors and identify any unusual patterns or behaviors.
+For example, if an executor suddenly starting to consume lot of resources, it could indicate a performance issue, or it could be a sign of malicious activity, such as a denial-of-service attack.
+Automatic montiring can for example be implemented using anonomaly detection algorithms and machine learning to grade the performance each executor.
+Administrators can then be automatically notified if the grade of an executor exceeds a certain threshold, enabling them to take appropriate action.
 
@@
 
+\newpage
+\appendix
+\section{Appendix - Colonies operations}
+
+\begin{table}[h]
+	\centering
+	\begin{tabular}{llcl}
+		\toprule
+		\cmidrule(r){1-4}
+        Operations                 & Role         & Assignment & Note \\
+		\midrule
+        \(add\_colony\)            & Server owner &            & Registers a new colony to the Colonies server. \\
+        \(delete\_colony\)         & Server owner &            & Unregisters a colony from the Colonies server. \\
+        \(add\_executor\)          & Colony owner &            & Register a new executor to a colony. \\
+        \(delete\_executor\)       & Colony owner &            & Unregister a executor from a colony. \\
+        \(approve\_executor\)      & Colony owner &            & Approve an executor to get process assignments. \\
+        \(reject\_executor\)       & Colony owner &            & Disaprove an executor from getting process assignments. \\
+        \(get\_executors\)         & Executor     &            & List all executors member of a colony. \\
+        \(get\_executor\)          & Executor     &            & Get info about an executor. \\
+        \(get\_colonies\)          & Executor     &            & List all colonies on the Colonies server. \\
+        \(get\_colony\)            & Executor     &            & Get info about a colony. \\
+        \(subscribe\_process\)     & Executor     &            & Subscribe to a particular process. \\
+        \(subscribe\_processes\)   & Executor     &            & Subscribe to processes status changes. \\
+        \(submit\)                 & Executor     &            & Submit a function specification. \\
+        \(assign\)                 & Executor     &            & Assign and binds a waiting process to an executor. \\
+        \(unassign\)               & Executor     & x          & Unassign a process from a executor. \\
+        \(close\)                  & Executor     & x          & Close a process as successful. \\
+        \(fail\)                   & Executor     & x          & Close a process as failed. \\
+        \(get\_processes\)         & Executor     &            & List all waiting, successful or failed processes. \\
+        \(get\_process\)           & Executor     &            & Get info about a process. \\
+        \(delete\_all\_processes\) & Colony owner &            & Delete all processes. \\
+        \(delete\_process\)        & Colony owner &            & Delete a process. \\
+        \(add\_attribute\)         & Executor     & x          & Add attribuite to a process. \\
+        \(remove\_attribute\)      & Executor     & x          & Remove attribute from a process. \\
+        \(subscribe\_attribute\)   & Executor     &            & Subscribe to a attribute. \\
+        \(submit\_workflow\)       & Executor     &            & Submit a workflow. \\
+        \(get\_workflow\)          & Executor     &            & Get info about a workflow. \\
+        \(get\_dag\)               & Executor     &            & Get info about a DAG. \\
+        \(add\_child\)             & Executor     & x          & Add a child to a process part of a DAG. \\
+        \(add\_generator\)         & Colony owner &            & Add a generator. \\
+        \(get\_generators\)        & Executor     &            & Get info about a generator. \\
+        \(pack\)                   & Executor     &            & Add input to an a generator. \\
+        \(remove\_generator\)      & Colony owner &            & Remove a generator. \\
+        \(add\_cron\)              & Colony owner &            & Add a cron to a colony.\\
+        \(remove\_cron\)           & Colony owner &            & Remove a cron from a colony.\\
+        \(get\_cron\)              & Executor     &            & Get info about a cron. \\
+        \(delete\_cron\)           & Colony owner &            & Delete a cron from a colony.\\
+        \(get\_crons\)             & Executor     &            & List all crons part of a colony.\\
+        \(add\_function\)          & Executor     &            & Register a function to an executor. \\
+        \(remove\_function\)       & Executor     &            & Unregister a function from an executor. \\
+        \(get\_functions\)         & Executor     &            & List all functions supported by the colony. \\
+		\bottomrule
+	\end{tabular}
+	\label{optable}
+\end{table}
+
 \end{document}

### 2023-04-09 paper_2023/paper.tex
wip

@@
 % Uncomment to override  the `A preprint' in the header
-\renewcommand{\headeright}{DRAFT Technical Report}
-\renewcommand{\undertitle}{DRAFT Technical Report}
+\renewcommand{\headeright}{Technical Report}
+\renewcommand{\undertitle}{Technical Report}
 %\renewcommand{\shorttitle}{\textit{arXiv} Template}
@@
 
-Built upon a stateless microservice architecture, Colonies enables complex workflows to be broken down into composable functions.
+Based on a stateless microservice architecture, Colonies enables complex workflows to be broken down into composable functions.
 With the use of an HTTP protocol, these composable functions can be implemented in any programming language and be executed by independent executors deployed across platforms, anywhere on the Internet, including cloud, edge, high-performance computing (HPC) environments.
@@
 In addition to a technical description of the Colonies framework, the paper also describes some potential use cases.
-The paper describe how Colonies can be leveraged to build a remote sensing platform on Kubernetes, serve as a building block for edge computing, implement a serverless FaaS (Function-as-a-Service) platform, and be integrated with the Slurm workload manager.
+The paper describes how Colonies can be leveraged to build a remote sensing platform on Kubernetes, serve as a building block for edge computing, implement a serverless FaaS (Function-as-a-Service) platform, and be integrated with the Slurm workload manager.
 Finally, the paper presents a performance and robustness investigation.
@@
 % keywords can be removed
-\keywords{Distributed computing, Serverless computing \and Parallel computing \and Workflow orchestration}
+\keywords{Distributed computing \and HPC \and Serverless computing \and Parallel computing \and Workflow orchestration}
 
@@
 In contrast to traditional IT workloads, both the data and the model must be managed in addition to the software itself.
-As most models require regular re-training or re-calibration, it must be possible to update deployed models and software seamlessly without losing information or breaking the system.
+As most models require regular retraining or re-calibration, it must be possible to update deployed models and software seamlessly without losing information or breaking the system.
 In many cases, there is a constant flow of data ingested into the system which must be managed even in case of failures.
-This becomes even more challenging when nodes or parts of the underlying infrastructure become unavailable due to maintenance such as software updates, hardware replacements and sometimes missconfiguration problems.
+This becomes even more challenging when nodes or parts of the underlying infrastructure become unavailable due to maintenance such as software updates, hardware replacements and sometimes misconfiguration problems.
 
@@
 The primary objective is to establish a platform where monolithic workflows can be decomposed into independent microservices that can easily be integrated with other systems or applications running anywhere on the Internet.
-The remainder of the paper describes the Colonies framework and how it can be used to create robust and scalable AI systems across plattforms.
+The remainder of the paper describes the Colonies framework and how it can be used to create robust and scalable AI systems across platforms.
 
@@
 Argo \cite{argowf} is an open source container-native workflow engine for orchestrating parallel jobs on Kubernetes.
-It is can be used for running CI/CD pipelines or compute intensive machine learning or data processing tasks where each job runs as a container.
+It can be used for running CI/CD pipelines or compute intensive machine learning or data processing tasks where each job runs as a container.
 In contrast, Colonies offers a more versatile approach, allowing jobs to be launched within an already started container.
-As launching new containers on Kubernetes can occasionally be time-consuming, Colonies can delivers higher throughput as the costs of starting new jobs are minimal.
-This is particular useful when launching large container images, or workloads (e.g.
-Julia scripts) that can take long time to start.
+As launching new containers on Kubernetes can occasionally be time-consuming, Colonies can deliver higher throughput as the costs of starting new jobs are minimal.
+This is particularly useful when launching large container images, or workloads (e.g.
+Julia scripts) that can take a long time to start.
 
 Today, utilization of serverless computing is experiencing a significant growth \cite{cognit}, primarily due to its potential to liberate developers from the burden of managing underlying cloud infrastructures.
-Attempt are currently being made to implement serverless workflow management systems.
+Attempts are currently being made to implement serverless workflow management systems.
 For example, OpenWolf \cite{openwolf} is a serverless workflow engine designed to utilize the Function-as-a-Service (FaaS) paradigm for composing complex scientific workflows.
@@
 This will be further discussed in Section \ref{faas}.
-However, in contrast to previous work, Colonies is plattform independent and does not require Kubernetes.
-By using a zero-trust security protocol, functions can be securely executed by distributed executor deployed anywhere on the Internet.
+However, in contrast to previous work, Colonies is platform independent and does not require Kubernetes.
+By using a zero-trust security protocol, functions can be securely executed by distributed executors deployed anywhere on the Internet.
 It is important to point out that Colonies does not provide an infrastructure for function execution.
-Instead, Colonies primary role is to serve as a platform for coordinating function executions which are carried out by distributed executor.
+Instead, Colonies primary role is to serve as a platform for coordinating function executions which are carried out by distributed executors.
 
@@
 \cite{GarciaRepresa1740746, GarciaRepresa1640771} explore various challenges associated with developing microservice-based workflow management for industrial automation within the context of the Arrowhead project \cite{delsing2017iot}.
-The authors conclude that microservice-based workflow technologies is viable for industrial applications, particularly due to their inherent flexibility.
+The authors conclude that microservice-based workflow technologies are viable for industrial applications, particularly due to their inherent flexibility.
 The primary contribution of this paper is a comprehensive technical description of how to implement a distributed workflow engine based on microservice principles, extending beyond orchestrating microservices to execute automation tasks within one platform.
-The vision is to create a distributed architecuture where microservices can reside anywhere on the Internet and still function as a cohesive unit to execute cross-platform workflows.
+The vision is to create a distributed architecture where microservices can reside anywhere on the Internet and still function as a cohesive unit to execute cross-platform workflows.
 
@@
 A core concept of the Colonies framework is the notion of \emph{processes}.
-A process contains meta-information about computations executed by a remote computer programs, which are referred to as \emph{Distributed executors}.
+A process contains meta-information about computations executed by remote computer programs, referred to as \emph{distributed executors}.
 Specifically, a process contains a definition of a function to be executed, as well as contextual information such as execution status, including the result of the computation.
-One can think of a process as a digital twin of a real computing operation process.
-It is essential to note that a process does not necessarily have to be an operating system process; it can represent any type of computation, e.g.
+One can think of a process as a digital twin of a real computing process.
+It is essential to note that a process does not necessarily have to be an operating system process, but can represent any type of computation, e.g.
 a remote procedure call executed by any kind of software.
@@
 \emph{Distributed executors} are responsible for executing processes assigned by the Colonies Servers.
-Several distributed executors form a \emph{Colony}, which is a collection of executors operating as a cohesive unit where each executor is responsible for executing specific type of processes.
-One could view a Colony as an organization or a society of distributed computer softwares acting a single virtual computer system.
-To interact with other executors, executors must prove their Colony membership using a cryptographic protocol that follows the zero-trust security principle \emph{never trust, always verify}.
-This security protocol ensures that users can keep control even when executors are distributed across plattforms.
+Several distributed executors form a \emph{Colony}, which is a collection of executors operating as a cohesive unit where each executor is responsible for executing specific types of processes.
+One could view a Colony as an organization or a society of distributed computer softwares acting as a single virtual computer system.
+To interact with other executors, executors must prove their Colony membership using a cryptographic protocol that follows the zero-trust security principle: \emph{never trust, always verify}.
+This security protocol ensures that users can keep control even when executors are distributed across platforms.
 Zero-trust security is fundamental to the Colonies framework and will be further discussed in Section \ref{zerotrustsecurity}.
-The following sections outlines the underlying design principles and describes the Colonies framework in more details.
+The following section outlines the underlying design principles and describe the Colonies framework in more detail.
 
@@
 
-Although microservices has attractive properties and simplifes both development and deployment of executors, they also introduce complexity to the Colonies servers.
+Although microservices have attractive properties and simplify both development and deployment of executors, they also introduce complexity to the Colonies servers.
 To implement a workflow management framework supporting distributed microservices, the following challenges must be addressed:
@@
 \item \textbf{Process management:} Colonies must be able to distribute processes among available executors based on their capabilities and current workload.
-This involves assigning processes to the most appropriate executor and then monitor process progress.
+This involves assigning processes to the most appropriate executor and then monitoring process progress.
 
@@
 
-\item \textbf{Workflow orchestration:} The Colonies framework must coordinate and orchestrate complex multi-step workflows executed by several executors, sometimes in parallel where executors run on different plattforms.
-Colonies must define the sequence in which processes should be executed, manages dependencies among processes, and ensure that workflows execute successfully to completion.
+\item \textbf{Workflow orchestration:} The Colonies framework must coordinate and orchestrate complex multi-step workflows executed by several executors, sometimes in parallel where executors run on different platforms.
+Colonies must define the sequence in which processes should be executed, manage dependencies among processes, and ensure that workflows execute successfully to completion.
 
@@
 For example, HTTP APIs can be used to abstract away implementation detail and provide a clear and simple interface for interacting with a particular service.
-However, HTTP protocols alone are insufficient for handling dynamic environments where components frequently fails or the underlying infrastructures is constantly changing.
+However, HTTP protocols alone are insufficient for handling dynamic environments where components frequently fail or the underlying infrastructure is constantly changing.
 To address such environments, an alternative mechanism is necessary.
@@
 Queues enable software services to communicate indirectly by acting as a buffer between them.
-Queues makes it possible to decouple each executor and make them operate independently, e.g.
+Queues make it possible to decouple each executor and make them operate independently, e.g.
 an executor can be updated or replaced without affecting other executors.
-Queues also allow for asynchronous communication between executors, enabling them to process tasks at their own pace.
+Queues also enable asynchronous communication between executors, enabling them to process tasks at their own pace.
 This ensures that slower executors do not bottleneck faster ones, leading to a more efficient and scalable system.
-Most importantely, queues enable load balancing by distributing tasks among multiple executors, thus making it possible to parallelize workflow execution.
+Most importantly, queues enable load balancing by distributing tasks among multiple executors, thus making it possible to parallelize workflow execution.
 
-Queues can be implemented in different ways, and while message brokers are a common solution, Colonies adopts an alternative strategy and leverage a standard database and querying it for tasks to assign to different executors.
+Queues can be implemented in different ways, and while message brokers are a common solution, Colonies adopt an alternative strategy and leverage a standard database and query it for tasks to assign to different executors.
 One key advantage of this approach is that it enables fine-grained process assignments, making it possible to assign specific processes to particular executors.
 For instance, an executor of the browser type can be limited to only executing processes in web applications.
-This level of granularity cannot easily be implemented using message brokers which generally do not offer introspection of queue, or provide the ability to pull specific messages out of the queue.
-Generarally, the only way to retrieve a specific message is to pull all messages from the queue, obtain the message, and then place all remaining messages back into the queue in the same order.
+This level of granularity cannot easily be implemented using message brokers which generally do not offer introspection of queues, or provide the ability to pull specific messages out of the queue.
+Generally, the only way to retrieve a specific message is to pull all messages from the queue, obtain the message, and then place all remaining messages back into the queue in the same order.
 In contrast, a database can function as a queue and a query can match any columns thus making it possible to assign specific executors to specific processes.
@@
 
-When an executor connects to the Colonies server, the server hang the incomming HTTP connection\footnote{An alternative protocol is to use WebSockets or gPRC to communicate with the Colonies server.} until the executor is assigned a process, or until a connection timer expires.
+When an executor connects to the Colonies server, the server hangs the incoming HTTP connection\footnote{An alternative protocol is to use WebSockets or gPRC to communicate with the Colonies server.} until the executor is assigned a process, or until a connection timer expires.
 Note that the Colonies server does not connect to the executors.
@@
     \includegraphics[scale=0.35]{function_spec.png}
-	\caption{Example of a function spec.}
+	\caption{Example of a function specification.}
 	\label{fig:function_spec}
@@
 When registering, executors have to specify to the Colonies server which functions they are capable of executing.
-The Colonies server then makes sure that the conditions (i.e requirmenets) part of the function specification matches the capability of an executor.
+The Colonies server then makes sure that the conditions (i.e requirements) part of the function specification matches the capability of an executor.
 
@@
 By using the \emph{priority time} column to sort the processes, the process table serves as a queue.
-This can be done in SQL by utilizing the \emph{order by} to sort processes according their submission time.
-To make it possible to handle priorites, the submission time is adjusted so that higher priority processes are processed before lower priority processes.
+This can be done in SQL by utilizing the \emph{order by} to sort processes according to their submission time.
+To make it possible to handle priorities, the submission time is adjusted so that higher priority processes are processed before lower priority processes.
 When a process is submitted, a \emph{priority time} value is calculated and stored in the process table.
@@
 The server then regularly checks for any running processes that have exceeded their deadlines.
-If such process is detected, it is reset, allowing it to be re-assigned to other executors.
+If such a process is detected, it is reset, allowing it to be re-assigned to other executors.
 
@@
 Scaling down, however, can be more challenging.
-One solution is to select a set of executors to be removed and then starv them out by denying them new process assignments.
-Another, simpler solution, is to immediately destroy the executors and use the \emph{maxexectime} failsafe mechanism move back processes from defunct executors to the queue.
+One solution is to select a set of executors to be removed and then starve them out by denying them new process assignments.
+Another, simpler solution, is to immediately destroy the executors and use the \emph{maxexectime} failsafe mechanism to move back processes from defunct executors to the queue.
 The \emph{maxexectime} failsafe mechanism ensures that processes will eventually be executed even in the case of failures.
-This mechanism also relieves the burden of user to check if a process has been executed or not, as they can simply look up the process in the database to get its current status.
+This mechanism also relieves the burden of the user to check if a process has been executed or not, as they can simply look up the process in the database to get its current status.
 
@@
 \subsubsection{Data consistency and distributed consensus}
-Synchronization is esstential to prevent data inconsistency and race conditions when accessing shared resources concurrently with multiple threads.
-However, synchronization can also slows down execution as only one thread can access critical sections at a time.
-By carefully designing multithreaded applications and employing the right synchronization techniques, it possible to minimize the performance impact while still ensuring data consistency and correctness.
+Synchronization is essential to prevent data inconsistency and race conditions when accessing shared resources concurrently with multiple threads.
+However, synchronization can also slow down execution as only one thread can access critical sections at a time.
+By carefully designing multithreaded applications and employing the right synchronization techniques, it is possible to minimize the performance impact while still ensuring data consistency and correctness.
 
-The assign API request binds a process from the database to an executor.
-Given the multi-threaded nature of the Colonies server, it is essential that the assign request is synchronized to ensure that only one thread at a time can modify the database and update the process table, thus preventing multiple executors from being assigned to the same process.
-To ensure that only one executor can be assigned to a process, the assign request must be synchronization.
+The \emph{assign} API request binds a process from the database to an executor.
+Given the multi-threaded nature of the Colonies server, it is essential that the \emph{assign} request is synchronized to ensure that only one thread at a time can modify the database and update the process table, thus preventing multiple executors from being assigned to the same process.
+To ensure that only one executor can be assigned to a process, the \emph{assign} request must be synchronized.
 It is worth noting that synchronization is not necessary for other requests.
-For example, as the submit request only add new entries to the process table, there is no race conditions.
-The close request set the output of the function innovation and updates the process state to either successful or failed in the process table.
-Since there can only be one executor assigned to a process there is no race conditions and consequently no need for synchronization.
+For example, as the \emph{submit} request only adds new entries to the process table, there are no race conditions.
+The \emph{close} request sets the output of the function innovation and updates the process state to either successful or failed in the process table.
+Since there can only be one executor assigned to a process there are no race conditions and consequently no need for synchronization.
 
 To minimize downtime, Colonies supports high-availability deployments.
-If one Coloniser server crashes, an executor simply need to re-send the failed request, which will then be served by another Colonies server replica.
+If one Coloniser server crashes, an executor simply need to resend the failed request, which will then be served by another Colonies server replica.
 However, by introducing multiple Colonies servers, there is again a risk of race conditions when assigning processes to executors.
-This means that the Colonies server replicas must coordintate which replica server incomming assign requests so that precisely one executor is assigned to a process.
+This means that the Colonies server replicas must coordinate which replica server incoming assign requests so that precisely one executor is assigned to a process.
 
@@
 
-Incorporting Raft with the Colonies framework allows incoming assign requests to be directed towards the leading Colonies server, thereby ensuring that only one Colonies server replica handles such requests.
+Adding Raft with the Colonies framework allows incoming assign requests to be directed towards the leading Colonies server, thereby ensuring that only one Colonies server replica handles such requests.
 Figure \ref{fig:ha_deployment} shows an overview of a high-availability Colonies deployment.
@@
 In Colonies, workflows are represented as directed acyclic graphs (DAGs) where nodes represent processes and edges represent dependencies and data flow between processes.
-Similar to processes, workflows are managed completely statelessly.
+Similar to processes, workflows are managed completely stateless.
 When a workflow is submitted, all processes part of the workflow are submitted and added to the process table similar to how ordinary processes are handled.
 To control the order processes can be assigned, Colonies sets a flag, \emph{wait for parents} to prevent processes to be assigned to an executor before their dependencies have been satisfied.
-Note that processes may run in parallel if sufficient number of executors are available.
+Note that processes may run in parallel if a sufficient number of executors are available.
 
@@
 To be able to generate a DAG, an additional database table is required.
-Table \ref{deptable} shows an example of a dependency table stroing relationship and dependencies between processes.
+Table \ref{deptable} shows an example of a dependency table storing relationships between processes.
 The table is updated when a workflow is submitted to a Colonies server.
 As mentioned before, when an executor is assigned a process it obtains exclusive rights to that process within the process table.
-This exclusivity also means that is possible for an exector to dynamically add new children processes the workflow DAG on the fly.
+This exclusivity also means that it is possible for an exector to dynamically add new children processesto  the workflow DAG on the fly.
 As a result, it becomes possible to implement MapReduce \cite{mapreduce} data processing patterns similar to Hadoop \cite{hadoop}.
@@
 
-Cron is a time-based job scheduler designed to trigger workflows at pre-defined intervals.
+Cron is a time-based job scheduler designed to trigger workflows at predefined intervals.
 It can for example be used to automate tasks such as routinely fetching data from external systems.
 To make Colonies robust and easy to implement it is essential to make Cron workflows stateless.
-This can be achieved by introducing an additional database table to store information about Cron workflow, and let the elected Colonies server leader assumes the responsibility of managing Cron workflows through a two-step process.
+This can be achieved by introducing an additional database table to store information about Cron workflow, and let the elected Colonies server leader assume the responsibility of managing Cron workflows through a two-step process.
 
@@
 When the current time surpasses the calculated deadline for a particular workflow, a new workflow is automatically submitted.
-Note that this protocol maintains complete statefullness as no session information is stored in memory between cron scans.
+Note that this protocol is completely stateful as no session information is stored in memory between cron scans.
 In the event that the leader Colonies server experiences a crash, a new leader is simply selected and assumes the responsibility of managing Cron workflows according to the two-step protocol.
@@
 \subsubsection{Generators}
-Another features of the Colonies framework are so-called generators.
+Another feature of the Colonies framework are so-called generators.
 Generators are workflow templates that are automatically triggered when specific conditions are met.
-Generators can be used accumulate input data into bulks to improve performance or parallelism.
-For example, an external service downloads satellite images and use a generator to automatically be triggered a workflow when 10 images are collected.
+Generators can be used to accumulate input data into bulks to improve performance or parallelism.
+For example, an external service downloads satellite images and can use a generator to automatically triggered a workflow when 10 images are collected.
 This is achieved by sending a special \emph{pack request} containing input data to a Colonies server, targeting a certain generator.
@@
 Generators facilities integration with third-party systems as it allows for stateless implementation where interactions with Colonies can be done in a fire-and-forget style.
-Colonies guarantees that all pack requests submitted to a generator is processed correctly, even if a Colonies server is restarted or fails.
+Colonies guarantees that all pack requests submitted to a generator are processed correctly, even if a Colonies server is restarted or fails.
 Without generators, third-party integration software would have to use an alternative method such as a message broker or a database to enqueue input data and periodically empty the queues to generate workflows themselves.
-This adds complexity to the integration process and may requires additional infrastructure components.
-Generators are a powerful feature of the Colonies framework that can help simplify integration and implementing robust data processing pipelines.
+This adds complexity to the integration process and may require additional infrastructure components.
+Generators are a powerful feature of the Colonies framework that can help simplify integration and implement robust data processing pipelines.
 
-Generators can be implemented using a similar method to how Cron workflow are implemented.
+Generators can be implemented using a similar method to how Cron workflows are implemented.
 The elected leader Colonies server is responsible for scanning a special generator table to decide if a workflow should be triggered.
@@
 This ensures that the generation of workflows is not impacted by server failures, and the system can continue to operate smoothly even in case of failures.
-It can be worth to point out that since the pack request only adds information to the generator table and does not manipulate any states, it can be processed by any Colonies server without synchronization precautious.
+It can be worth pointing out that since the pack request only adds information to the generator table and does not manipulate any states, it can be processed by any Colonies server without synchronization precautious.
 This means that any Colonies server in the cluster can handle incoming pack requests, which increases the scalability and fault-tolerance of the system as a whole.
@@
 As previously stated, a colony is a collection of distributed executors.
-Executors part of the same colony trust each other and can thus submit function specs and get process execution assignments.
-To implement such a scheme, the Colonies server must to be able to verify the identity of the executors and then check their colony membership.
+Executors part of the same colony trust each other and can thus submit function specifications and get process execution assignments.
+To implement such a scheme, the Colonies server must be able to verify the identity of the executors and then check their colony membership.
 This can be accomplished in various ways.
@@
 ECDSA (Elliptic Curve Digital Signature Algorithm) \cite{ecdsa} is a digital public key encryption signature algorithm.
-It is a widely used in blockchains such as Bitcoin and Ethereum to verify transactions.
+It is widely used in blockchains such as Bitcoin and Ethereum to verify transactions.
 One of the advantages of ECDSA is its ability to recover public keys from received messages and signatures without explicitly transmitting the public keys.
@@
 Table \ref{coltable} shows an example of a colony table consisting of three executors.
-In this case, to register a new executor, the SHA-3 hash of the recovered signature needs to match the idendity \(4787a5071856a4acf702b2ffcea422e3237a679c681314113d86139461290cf4\) stored in the colony table.
+In this case, to register a new executor, the SHA-3 hash of the recovered signature needs to match the identity \(4787a5071856a4acf702b2ffcea422e3237a679c681314113d86139461290cf4\) stored in the colony table.
 This operation can only be performed by the colony owner possessing the corresponding private key.
 Similarly, when an executor connects to the Colonies server to either submit a function specification or get a process assignment, it needs to sign the API message with its private key.
-The Colonies server will then recovery the identity of the executor and look up the colony table to check if the executor is a rightful colony member.
+The Colonies server will then recover the identity of the executor and look up the colony table to check if the executor is a rightful colony member.
+See Appendix A for additional details on how the various roles apply to different Colonies operations.
 
-An advantage of using a stateless architecture for workflow managament as proposed in the paper is the ability to monitor executor performance to detect any unusual or suspicious behaviors.
+An advantage of using a stateless architecture for workflow management as proposed in the paper is the ability to monitor executor performance to detect any unusual or suspicious behaviors.
 Because all transactions are recorded in a database, it is possible to track the activity of individual executors and identify any unusual patterns or behaviors.
-For example, if an executor suddenly starting to consume lot of resources, it could indicate a performance issue, or it could be a sign of malicious activity, such as a denial-of-service attack.
-Automatic montiring can for example be implemented using anonomaly detection algorithms and machine learning to grade the performance each executor.
+For example, if an executor suddenly starts to consume a lot of resources, it could indicate a performance issue, or it could be a sign of malicious activity, such as a denial-of-service attack.
+Automatic monitoring can for example be implemented using anomaly detection algorithms and machine learning to grade the performance of each executor.
 Administrators can then be automatically notified if the grade of an executor exceeds a certain threshold, enabling them to take appropriate action.
@@
 \section{Implementation}
+There are SDKs is Golang\footnote{https://github.com/colonyos/colonies/tree/main/pkg/client}, Rust\footnote{https://github.com/colonyos/rust}, Julia\footnote{https://github.com/colonyos/Colonies.jl}, JavaScript\footnote{https://github.com/colonyos/colonies.js}, Python\footnote{https://github.com/colonyos/pycolonies}, and Haskell\footnote{https://github.com/colonyos/haskell}.
+
+\begin{figure}[h]
+	\centering
+    \includegraphics[scale=0.35]{executor_code.png}
+	\caption{Register an executor to a colony in Python.}
+	\label{fig:dashboard1}
+\end{figure}
+
+\begin{figure}[h]
+	\centering
+    \includegraphics[scale=0.35]{assign_code.png}
+	\caption{Assigning a process to an executor in Python.}
+	\label{fig:dashboard1}
+\end{figure}
+
+\begin{figure}[h]
+	\centering
+    \includegraphics[scale=0.25]{dashboard1.png}
+	\caption{Process information. The timeline makes it possible to follow the execution of process in realtime.}
+	\label{fig:dashboard1}
+\end{figure}
+
+\begin{figure}[h]
+	\centering
+    \includegraphics[scale=0.25]{dashboard2.png}
+	\caption{Visualization of a DAG.}
+	\label{fig:dashboard2}
+\end{figure}
 
 \section{Use cases}
-\subsection{Remote sensing}
-\subsection{Edge computing}
 \subsection{Serverless computing}
 \label{faas}
+\subsection{Edge computing}
+\subsection{Remote sensing}
+\subsection{HPC integration}
 

### 2023-04-10 paper_2023/paper.tex
wip

@@
 \usepackage{subcaption}
+\usepackage{xcolor}
+
+\colorlet{punct}{red!60!black}
+\definecolor{background}{HTML}{EEEEEE}
+\definecolor{delim}{RGB}{20,105,176}
+\colorlet{numb}{magenta!60!black}
+
+\usepackage{bera}% optional: just to have a nice mono-spaced font
+
+\lstdefinelanguage{json}{
+    basicstyle=\normalfont\ttfamily,
+    numbers=left,
+    numberstyle=\scriptsize,
+    stepnumber=1,
+    numbersep=8pt,
+    showstringspaces=false,
+    breaklines=true,
+    frame=lines,
+    backgroundcolor=\color{background},
+    literate=
+      {:}{{{\color{punct}{:}}}}{1}
+      {,}{{{\color{punct}{,}}}}{1}
+      {\{}{{{\color{delim}{\{}}}}{1}
+      {\}}{{{\color{delim}{\}}}}}{1}
+      {[}{{{\color{delim}{[}}}}{1}
+      {]}{{{\color{delim}{]}}}}{1},
+}
 
@@
 % Uncomment to override  the `A preprint' in the header
-\renewcommand{\headeright}{Technical Report}
-\renewcommand{\undertitle}{Technical Report}
+%\renewcommand{\headeright}{Technical Report}
+%\renewcommand{\undertitle}{Technical Report}
 %\renewcommand{\shorttitle}{\textit{arXiv} Template}
@@
 
-To address the aforementioned challenges, the paper introduces an open-source framework called Colonies\footnote{https://github.com/colonyos/colonies}, which is designed to facilitate execution of computational workloads across a diverse range of platforms.
+To address these challenges, the paper presents an open-source framework called Colonies\footnote{https://github.com/colonyos/colonies}, which facilitates execution of computational workloads across a diverse range of platforms.
 Colonies enables development of a \emph{grid-like distributed virtual computer}, called a \emph{Colony}, which can effortlessly integrate with any third-party application or other workflow management systems.
 
-Based on a stateless microservice architecture, Colonies enables complex workflows to be broken down into composable functions.
-With the use of an HTTP protocol, these composable functions can be implemented in any programming language and be executed by independent executors deployed across platforms, anywhere on the Internet, including cloud, edge, high-performance computing (HPC) environments.
-A zero-trust security protocol enables a collection of distributed executors to operate as a unified entity, thus establishing seamless compute continuums across multiple platforms.
+Based on a stateless microservice architecture, complex workflows to be broken down into composable functions.
+These composable functions can be implemented in any programming language and be executed by distributed executors deployed across platforms, anywhere on the Internet such as IoT, cloud, edge, and supercomputers.
+A zero-trust security protocol enables a collection of distributed executors to operate as a unified entity, thus establishing seamless compute continuums across platforms.
 
 In addition to a technical description of the Colonies framework, the paper also describes some potential use cases.
-The paper describes how Colonies can be leveraged to build a remote sensing platform on Kubernetes, serve as a building block for edge computing, implement a serverless FaaS (Function-as-a-Service) platform, and be integrated with the Slurm workload manager.
-Finally, the paper presents a performance and robustness investigation.
+The paper describes how Colonies can be leveraged to build a remote sensing platform on Kubernetes, serve as a building block for edge and serverless computing, and be integrated with the Slurm workload manager.
 
-In summary, Colonies is a highly versatile and scalable framework that can streamline development and deployment of computational workloads across heterogeneous platforms while ensuring full traceability and zero-trust security.
+In summary, Colonies is a highly versatile and scalable framework that can streamline development and deployment of computational workloads across heterogeneous platforms while ensuring scalability, robustness, traceability and zero-trust security.
 \end{abstract}
 
-% keywords can be removed
 \keywords{Distributed computing \and HPC \and Serverless computing \and Parallel computing \and Workflow orchestration}
@@
 		\cmidrule(r){1-2}
-        Process Id & Function Spec & Wait for Parents & Assigned Executor Id & State      & Priority Time \\
+        Process Id & Function Spec & Wait for Parents & Executor  & State      & Priority Time \\
 		\midrule
@@
 
-\begin{figure}[h]
-	\centering
-    \includegraphics[scale=0.35]{function_spec.png}
-	\caption{Example of a function specification.}
-	\label{fig:function_spec}
-\end{figure}
+\begin{lstlisting}[basicstyle=\small, label=fig:function_spec, language=json, basicstyle=\small, caption=Example of a function specification.]
+{
+   "conditions": {
+       "colonyid": "0c1168fe986ffe39fad14f17e0bd9e5896f6d968405ac0fb3380154109ee4022",
+       "executortype": "helloworld_executor"
+   },
+   "funcname": "helloworld",
+   "args": ["hello world"],
+   "maxwaittime": 10,
+   "maxexectime": 100,
+   "maxretries": 3,
+   "priority": 1
+}
+\end{lstlisting}
 
@@
 
+\begin{figure}[h]
+	\centering
+    \includegraphics[scale=0.5]{raft.png}
+	\caption{High-availability deployment of Colonies.}
+	\label{fig:ha_deployment}
+\end{figure}
+
 \subsection{A stateless failsafe mechanism}
+\label{stateless_error_management}
 The primary objective of all Colonies API requests is to alter some state stored in the database or retrieve information from the database.
@@
 The constant flux of executor replacements ensures that the system is capable of gracefully tolerating failures.
-
-\begin{figure}[h]
-	\centering
-    \includegraphics[scale=0.5]{raft.png}
-	\caption{High-availability deployment of Colonies.}
-	\label{fig:ha_deployment}
-\end{figure}
 
@@
 When the current time surpasses the calculated deadline for a particular workflow, a new workflow is automatically submitted.
-Note that this protocol is completely stateful as no session information is stored in memory between cron scans.
+Note that this protocol is completely stateless as no session information is stored in memory between cron scans.
 In the event that the leader Colonies server experiences a crash, a new leader is simply selected and assumes the responsibility of managing Cron workflows according to the two-step protocol.
@@
 Generators can be implemented using a similar method to how Cron workflows are implemented.
-The elected leader Colonies server is responsible for scanning a special generator table to decide if a workflow should be triggered.
+The elected leader Colonies server takes on the responsibility for scanning a special generator table to decide if a workflow should be triggered.
 This approach ensures that there is a single point of control for generating workflows and prevents generation of duplicate workflows.
@@
 \section{Implementation}
-There are SDKs is Golang\footnote{https://github.com/colonyos/colonies/tree/main/pkg/client}, Rust\footnote{https://github.com/colonyos/rust}, Julia\footnote{https://github.com/colonyos/Colonies.jl}, JavaScript\footnote{https://github.com/colonyos/colonies.js}, Python\footnote{https://github.com/colonyos/pycolonies}, and Haskell\footnote{https://github.com/colonyos/haskell}.
+Colonies is implemented in Golang and is publicly available on GitHub\footnote{https://github.com/colonyos/colonies} under the MIT license.
+It consists of a statically compiled binary that offers a CLI tool for deploying Colonies servers or managing the system.
+Furthermore, there are several software development kits (SDKs) available for various programming languages, including Golang\footnote{https://github.com/colonyos/colonies/tree/main/pkg/client}, Rust\footnote{https://github.com/colonyos/rust}, Julia\footnote{https://github.com/colonyos/Colonies.jl}, JavaScript\footnote{https://github.com/colonyos/colonies.js}, Python\footnote{https://github.com/colonyos/pycolonies}, and Haskell\footnote{https://github.com/colonyos/haskell}.
+As an introduction to Colonies, the following section gives a brief overview of the Colonies tools and how to use the Python SDK.
 
-\begin{figure}[h]
-	\centering
-    \includegraphics[scale=0.35]{executor_code.png}
-	\caption{Register an executor to a colony in Python.}
-	\label{fig:dashboard1}
-\end{figure}
+\subsection{Python SDK}
+A colonies application typically consists of a network of executors written in different programming languages and deployed across multiple platforms.
+The executors are small microservices that interact with one another by submitting function specifications or workflows, which are then executed by other executors that are part of the same colony.
 
-\begin{figure}[h]
-	\centering
-    \includegraphics[scale=0.35]{assign_code.png}
-	\caption{Assigning a process to an executor in Python.}
-	\label{fig:dashboard1}
-\end{figure}
+The first step in implementing an executor in Python is to create a private key and identity, followed by registering the executor with a colony.
+Registering an executor can only be carried out by the colony owner with access to the colony private key.
+Code in listening \ref{code:regexecutor} illustrates how to register an executor and which functions a specific executor is capable of running.
+In the example, an executor of type \emph{helloworld\_executor} is registered to run a function named \emph{helloworld}.
 
+\begin{lstlisting}[showstringspaces=false, frame=lines, numbers=left, numberstyle=\scriptsize, backgroundcolor=\color{background}, basicstyle=\small, label=code:regexecutor, language=Python, caption=Register an executor to a colony in Python.]
+colonies = Colonies("localhost", 50080)
+
+crypto = Crypto()
+colonyid = "4787a5071856a4acf702b2ffcea422e3237a679c681314113d86139461290cf4"
+colony_prvkey = "ba949fa134981372d6da62b6a56f336ab4d843b22c02a4257dcf7d0d73097514"
+executor_prvkey = crypto.prvkey()
+executorid = crypto.id(executor_prvkey)
+
+executor = {
+    "executorname": "helloworld_executor",
+    "executorid": executorid,
+    "colonyid": colonyid,
+    "executortype": "helloworld _executor"
+}
+
+colonies.add_executor(executor, colony_prvkey)
+colonies.approve_executor(executorid, colony_prvkey)
+
+# register capability run the helloworld function
+colonies.add_function(executorid, colonyid, "helloworld",  executor_prvkey)
+\end{lstlisting}
+
+After registration, the executor must connect to the Colonies server and request process assignments of the type \emph{helloworld\_executor}.
+Note that executor properties, such as the executor type, are not explicitly specified, but rather derived implicitly by the Colonies server from the executor identity, which is derived from the signature generated by executor's the private key.
+This approach ensures that the executor can only be assigned processes that match its registered capabilities.
+
+\begin{lstlisting}[showstringspaces=false, frame=lines, numbers=left, numberstyle=\scriptsize, backgroundcolor=\color{background}, basicstyle=\small, language=Python, label=code:assign, caption=Assigning and executing a process in Python.]
+while True:
+    process = colonies.assign(colonyid, 10, executor_prvkey)
+    if process["spec"]["funcname"] == "helloworld":
+       colonies.close(process["processid"], ["hello world"], executor_prvkey)
+\end{lstlisting}
+
+This \emph{assign} request is always initiated by the executor using an HTTP-based protocol, allowing executors to be deployed behind firewalls.
+When receiving the request, the Colonies server hangs the request until a matching process is found or a timer expires.
+In the example in Listing \ref{code:assign}, the timer is set for 10 seconds, after which an exception is raised, and the executor must reconnect to the server.
+
+When assigned to a process, the executor interprets the metadata stored in the process data structure, including the function name and arguments, and performs some kind of computation such as preprocessing some data or training a neural network.
+Upon completion, the executor closes the process and sets the output.
+In the example, the output is set to the the string \emph{helloworld}.
+
+\begin{lstlisting}[showstringspaces=false, frame=lines, numbers=left, numberstyle=\scriptsize, backgroundcolor=\color{background}, basicstyle=\small, language=Python, label=code:submit, caption=Submitting a function specification.]
+func_spec = create_func_spec(func="helloworld",
+                             colonyid=colonyid,
+                             executortype="helloworld_executor",
+                             priority=0,
+                             maxexectime=100,
+                             maxretries=3)
+process = colonies.submit(func_spec, executor_prvkey)
+\end{lstlisting}
+
+Listing \ref{code:submit} demonstrates how to submit a function specification using the Python SDK and call the \emph{helloworld} function.
+Note that the executor must complete the process within 100 seconds.
+If the execution takes longer, the process is reassigned to another executor, and the previosuly executor receives an error upon closure.
+As discussed in Section \ref{stateless_error_management}, establishing boundaries on computations is critical to achieving robustness and supporting infrastructural changes that may result from DevOps or IT operations.
+
+The Colonies CLI tool can be used to 
 \begin{figure}[h]
@@
 	\label{fig:dashboard1}
-\end{figure}
-
-\begin{figure}[h]
-	\centering
-    \includegraphics[scale=0.25]{dashboard2.png}
-	\caption{Visualization of a DAG.}
-	\label{fig:dashboard2}
 \end{figure}
@@
 
-\section{Evaluation}
 \section{Discussion}

### 2023-04-10 paper_2023/paper.tex
wip

@@
 The primary contribution of this paper is a comprehensive technical description of how to implement a distributed workflow engine based on microservice principles, extending beyond orchestrating microservices to execute automation tasks within one platform.
-The vision is to create a distributed architecture where microservices can reside anywhere on the Internet and still function as a cohesive unit to execute cross-platform workflows.
+The vision is to create a distributed architecture where microservices can reside anywhere on the Internet and still function as a cohesive unit to execute cross-platform workloads.
 
@@
 After registration, the executor must connect to the Colonies server and request process assignments of the type \emph{helloworld\_executor}.
-Note that executor properties, such as the executor type, are not explicitly specified, but rather derived implicitly by the Colonies server from the executor identity, which is derived from the signature generated by executor's the private key.
+Note that executor properties, such as the executor type, are not explicitly specified, but rather derived implicitly by the Colonies server from the executor identity, which is derived from the signature generated by the executor's private key.
 This approach ensures that the executor can only be assigned processes that match its registered capabilities.
@@
 Upon completion, the executor closes the process and sets the output.
-In the example, the output is set to the the string \emph{helloworld}.
+In the example, the output is set to the string \emph{helloworld}.
 
@@
 Note that the executor must complete the process within 100 seconds.
-If the execution takes longer, the process is reassigned to another executor, and the previosuly executor receives an error upon closure.
+The process is reassigned to another executor if the execution takes too long a time.
+The previous executor then receives an error when trying to send a \emph{close} request.
 As discussed in Section \ref{stateless_error_management}, establishing boundaries on computations is critical to achieving robustness and supporting infrastructural changes that may result from DevOps or IT operations.
 
-The Colonies CLI tool can be used to 
 \begin{figure}[h]
 	\centering
-    \includegraphics[scale=0.25]{dashboard1.png}
+    \includegraphics[scale=0.23]{dashboard1.png}
 	\caption{Process information. The timeline makes it possible to follow the execution of process in realtime.}
@@
 \end{figure}
+
+Figure \ref{fig:dashboard1} depicts a screenshot of the Colonies dashboard tool, which enables users to monitor the status of processes and executors.
+The timeline depicted in the figure allows for real-time monitoring of the execution of processes, similar to tracking a physical package sent through a postal service.
+As previously discussed, it's also possible to subscribe to a process and receive information about its progress, which can be used as another point of integration.
+For example, sending a notification on Slack or by email if a process fails.
+
+\subsection{Workflows}
+A workflow is just a set of processes with dependencies represented as a DAG internally in Colonies.
+Workflows can either be submitted using SDKs or the Colonies CLI, which supports submission of JSON files.
+Listening \ref{code:workflow_spec} shows an example of a workflow consisting of four processes, each requiring different types of executors.
+The management of workflows is entirely handled by Colonies, and each executor executes only processes that are part of the graph for which it is responsible for.
+This approach significantly reduces the complexity of designing, implementing and operating large-scale AI applications and data pipelines.
+
+\begin{lstlisting}[basicstyle=\small, label=code:workflow_spec, language=json, basicstyle=\small, caption=Example of a workflow]
+[ { "nodename": "task_a",
+    "funcname": "echo",
+    "conditions": { "executortype": "executor_type1", "dependencies": [] } },
+  { "nodename": "task_b",
+    "funcname": "echo",
+    "conditions": { "executortype": "executor_type2", "dependencies": ["task_a"] } },
+  { "nodename": "task_c",
+    "funcname": "echo",
+    "conditions": { "executortype": "executor_type3", "dependencies": ["task_a"] } },
+  { "nodename": "task_d",
+    "funcname": "echo",
+    "conditions": { "executortype": "executor_type4", "dependencies": ["task_b", "task_c"] }
+  } ]
+\end{lstlisting}
+
+\begin{figure}[h]
+	\centering
+    \includegraphics[scale=0.23]{dashboard2.png}
+	\caption{Screenshot of a workflow in the Colonies dashboard.}
+	\label{fig:dashboard2}
+\end{figure}
+
+In addition to monitoring processes and executors, the Colonies dashboard can be used to visualize workflows.
+Figure \ref{fig:dashboard2} illustrates the DAG created when submitting the workflow shown in Listing \ref{code:workflow_spec}.
+This functionality makes it possible to follow workflow execution progress.
 

### 2023-04-10 paper_2023/paper.tex
wip

@@
     To address these challenges, the paper presents an open-source framework called Colonies\footnote{https://github.com/colonyos/colonies}, which facilitates execution of computational workloads across a diverse range of platforms.
-Colonies enables development of a \emph{grid-like distributed virtual computer}, called a \emph{Colony}, which can effortlessly integrate with any third-party application or other workflow management systems.
+Colonies enables development of a \emph{grid-like distributed virtual computer}, called a \emph{Colony}, which can effortlessly integrate with any third-party application (or other workflow management systems).
 
@@
 	\centering
-    \includegraphics[scale=0.5]{raft.png}
+    \includegraphics[scale=0.6]{raft.png}
 	\caption{High-availability deployment of Colonies.}
@@
 
-\subsubsection{Workflows}
-A workflow is a series of processes that need to be completed in a specific order.
-In Colonies, workflows are represented as directed acyclic graphs (DAGs) where nodes represent processes and edges represent dependencies and data flow between processes.
-Similar to processes, workflows are managed completely stateless.
-When a workflow is submitted, all processes part of the workflow are submitted and added to the process table similar to how ordinary processes are handled.
-To control the order processes can be assigned, Colonies sets a flag, \emph{wait for parents} to prevent processes to be assigned to an executor before their dependencies have been satisfied.
-Note that processes may run in parallel if a sufficient number of executors are available.
-
 \begin{table}[h]
@@
 \end{table}
+
+\subsubsection{Workflows}
+A workflow is a series of processes that need to be completed in a specific order.
+In Colonies, workflows are represented as directed acyclic graphs (DAGs) where nodes represent processes and edges represent dependencies and data flow between processes.
+Similar to processes, workflows are managed completely stateless.
+When a workflow is submitted, all processes part of the workflow are submitted and added to the process table similar to how ordinary processes are handled.
+To control the order processes can be assigned, Colonies sets a flag, \emph{wait for parents} to prevent processes to be assigned to an executor before their dependencies have been satisfied.
+Note that processes may run in parallel if a sufficient number of executors are available.
 
@@
 
-\begin{lstlisting}[basicstyle=\small, label=code:workflow_spec, language=json, basicstyle=\small, caption=Example of a workflow]
+\begin{lstlisting}[basicstyle=\small, label=code:workflow_spec, language=json, basicstyle=\small, caption=Example of a workflow expressed in JSON.]
 [ { "nodename": "task_a",
@@
 \label{faas}
+
+\begin{figure}[h]
+	\centering
+    \includegraphics[scale=0.44]{cognite_faas.png}
+	\caption{A Colonies-based FaaS serverless computing platform.}
+	\label{fig:ha_deployment}
+\end{figure}
 \subsection{Edge computing}
@@
 \subsection{HPC integration}
+
+\begin{figure}[h]
+	\centering
+    \includegraphics[scale=0.43]{hpc_workflow.png}
+	\caption{HPC workflow.}
+	\label{fig:dashboard2}
+\end{figure}
 

### 2023-04-10 paper_2023/paper.tex
wip

@@
 In addition to a technical description of the Colonies framework, the paper also describes some potential use cases.
-The paper describes how Colonies can be leveraged to build a remote sensing platform on Kubernetes, serve as a building block for edge and serverless computing, and be integrated with the Slurm workload manager.
-
 In summary, Colonies is a highly versatile and scalable framework that can streamline development and deployment of computational workloads across heterogeneous platforms while ensuring scalability, robustness, traceability and zero-trust security.
@@
      \centering
-     \begin{subfigure}[b]{0.3\textwidth}
+     \begin{subfigure}[b]{0.4\textwidth}
          \centering
-         \includegraphics[scale=0.37]{arch.png}
+         \includegraphics[scale=0.45]{arch.png}
          \caption{Process assignment steps.}
@@
      \hfill
-     \begin{subfigure}[b]{0.3\textwidth}
+     \begin{subfigure}[b]{0.4\textwidth}
          \centering
-         \includegraphics[scale=0.37]{process.png}
+         \includegraphics[scale=0.45]{process.png}
          \caption{Manipulation of a process.}
@@
 
-\begin{figure}[h]
+\begin{figure}
      \centering
-    \includegraphics[scale=0.23]{dashboard1.png}
-	\caption{Process information. The timeline makes it possible to follow the execution of process in realtime.}
+     \begin{subfigure}[b]{0.49\textwidth}
+         \centering
+         \includegraphics[scale=0.17]{dashboard1.png}
+         \caption{Screenshot of process information.}
      	 \label{fig:dashboard1}
+     \end{subfigure}
+     \hfill
+     \begin{subfigure}[b]{0.49\textwidth}
+         \centering
+         \includegraphics[scale=0.17]{dashboard2.png}
+      	 \caption{Screenshot of a workflow.}
+     \end{subfigure}
+     \caption{Screenshot of the Colonies Dashboard tool.}
 \end{figure}
+
+% \begin{figure}[h]
+% 	\centering
+%     \includegraphics[scale=0.23]{dashboard1.png}
+% 	\caption{Process information. The timeline makes it possible to follow the execution of process in realtime.}
+% 	\label{fig:dashboard1}
+% \end{figure}
 
@@
 
-\begin{figure}[h]
-	\centering
-    \includegraphics[scale=0.23]{dashboard2.png}
-	\caption{Screenshot of a workflow in the Colonies dashboard.}
-	\label{fig:dashboard2}
-\end{figure}
+% \begin{figure}[h]
+% 	\centering
+%     \includegraphics[scale=0.23]{dashboard2.png}
+% 	\caption{Screenshot of a workflow in the Colonies dashboard.}
+% 	\label{fig:dashboard2}
+% \end{figure}
 
@@
 \section{Use cases}
+Colonies is currently used commerically by a startup company called RockSigma AB\cite{https://www.rocksigma.com} to implement a compute engine for analysing seismic activities in underground mines.
+This section will not dig into the details of this specific implementation; instead it will provide an outline of potential future use cases in which Colonies could make a considerable impact.
+
 \subsection{Serverless computing}
 \label{faas}
+Serverless computing is a cloud-based architecture that allows developers to build and deploy applications without the need to manage the underlying infrastructure.
+In this model, the cloud service provider dynamically allocates and manages resources, such as computing power and storage, as required by the application.
 
@@
 \end{figure}
-\subsection{Edge computing}
-\subsection{Remote sensing}
-\subsection{HPC integration}
+
+\subsection{Slurm integration}
+\label{hpc_use_case}
 
@@
 
+\subsection{OpenEO example}
+OpenEO \cite{openeo} is an open-source project that aimes at providing a standardized and interoperable solution for accessing, processing, and analyzing Earth observation (EO) data.
+Its primary objective is to facilitate seamless user interaction with diverse data sources and processing backends through a unified API.
+An overview of various OpenEO backends is presented in \cite{openeo_backends}.
+The paper concludes that there is a need for a more decentralized and federated EO data processing platform, and OpenEO represents a significant step in this direction.
+Nevertheless, the paper also acknowledges that the implementation of underlying datacubes, storage solutions, and service plans cannot be entirely abstract away by the OpenEO API.
+
+% \begin{figure}[h]
+% 	\centering
+%     \includegraphics[scale=0.43]{openeo.png}
+% 	\caption{OpenEO integration}
+% 	\label{fig:dashboard2}
+% \end{figure}
+
+Colonies can be used to implement a decentralized, grid-like compute infrastructure for executing EO workloads across diverse platforms.
+By encapsulating computational EO tasks in Colony processes, interoperability among platforms can be significantly enhanced.
+Executors could also be used to provision computational resources if needed.
+For example, a provisioning engine similar to serverless engine presented in \ref{faas} can be used to deploy extra cloud executors to speed up execution of demaning EO workloads, or take advantage of HPC resources as discussed in Section \ref{hpc_use_case}.
+
 \section{Discussion}
+This paper has presented an open-source framework called Colonies, which streamlines execution of computational workloads across platforms.
+Colonies enables creation of a grid-like distributed virtual computer, called a \emph{Colony}, which is a collective machine consisting of distributed executors running on the Internet.
+Unlike many other systems where security sometimes is added later on, Colonies builds on the principle, \emph{security first}.
+
+Public key encryption allows administrators and users to maintain strict control over a colony, even when executors are dispersed throughout the Internet.
+Each colony is governed by a private key that serves as a central authority, analogous to \emph{one ring to rule them all}, making it possible to control a colony as if it was a single machine.
+The zero-trust principle, \emph{trust, but verify} enables development of immune system-like functionality where malfunctional executors can be excluded and replaced with fresh new ones, ensuring that the overall performance and reliability of the system is not compromised.
+
+A significant portion of the paper has focused on how to implement robustness, which allows software components to fail without compromising the entire system.
+In Colonies, robustness is a unique feature that can be leveraged not only to build stable data pipelines, but also to efficiently handle IT operations such as continuous deployments and software upgrades.
+When used together with Kubernetes, Colonies can significantly simplify software updates simply by killing pods in a live system, and let Kubernetes replace them with new versions, assuming the new pods have been thoroughly tested and are compatible with the old version.
+This approach can be further enhanced by applying chaos engineering and randomly killing executors to ensure that the system is capable of gracefully tolerating failures.
+In theory, by controlling the rate at which executors are killed, it would be possible to dynamically scale the system up or down, somewhat similar to programmed cell death mechanisms used in biological systems \footnote{https://en.wikipedia.org/wiki/Programmed\_cell\_death}.
+
+Security and robustness are crucial factors to consider when developing a workflow management system that is capable of functioning across various platforms and handling constantly changing infrastructures.
+Another important mechanism is the use of queues, which makes it possible to separate responsibilities between different system components and ensure that each component can operate independently.
+
+The use of queues also enables a split architecture, where executors can be dynamically managed.
+In Colonies, it is possible to submit a workflow even if no executor is available to execute the workload.
+The processes associated with the workflow are simply enqueued until an executor capable of handling the workload is deployed.
+By designing each executor as a standalone microservice, it becomes easy to scale the system by simply deploying more executors.
+However, if too few executors are available, a queue of pending processes will begin to build up.
+This opens up an exciting possibility to delegate colony management responsibility to another software capable of automatically provisioning and deploying executors whenever they are needed, and then also scale down the system to conserve resources and save costs.
+In line with this, the Cognit project \cite{cognit} aims over the next three years to develop an AI orchestrator capable of automatically managing serverless computing infrastructures.
+
+It has been mentioned in the paper more than once that Colonies is designed to simplify integration with other systems.
+What feature in Colonies is it that makes the integration easier?
+A core idea in the Colonies framework is to package meta information and instructions how to carry out a computation in a database record called a process, and then provide an API to manipulate the process record, e.g.
+assigning it to an executor.
+As Colonies is designed to be stateless and all data is stored in a database, Colonies provides full transparency and third-party can simply lookup the process to retrieve its status.
+As a result, third-party systems do not have to actively monitor submitted jobs, but can instead trust that the data will eventually be processed.
+This approach offloads a significant portion of error handling onto Colonies instead of the third-party software.
+
+Microservices is another design principle of the Colonies framework that simplifies integration.
+As executors are designed to be loosly-coupled and can be implemented in any language (supported by Colonies), the whole system becomes easier to develop and maintain.
+For example, different teams can take on responsibility to develop a certain executor, which can then be used as a part of a complex workflow running across platforms.
+
+As a final remark, it is worth highlighting once again the significant advantage of Colonies as a security solution.
+As a colony is a trusted community of computer software, Colonies can be used to share secrets and credentials among its executor members.
+By storing credentials, such as one-time passwords or authentication keys in process records, executors can just execute a process in run-and-forget style and be able to access third-party infrastructures without the need to store the and manage credentials locally.
+Additionally, by targeting specific executor types when submitting a function specification, a colony can be further segmented so that only a subset of the executors in a colony can access the credentials, further improving the security.
 
@@
 \section{Appendix - Colonies operations}
-
 \begin{table}[h]
@@
 		\cmidrule(r){1-4}
-        Operations                 & Role         & Assignment & Note \\
+        Operation                  & Role         & Assign required & Note \\
 		\midrule
-        \(add\_colony\)            & Server owner &            & Registers a new colony to the Colonies server. \\
-        \(delete\_colony\)         & Server owner &            & Unregisters a colony from the Colonies server. \\
+        \(add\_colony\)            & Server owner &            & Registers a new colony. \\
+        \(delete\_colony\)         & Server owner &            & Unregisters a colony. \\
         \(add\_executor\)          & Colony owner &            & Register a new executor to a colony. \\
         \(delete\_executor\)       & Colony owner &            & Unregister a executor from a colony. \\
-        \(approve\_executor\)      & Colony owner &            & Approve an executor to get process assignments. \\
-        \(reject\_executor\)       & Colony owner &            & Disaprove an executor from getting process assignments. \\
+        \(approve\_executor\)      & Colony owner &            & Approve an executor. \\
+        \(reject\_executor\)       & Colony owner &            & Disaprove an executor. \\
         \(get\_executors\)         & Executor     &            & List all executors member of a colony. \\
@@
         \(submit\)                 & Executor     &            & Submit a function specification. \\
-        \(assign\)                 & Executor     &            & Assign and binds a waiting process to an executor. \\
+        \(assign\)                 & Executor     &            & Assign a waiting process to an executor. \\
         \(unassign\)               & Executor     & x          & Unassign a process from a executor. \\
@@
         \(fail\)                   & Executor     & x          & Close a process as failed. \\
-        \(get\_processes\)         & Executor     &            & List all waiting, successful or failed processes. \\
+        \(get\_processes\)         & Executor     &            & List waiting, successful or failed processes. \\
         \(get\_process\)           & Executor     &            & Get info about a process. \\

### 2023-04-10 paper_2023/paper.tex
wip

@@
 
+\begin{figure}[h]
+	\centering
+    \includegraphics[scale=0.44]{cognite_faas.png}
+	\caption{A Colonies-based FaaS serverless computing platform.}
+	\label{fig:faas}
+\end{figure}
+
 \section{Use cases}
@@
 Serverless computing is a cloud-based architecture that allows developers to build and deploy applications without the need to manage the underlying infrastructure.
-In this model, the cloud service provider dynamically allocates and manages resources, such as computing power and storage, as required by the application.
+In this model, the cloud service provider dynamically allocates and manages resources such as adding computing power or storage required by an application.
+Function-as-a-Service (FaaS) is a serverless computing model where developers provide a code snippet that should be executed, and the cloud provider takes care of scaling, configuration and deployment.
+FaaS is often used in event-driven architectures where functions are triggered by specific events, thus eliminating the need to keep servers continuously running.
+The Cognit project \cite{cognit} aims to extend the FaaS concept by supporting geographically distributed edge servers in addition to cloud resources.
+The project is also going to develop an AI-enabled orchestrator that can decide where to deploy FaaS code.
 
-\begin{figure}[h]
-	\centering
-    \includegraphics[scale=0.44]{cognite_faas.png}
-	\caption{A Colonies-based FaaS serverless computing platform.}
-	\label{fig:ha_deployment}
-\end{figure}
+Colonies offer functionality to call functions in remote executors, a fundamental requirement to implement a FaaS platform.
+Colonies could also provide service discovery and provide security isolation to prevent non-authorized function calls.
+In some cases, it could make sense to manually manage a colony by an administrator, and manually registering and deploying executors.
+Alternatively, a software solution could be developed to automate the administration and automatically deploy or undeploy executors.
+Such an approach would be very similar to a Function-as-a-Service (FaaS) platform.
+
+Figure \ref{fig:faas} depicts a Colonies-based FaaS platform.
+Similar to AWS Lambda or Google Cloud Functions, a developer would need to upload a code snippet, which would then be packaged and published as a container.
+To invoke a function, a function specification needs to be submitted  to the Colonies server indicating which function to call.
+
+The advantage of Colonies is that it introduces an extra level of separation by offering a queuing system.
+If an executor is unavailable, the process is simply enqueued in the process table, enabling development of a loosely coupled system.
+A software agent could then monitor pending processes on the Colonies server, predict the system load, and automatically deploy executors to meet the load, resulting in a highly efficient and scalable system.
+In the future, the Cognit AI orchestrator could also be used together with Colonies to reduce energy consumption or minimize network latency by efficiently utilizing edge servers.
 
@@
 Executors could also be used to provision computational resources if needed.
-For example, a provisioning engine similar to serverless engine presented in \ref{faas} can be used to deploy extra cloud executors to speed up execution of demaning EO workloads, or take advantage of HPC resources as discussed in Section \ref{hpc_use_case}.
+For example, automatically provision cloud resources and setup EO software, or deploy workflows at HPC systems to handle high-demand EO workloads.
 

### 2023-04-10 paper_2023/paper.tex
wip

@@
 
-Based on a stateless microservice architecture, complex workflows to be broken down into composable functions.
+Based on a stateless microservice architecture, complex workflows can be broken down into composable functions.
 These composable functions can be implemented in any programming language and be executed by distributed executors deployed across platforms, anywhere on the Internet such as IoT, cloud, edge, and supercomputers.
-A zero-trust security protocol enables a collection of distributed executors to operate as a unified entity, thus establishing seamless compute continuums across platforms.
+A zero-trust security protocol enables a collection of distributed executors to operate as a unified entity, establishing seamless compute continuums across platforms.
 

### 2023-04-11 paper_2023/paper.tex
wip

@@
 
-\subsection{Slurm integration}
+\subsection{HPC integration}
 \label{hpc_use_case}
+Slurm \cite{slurm} is a workload manager and job scheduler commonly used in HPC environments.
+In constrast to Colonies, it provides functionality to reserve resources.
+This is particularly useful when running complex simulations that requires a large number of of CPUs or a large amount of memory.
+
+A challenge associated with Slurm and HPC environments in general is that they can be difficult to set up and use.
+Submitting a job typically requires users to develop job scripts and manually transfer data to the HPC environment before a job can be executed.
+Compared to cloud platforms, Slurm is less mature when it comes to APIs and users have to develop their own automation tools, which can make the usage more time-consuming and manual.
+
+Colonies could be used to fill the gap between cloud and HPC and create workflows that can automate processing of data in the cloud while also taking advantage of the vaste compuational capacity offered by HPC.
+For example, a company could develop a workflow that automatically copy training data to a HPC center, train a neural network, move the result back to the cloud, and finally re-deploy a Kubernetes service to use the new model.
+
+Figure \ref{hpc_workflow} shows an example how Colonies could be used in a HPC environment.
+Several executors runs at the HPC system, each with their own responsibilities.
+The Container Importer Executor is responsible for converting container images to a format that is supported by Slurm, e.g.
+Singularity containers\footnote{https://docs.sylabs.io}.
+The Data Importer Executors downloads data to the HPC file system.
+The Slurm Executor is responsible for submitting and overseeing Slurm jobs, while the Data Exporter executor is responsible for uploading result data to the cloud.
+
+A challenge developing cross-platform workflows is the integration with different autentication services.
+For example, the Data Importer may need to access an S3 services to fetch data.
+The Container Importer must authenticate with a Docker registry to be able to pull a Docker container to the HPC system.
+Generally, it is not straigtforward connect to the HPC system.
+In many cases, users have to manually set up SSH tunnels.
+
+When using Colonies, users would not direclty connect to the HPC system, but indirectly control it by submitting Colonies processes that are subsequently picked up by HPC executors.
+
+picked up the HPC executors.
+Since the security model offered by Colonies allows for different executors to trust each other, Colonies also provide a way of sharing secrets, e.g.
+S3 secrets.
+As the executors running at the HPC system would get all needed information in the assigned process records, they become easy to configure and wouldn't need to store credentials locally at the HPC system.
 
@@
 	\caption{HPC workflow.}
-	\label{fig:dashboard2}
+	\label{fig:hpc_workflow}
 \end{figure}
@@
 \subsection{OpenEO example}
-OpenEO \cite{openeo} is an open-source project that aimes at providing a standardized and interoperable solution for accessing, processing, and analyzing Earth observation (EO) data.
+OpenEO \cite{openeo} is an open-source project that aims at providing a standardized and interoperable solution for accessing, processing, and analyzing Earth observation (EO) data.
 Its primary objective is to facilitate seamless user interaction with diverse data sources and processing backends through a unified API.
@@
 
-Colonies can be used to implement a decentralized, grid-like compute infrastructure for executing EO workloads across diverse platforms.
-By encapsulating computational EO tasks in Colony processes, interoperability among platforms can be significantly enhanced.
-Executors could also be used to provision computational resources if needed.
-For example, automatically provision cloud resources and setup EO software, or deploy workflows at HPC systems to handle high-demand EO workloads.
+Colonies offer a promising solution for implementing a decentralized and grid-like compute infrastructure, capable of executing Earth Observation (EO) workloads across diverse platforms.
+By encapsulating computational EO tasks into Colonies processes, interoperability among platforms can be further enhanced.
+For instance, it becomes possible to create a broker-based architecture that enables OpenEO applications to submit job specifications indirectly to a Colonies server.
+These specifications are then picked up by executors connecting to different backend providers, e.g.
+by an executor connected to a Data and Information Access Services (DIAS).
+Executors can also be leveraged to provision computational resources.
+For example, automatically provision AWS cloud resources or deploy workflows at HPC systems to handle high-demand EO workloads.
 
@@
 This paper has presented an open-source framework called Colonies, which streamlines execution of computational workloads across platforms.
-Colonies enables creation of a grid-like distributed virtual computer, called a \emph{Colony}, which is a collective machine consisting of distributed executors running on the Internet.
-Unlike many other systems where security sometimes is added later on, Colonies builds on the principle, \emph{security first}.
+Colonies enables creation of a \emph{grid-like distributed virtual computer}, called a \emph{Colony}, which can be seen as a collective computer machine consisting of distributed executor software running on the Internet.
+Unlike many other systems where security sometimes is added later on, Colonies builds on the principle: \emph{security first}.
 
@@
 As a final remark, it is worth highlighting once again the significant advantage of Colonies as a security solution.
-As a colony is a trusted community of computer software, Colonies can be used to share secrets and credentials among its executor members.
-By storing credentials, such as one-time passwords or authentication keys in process records, executors can just execute a process in run-and-forget style and be able to access third-party infrastructures without the need to store the and manage credentials locally.
-Additionally, by targeting specific executor types when submitting a function specification, a colony can be further segmented so that only a subset of the executors in a colony can access the credentials, further improving the security.
+As a colony is a trusted community of execturs, it can be used to share secrets and credentials among its executor members.
+By storing credentials, such as one-time passwords or authentication keys in process records, executors can just execute process in run-and-forget style and still be able to access third-party infrastructures without the need to store and manage credentials locally.
+Additionally, by targeting specific executor types when submitting a function specification, a colony can be further segmented so that only a subset of the executors in a colony can access the credentials, further improving security.
 

### 2023-04-11 paper_2023/paper.tex
first draft

@@
 This will be further discussed in Section \ref{faas}.
-However, in contrast to previous work, Colonies is platform independent and does not require Kubernetes.
+However, in contrast to previous work, Colonies is designed to be platform independent and does therefore not require Kubernetes.
 By using a zero-trust security protocol, functions can be securely executed by distributed executors deployed anywhere on the Internet.
@@
 
-Although microservices have attractive properties and simplify both development and deployment of executors, they also introduce complexity to the Colonies servers.
+Although microservices have attractive properties and simplify both development and deployment of executors, they also introduce additional complexity that must be managed by the Colonies servers.
 To implement a workflow management framework supporting distributed microservices, the following challenges must be addressed:
@@
 
-\item \textbf{Service discovery:} The Colonies framework must enable executors to dynamically register and deregister to Colonies servers.
+\item \textbf{Service discovery:} The Colonies framework must enable executors to dynamically register and deregister from Colonies servers.
 It must be possible to deploy executors anywhere on the Internet, even behind firewalls.
@@
 A fundamental design principle is statelessness, which makes Colonies simpler, more reliable and scalable, and easier to implement.
-To ensure reliability and data consistency, distributed consensus algorithms are needed to offer high-availability and handle server crashes.
+To ensure reliability and data consistency, a distributed consensus algorithm is needed to handle high-availability and server crashes.
 The remainder of this section describes how the Colonies framework is implemented.
-The next discusses the role of queues to support batch jobs and realtime processing, but also how queues can be used to enable loose coupling of system components.
+The next discusses the role of queues to support batch jobs and realtime processing, but also how to implement a loosely coupled system.
 
@@
 To assign a process to an executor, the Colonies server searches in the process table to find a process matching a waiting executor.
-By using the \emph{priority time} column to sort the processes, the process table serves as a queue.
-This can be done in SQL by utilizing the \emph{order by} to sort processes according to their submission time.
+By using the \emph{priority time} column to sort the processes, the process table can function as a queue.
+This can be done in SQL by utilizing the \emph{order by} clause to sort processes according to their submission time.
 To make it possible to handle priorities, the submission time is adjusted so that higher priority processes are processed before lower priority processes.
@@
 To minimize downtime, Colonies supports high-availability deployments.
-If one Coloniser server crashes, an executor simply need to resend the failed request, which will then be served by another Colonies server replica.
+If one Colonies server crashes, an executor simply need to resend the failed request, which will then be served by another Colonies server replica.
 However, by introducing multiple Colonies servers, there is again a risk of race conditions when assigning processes to executors.
-This means that the Colonies server replicas must coordinate which replica server incoming assign requests so that precisely one executor is assigned to a process.
+This means that the Colonies must coordinate which replica server should serve incoming assign requests so that precisely one executor is assigned to a particular process.
 
 Raft \cite{raft} is a consensus algorithm specifically designed to manage a replicated log within a distributed system.
-It functions within a cluster of servers, where a single server takes on the role of leader while the remaining servers act as followers.
+Raft makes it possible to implement distributed leader elections where a single server takes on the role of leader while the remaining servers act as followers.
 The leader is responsible for managing the replicated log, processing client requests, and replicating entries to the followers.
 Followers passively replicate the leader's log and participate in leader elections.
-The leadership can change over time due to elections triggered by timeouts or other factors.
 
-Adding Raft with the Colonies framework allows incoming assign requests to be directed towards the leading Colonies server, thereby ensuring that only one Colonies server replica handles such requests.
+By using the Raft protocol, Colonies can redirect incoming \emph{assign} requests to the leading Colonies server, thereby ensuring that only one Colonies server replica handles such requests.
+A new leader is automatically elected in the event that a Colonies server replica fails.
 Figure \ref{fig:ha_deployment} shows an overview of a high-availability Colonies deployment.
-A new leader is elected in the event that a Colonies server replica fails.
-The Raft protocol also enables seamless updates to the Colonies server software by making it possible to upgrade each replica individually.
-Consequently, Colonies is well-suited for Kubernetes deployments, ensuring high levels of availability and fault tolerance.
+In addition, the Raft protocol makes it possible to seamlessly update Colonies software by making it possible to upgrade each Colonies server replica individually.
+Consequently, making Colonies well-suited for Kubernetes deployments.
 
@@
 To control the order processes can be assigned, Colonies sets a flag, \emph{wait for parents} to prevent processes to be assigned to an executor before their dependencies have been satisfied.
-Note that processes may run in parallel if a sufficient number of executors are available.
+Note that processes may run in parallel if a sufficient number of executors are available, assuming their parent processes have finished.
 
@@
 If all parents of a child process have finished, the server updates the process table in the database by setting the flag \emph{wait for parents} to false, which allows child processes to be assigned to executors.
-This operation is also done statelessly when processing the close request.
+This operation is also done statelessly when processing the \emph{close} request.
 
@@
 As mentioned before, when an executor is assigned a process it obtains exclusive rights to that process within the process table.
-This exclusivity also means that it is possible for an exector to dynamically add new children processesto  the workflow DAG on the fly.
+This exclusivity also means that it is possible for an exector to dynamically add new children processes to the workflow DAG on the fly.
 As a result, it becomes possible to implement MapReduce \cite{mapreduce} data processing patterns similar to Hadoop \cite{hadoop}.
@@
 It can for example be used to automate tasks such as routinely fetching data from external systems.
-To make Colonies robust and easy to implement it is essential to make Cron workflows stateless.
+To make Colonies robust and easy to implement it is essential that Cron workflows also are managed stateless.
 This can be achieved by introducing an additional database table to store information about Cron workflow, and let the elected Colonies server leader assume the responsibility of managing Cron workflows through a two-step process.
@@
 For example, an external service downloads satellite images and can use a generator to automatically triggered a workflow when 10 images are collected.
-This is achieved by sending a special \emph{pack request} containing input data to a Colonies server, targeting a certain generator.
+This is achieved by sending a special \emph{pack} request containing input data to a Colonies server, targeting a certain generator.
 
@@
 Automatic monitoring can for example be implemented using anomaly detection algorithms and machine learning to grade the performance of each executor.
-Administrators can then be automatically notified if the grade of an executor exceeds a certain threshold, enabling them to take appropriate action.
+Administrators can then be automatically notified if the grade of an executor exceeds a certain threshold, enabling them to take appropriate actions.
 
@@
 \subsection{Python SDK}
-A colonies application typically consists of a network of executors written in different programming languages and deployed across multiple platforms.
+A Colonies application typically consists of a set executors written in different programming languages and deployed across multiple platforms.
 The executors are small microservices that interact with one another by submitting function specifications or workflows, which are then executed by other executors that are part of the same colony.
@@
 Registering an executor can only be carried out by the colony owner with access to the colony private key.
-Code in listening \ref{code:regexecutor} illustrates how to register an executor and which functions a specific executor is capable of running.
+Code in Listening \ref{code:regexecutor} illustrates how to register an executor and which functions a specific executor is capable of running.
 In the example, an executor of type \emph{helloworld\_executor} is registered to run a function named \emph{helloworld}.
@@
       	 \caption{Screenshot of a workflow.}
+     	 \label{fig:dashboard2}
      \end{subfigure}
@@
 
-Figure \ref{fig:dashboard1} depicts a screenshot of the Colonies dashboard tool, which enables users to monitor the status of processes and executors.
-The timeline depicted in the figure allows for real-time monitoring of the execution of processes, similar to tracking a physical package sent through a postal service.
+Figure \ref{fig:dashboard1} shows a screenshot of the Colonies dashboard tool, which enables users to monitor the status of processes and executors.
+The timeline depicted in the figure allows for real-time monitoring of process execution, similar to tracking a physical package sent through a postal service.
 As previously discussed, it's also possible to subscribe to a process and receive information about its progress, which can be used as another point of integration.
-For example, sending a notification on Slack or by email if a process fails.
+For example, sending a notification on Slack if a process fails.
 
 \subsection{Workflows}
-A workflow is just a set of processes with dependencies represented as a DAG internally in Colonies.
+A workflow is just a set of processes with dependencies, which is represented as a DAG internally in Colonies.
 Workflows can either be submitted using SDKs or the Colonies CLI, which supports submission of JSON files.
@@
 In addition to monitoring processes and executors, the Colonies dashboard can be used to visualize workflows.
-Figure \ref{fig:dashboard2} illustrates the DAG created when submitting the workflow shown in Listing \ref{code:workflow_spec}.
+Figure \ref{fig:dashboard2} shows a DAG created when submitting the workflow code shown in Listing \ref{code:workflow_spec}.
 This functionality makes it possible to follow workflow execution progress.
@@
 \section{Use cases}
-Colonies is currently used commerically by a startup company called RockSigma AB\cite{https://www.rocksigma.com} to implement a compute engine for analysing seismic activities in underground mines.
+Colonies is currently used commerically by a startup company called RockSigma AB\footnote{https://www.rocksigma.com} to implement a compute engine for analysing seismic activities in underground mines.
 This section will not dig into the details of this specific implementation; instead it will provide an outline of potential future use cases in which Colonies could make a considerable impact.
@@
 In this model, the cloud service provider dynamically allocates and manages resources such as adding computing power or storage required by an application.
-Function-as-a-Service (FaaS) is a serverless computing model where developers provide a code snippet that should be executed, and the cloud provider takes care of scaling, configuration and deployment.
+Function-as-a-Service (FaaS) is a serverless computing model where developers provide a code snippet that should be executed, and the cloud provider takes care of scaling, configuration and deployment of that code snippet.
 FaaS is often used in event-driven architectures where functions are triggered by specific events, thus eliminating the need to keep servers continuously running.
-The Cognit project \cite{cognit} aims to extend the FaaS concept by supporting geographically distributed edge servers in addition to cloud resources.
+The Cognit project \cite{cognit} aims to extend the FaaS concept by supporting geographically distributed edge servers in addition to cloud server resources.
 The project is also going to develop an AI-enabled orchestrator that can decide where to deploy FaaS code.
@@
 Colonies offer functionality to call functions in remote executors, a fundamental requirement to implement a FaaS platform.
-Colonies could also provide service discovery and provide security isolation to prevent non-authorized function calls.
+Colonies could also provide service discovery and security isolation to prevent non-authorized function calls.
 In some cases, it could make sense to manually manage a colony by an administrator, and manually registering and deploying executors.
@@
 Figure \ref{fig:faas} depicts a Colonies-based FaaS platform.
-Similar to AWS Lambda or Google Cloud Functions, a developer would need to upload a code snippet, which would then be packaged and published as a container.
+Similar to AWS Lambda or Google Cloud Functions, developers would need to upload code snippets, which would then be packaged and published as Docker containers.
 To invoke a function, a function specification needs to be submitted to the Colonies server indicating which function to call.
 
-The advantage of Colonies is that it introduces an extra level of separation by offering a queuing system.
-If an executor is unavailable, the process is simply enqueued in the process table, enabling development of a loosely coupled system.
+The advantage of Colonies for implementing a FaaS platform is that it introduces an extra level of separation and the possibility to develop a loosely coupled system.
+For example, if an executor is missing, the process is simply enqueued in the process table.
 A software agent could then monitor pending processes on the Colonies server, predict the system load, and automatically deploy executors to meet the load, resulting in a highly efficient and scalable system.
-In the future, the Cognit AI orchestrator could also be used together with Colonies to reduce energy consumption or minimize network latency by efficiently utilizing edge servers.
+In the future, the AI orchestrator developed in the Cognit project could also be used together with Colonies to reduce energy consumption or minimize network latency by efficiently utilizing edge servers.
 
@@
 Slurm \cite{slurm} is a workload manager and job scheduler commonly used in HPC environments.
-In constrast to Colonies, it provides functionality to reserve resources.
-This is particularly useful when running complex simulations that requires a large number of of CPUs or a large amount of memory.
+In contrast to Colonies, it provides functionality to reserve resources.
+This is particularly useful when running complex simulations that require a large number of CPUs or a large amount of memory.
 
@@
 Compared to cloud platforms, Slurm is less mature when it comes to APIs and users have to develop their own automation tools, which can make the usage more time-consuming and manual.
-
-Colonies could be used to fill the gap between cloud and HPC and create workflows that can automate processing of data in the cloud while also taking advantage of the vaste compuational capacity offered by HPC.
-For example, a company could develop a workflow that automatically copy training data to a HPC center, train a neural network, move the result back to the cloud, and finally re-deploy a Kubernetes service to use the new model.
-
-Figure \ref{hpc_workflow} shows an example how Colonies could be used in a HPC environment.
-Several executors runs at the HPC system, each with their own responsibilities.
-The Container Importer Executor is responsible for converting container images to a format that is supported by Slurm, e.g.
-Singularity containers\footnote{https://docs.sylabs.io}.
-The Data Importer Executors downloads data to the HPC file system.
-The Slurm Executor is responsible for submitting and overseeing Slurm jobs, while the Data Exporter executor is responsible for uploading result data to the cloud.
-
-A challenge developing cross-platform workflows is the integration with different autentication services.
-For example, the Data Importer may need to access an S3 services to fetch data.
-The Container Importer must authenticate with a Docker registry to be able to pull a Docker container to the HPC system.
-Generally, it is not straigtforward connect to the HPC system.
-In many cases, users have to manually set up SSH tunnels.
-
-When using Colonies, users would not direclty connect to the HPC system, but indirectly control it by submitting Colonies processes that are subsequently picked up by HPC executors.
-
-picked up the HPC executors.
-Since the security model offered by Colonies allows for different executors to trust each other, Colonies also provide a way of sharing secrets, e.g.
-S3 secrets.
-As the executors running at the HPC system would get all needed information in the assigned process records, they become easy to configure and wouldn't need to store credentials locally at the HPC system.
 
@@
 \end{figure}
+
+Colonies could be used to fill the gap between cloud and HPC and create workflows that can automate processing of data in the cloud while also taking advantage of the vaste computational capacity offered by HPC.
+For example, a company could develop a workflow that automatically copy training data to a HPC center, train a neural network, move the result back to the cloud, and finally re-deploy a Kubernetes service to use the new model.
+
+Figure \ref{fig:hpc_workflow} shows an example how Colonies could be used in a HPC environment.
+Several executors run at the HPC system, each with their own responsibilities.
+The Container Importer Executor is responsible for converting container images to a format that is supported by Slurm, e.g.
+Singularity containers\footnote{https://docs.sylabs.io}.
+The Data Importer Executors downloads data to the HPC file system.
+The Slurm Executor is responsible for submitting and overseeing Slurm jobs, while the Data Exporter executor is responsible for uploading result data to the cloud.
+
+\begin{figure}[h]
+	\centering
+    \includegraphics[scale=0.43]{hpc_overlay.png}
+	\caption{A global grid-like master-control submission system for HPC.}
+	\label{fig:hpc_overlay}
+\end{figure}
+
+In general, integrating with various authentication services presents a major challenge when developing HPC workflows.
+For example, the Data Importer may need to access an S3 service to fetch data.
+The Container Importer must authenticate with a Docker registry to be able to pull a Docker container to the HPC system.
+In addition, it is not straightforward to connect to the HPC system.
+In many cases, users have to manually connect using SSH terminals, making it time-consuming and difficult to automate HPC workflows.
+
+With Colonies, users can indirectly control HPC systems by submitting Colonies processes that are subsequently picked up by HPC executors, without requiring direct connection to the HPC system.
+Essentially, using Colonies as a cloud-based submission system for HPC, connecting different HPC centers together in a grid-like overlay network.
+Figure \ref{fig:hpc_overlay} depicts such a submission system.
+
+Additionally, the security model provided by Colonies allows different executors to trust one another, thereby enabling the sharing of confidential information, such as S3 secrets.
+As credentials are stored in the assigned process record, there is no need to store them locally on the HPC system.
+As a result, HPC executors become more straightforward to configure and deploy, further simplifying creation of HPC workflows.
 
@@
 
-Colonies offer a promising solution for implementing a decentralized and grid-like compute infrastructure, capable of executing Earth Observation (EO) workloads across diverse platforms.
+Colonies could offer a promising solution for implementing a decentralized and grid-like compute infrastructure, capable of executing Earth Observation (EO) workloads across diverse platforms.
 By encapsulating computational EO tasks into Colonies processes, interoperability among platforms can be further enhanced.
 For instance, it becomes possible to create a broker-based architecture that enables OpenEO applications to submit job specifications indirectly to a Colonies server.
-These specifications are then picked up by executors connecting to different backend providers, e.g.
-by an executor connected to a Data and Information Access Services (DIAS).
-Executors can also be leveraged to provision computational resources.
+These specifications could then be picked up by executors connected to different backend providers, e.g.
+a Data and Information Access Services (DIAS) executor.
+Executors can also be used to leveraged to provision computational resources.
 For example, automatically provision AWS cloud resources or deploy workflows at HPC systems to handle high-demand EO workloads.
@@
 This opens up an exciting possibility to delegate colony management responsibility to another software capable of automatically provisioning and deploying executors whenever they are needed, and then also scale down the system to conserve resources and save costs.
-In line with this, the Cognit project \cite{cognit} aims over the next three years to develop an AI orchestrator capable of automatically managing serverless computing infrastructures.
 
@@
 As a final remark, it is worth highlighting once again the significant advantage of Colonies as a security solution.
-As a colony is a trusted community of execturs, it can be used to share secrets and credentials among its executor members.
-By storing credentials, such as one-time passwords or authentication keys in process records, executors can just execute process in run-and-forget style and still be able to access third-party infrastructures without the need to store and manage credentials locally.
+As a colony is a trusted community, an executor can share secrets with other colony members.
+By storing credentials, such as one-time passwords or authentication keys in process records, executors can just execute process in run-and-forget style and be able to access third-party infrastructures without the need to store and manage credentials locally.
 Additionally, by targeting specific executor types when submitting a function specification, a colony can be further segmented so that only a subset of the executors in a colony can access the credentials, further improving security.

### 2023-04-12 paper_2023/paper.tex
wip

@@
 To address these challenges, the paper presents an open-source framework called Colonies\footnote{https://github.com/colonyos/colonies}, which facilitates execution of computational workloads across a diverse range of platforms.
-Colonies enables development of a \emph{grid-like distributed virtual computer}, called a \emph{Colony}, which can effortlessly integrate with any third-party application (or other workflow management systems).
+Colonies enables development of a \emph{decentralized virtual computer system}, called a \emph{Colony}, which can effortlessly integrate with any third-party application or other workflow systems.
 
@@
 Colonies is built around a loosely-coupled microservices architecture that separates workflow definitions from implementation and deployment.
-The primary objective is to establish a platform where monolithic workflows can be decomposed into independent microservices that can easily be integrated with other systems or applications running anywhere on the Internet.
+The primary objective is to establish a platform where workflows can be decomposed into independent microservices that can easily be integrated with other systems or applications running anywhere on the Internet.
 The remainder of the paper describes the Colonies framework and how it can be used to create robust and scalable AI systems across platforms.
@@
 \section{Related work}
-Workflow management has been extensively studied in both academic and industrial settings with numerous approaches \cite{service_wfs, schmitt2022workflow, GarciaRepresa1740746, Ouyang2010, NIKOLOV2021100440, workflow_in_bigdata} proposed to address the challenges in this field.
+Workflow management has been extensively studied in both academic and industrial settings with numerous approaches \cite{scafe, synapse, service_wfs, schmitt2022workflow, GarciaRepresa1740746, Ouyang2010, NIKOLOV2021100440, workflow_in_bigdata} proposed to address the challenges in this field.
 Recently, Apache Airflow \cite{apache_airflow} has become a popular open-source workflow management system for handling data engineering pipelines.
 Like Colonies, Apache Airflow enables developers to create custom operators and executors that can be integrated with various systems.
-Additionally, Apache Airflow offers an HTTP API that makes it possible to develop software development kits (SDKs) in various programming languages.
-However, Apache Airflow does not rely on a queuing system.
-Instead, it must be integrated with a message broker, such as RabbitMQ \cite{rabbitmq} or Kafka \cite{apache_kafka}, to implement task queues, resulting in a more complex architecture than Colonies.
+Additionally, Apache Airflow offers an HTTP API that makes it possible to develop Software Development Kits (SDKs) in various programming languages.
+However, Apache Airflow must be integrated with a message broker, such as RabbitMQ \cite{rabbitmq} or Kafka \cite{apache_kafka} to implement task queues, resulting in a more complex architecture than Colonies.
 Furthermore, Colonies is based on a distributed microservice architecture that makes it more suitable for DevOps software development.
@@
 Argo \cite{argowf} is an open source container-native workflow engine for orchestrating parallel jobs on Kubernetes.
-It can be used for running CI/CD pipelines or compute intensive machine learning or data processing tasks where each job runs as a container.
+It can be used for running CI/CD pipelines or compute intensive machine learning or data processing jobs where each job runs as a container.
 In contrast, Colonies offers a more versatile approach, allowing jobs to be launched within an already started container.
@@
 It is based on OpenFaaS \cite{openfaas}, which allows functions to run on Kubernetes clusters.
-The serverless workflow project \cite{serverlessworkflows} aims at providing a vendor-neutral workflow DSL.
-Synapse \cite{synapse} is a workflow management system similar to Argo, but that implements the serverless workflow specification.
-A similar engine is proposed in \cite{scafe}.
+
 Colonies can also be used to implement serverless workflow management systems.
@@
 The vision is to create a distributed architecture where microservices can reside anywhere on the Internet and still function as a cohesive unit to execute cross-platform workloads.
+o
 
@@
 A process contains meta-information about computations executed by remote computer programs, referred to as \emph{distributed executors}.
-Specifically, a process contains a definition of a function to be executed, as well as contextual information such as execution status, including the result of the computation.
+Specifically, a process contains a definition of a function to be executed as well as contextual information such as execution status, including the result of the computation.
 One can think of a process as a digital twin of a real computing process.
@@
 For instance, a data engineering team may handle the implementation of pre-processing functions, while a data science team oversees machine learning functions, and another team manages visualization or customer integration etc.
+
 \item \textbf{Scalability:} Executors can be deployed independently, which enables horizontal scaling. This allows for better resource utilization, parallelism, and improved performance.
+
 \item \textbf{Resilience:} The failure of a single executor does not compromise the entire application or workflow.
@@
 If an executor crashes during execution, the process is automatically reassigned to another executor.
+
 \item \textbf{DevOps and Continuous Integration:} Executors' inherent fault-tolerant design permits changes to individual executors without impacting the entire system.
 This makes it possible to seamlessly update the system.
+
 \item \textbf{Technology agnostic:} Executors can be implemented in any programming language, facilitating seamless integration with other applications and systems.
+
 \item \textbf{Decentralized governance:} As the Colonies framework is technology-agnostic, different software development teams can make independent technology and design decisions when developing executors, promoting greater flexibility and adaptability.
@@
 
-Although microservices have attractive properties and simplify both development and deployment of executors, they also introduce additional complexity that must be managed by the Colonies servers.
+Although microservices have attractive properties and can simplify both development and deployment of executors, they also introduce additional complexity that must be managed by the Colonies framework.
 To implement a workflow management framework supporting distributed microservices, the following challenges must be addressed:
@@
 \item \textbf{Process management:} Colonies must be able to distribute processes among available executors based on their capabilities and current workload.
-This involves assigning processes to the most appropriate executor and then monitoring process progress.
+This involves assigning processes to the most appropriate executor and then monitoring execution progress.
 
@@
 The remainder of this section describes how the Colonies framework is implemented.
-The next discusses the role of queues to support batch jobs and realtime processing, but also how to implement a loosely coupled system.
+The next discusses the role of queues to support batch jobs and realtime processing, but also queues roles in implementing loosely coupled systems.
 
@@
 For example, HTTP APIs can be used to abstract away implementation detail and provide a clear and simple interface for interacting with a particular service.
-However, HTTP protocols alone are insufficient for handling dynamic environments where components frequently fail or the underlying infrastructure is constantly changing.
-To address such environments, an alternative mechanism is necessary.
+However, HTTP APIs alone are insufficient for handling dynamic environments where components frequently fail or the underlying infrastructure is constantly changing.
+To address such environments, an alternative mechanism is needed.
 
 Queues enable software services to communicate indirectly by acting as a buffer between them.
-Queues make it possible to decouple each executor and make them operate independently, e.g.
-an executor can be updated or replaced without affecting other executors.
+Queues make it possible to decouple executor and make them operate independently, for example an executor can be updated or replaced without affecting other executors.
 Queues also enable asynchronous communication between executors, enabling them to process tasks at their own pace.
 This ensures that slower executors do not bottleneck faster ones, leading to a more efficient and scalable system.
-Most importantly, queues enable load balancing by distributing tasks among multiple executors, thus making it possible to parallelize workflow execution.
+Most importantly, queues enable load balancing by distributing tasks among executors, thus making it possible to parallelize workflow execution.
 
-Queues can be implemented in different ways, and while message brokers are a common solution, Colonies adopt an alternative strategy and leverage a standard database and query it for tasks to assign to different executors.
+Queues can be implemented in different ways.
+While message brokers are a common solution, Colonies adopt an alternative strategy and leverage a standard database and query it for processes to assign to different executors.
 One key advantage of this approach is that it enables fine-grained process assignments, making it possible to assign specific processes to particular executors.
-For instance, an executor of the browser type can be limited to only executing processes in web applications.
+For instance, an executor of the browser type can be limited to only executing processes capable of running in web browsers.
 This level of granularity cannot easily be implemented using message brokers which generally do not offer introspection of queues, or provide the ability to pull specific messages out of the queue.
 Generally, the only way to retrieve a specific message is to pull all messages from the queue, obtain the message, and then place all remaining messages back into the queue in the same order.
-In contrast, a database can function as a queue and a query can match any columns thus making it possible to assign specific executors to specific processes.
+In contrast, a database can function as a queue and a query can match any columns making it possible to assign specific executors to specific processes.
 
@@
 Once submitted, other executors can connect to the servers to receive process execution assignments.
-A new process entry is added to the process table database when a function specification is submitted to the Colonies server.
+A new process entry is automatically added to the process table database when a function specification is submitted to the Colonies server.
 
@@
 When an executor connects to the Colonies server, the server hangs the incoming HTTP connection\footnote{An alternative protocol is to use WebSockets or gPRC to communicate with the Colonies server.} until the executor is assigned a process, or until a connection timer expires.
-Note that the Colonies server does not connect to the executors.
+The Colonies server never connects to the executors.
 Rather, it is the responsibility of the executors to connect to the Colonies server.
@@
 When registering, executors have to specify to the Colonies server which functions they are capable of executing.
-The Colonies server then makes sure that the conditions (i.e requirements) part of the function specification matches the capability of an executor.
+The Colonies server then makes sure that the conditions part of the function specification matches the capability of an executor.
 
@@
 This can be done in SQL by utilizing the \emph{order by} clause to sort processes according to their submission time.
+
 To make it possible to handle priorities, the submission time is adjusted so that higher priority processes are processed before lower priority processes.
@@
 	\centering
-    \includegraphics[scale=0.6]{raft.png}
+    \includegraphics[scale=0.5]{raft.png}
 	\caption{High-availability deployment of Colonies.}
@@
 The primary objective of all Colonies API requests is to alter some state stored in the database or retrieve information from the database.
-The Colonies framework is designed to be stateless, meaning that the Colonies server does not keep any information in memory between requests.
+The Colonies framework is designed to be stateless, meaning that Colonies servers do not keep any information in memory between requests.
 Each request is handled independently without relying on information from previous requests.
@@
 However, synchronization can also slow down execution as only one thread can access critical sections at a time.
-By carefully designing multithreaded applications and employing the right synchronization techniques, it is possible to minimize the performance impact while still ensuring data consistency and correctness.
+By carefully designing multithreaded applications and using right synchronization techniques, it is possible to minimize the performance impact while still ensuring data consistency and correctness.
 
@@
 
-By using the Raft protocol, Colonies can redirect incoming \emph{assign} requests to the leading Colonies server, thereby ensuring that only one Colonies server replica handles such requests.
+By using the Raft protocol, Colonies can redirect incoming \emph{assign} requests to the leading Colonies server, thereby ensuring that only one Colonies server replica handles assign requests.
 A new leader is automatically elected in the event that a Colonies server replica fails.
@@
 In addition, the Raft protocol makes it possible to seamlessly update Colonies software by making it possible to upgrade each Colonies server replica individually.
-Consequently, making Colonies well-suited for Kubernetes deployments.
+Consequently, making Colonies servers well-suited for Kubernetes deployments.
 
@@
 When a process is closed, the Colonies server checks its child processes to see if all their parent processes have also completed.
-If all parents of a child process have finished, the server updates the process table in the database by setting the flag \emph{wait for parents} to false, which allows child processes to be assigned to executors.
+If all parents of a child process have finished, the server updates the process table in the database by setting the flag \emph{wait for parents} to \emph{false}, which allows child processes to be assigned to executors.
 This operation is also done statelessly when processing the \emph{close} request.
 
-Figure \ref{fig:workflowexec} portrays an execution timeline for a workflow.
+Figure \ref{fig:workflowexec} shows an execution timeline for a workflow.
 Upon submission, process \(P_{1}\) is assigned to Executor \(E_{1}\).
@@
 The table is updated when a workflow is submitted to a Colonies server.
-As mentioned before, when an executor is assigned a process it obtains exclusive rights to that process within the process table.
+As mentioned before, when an executor is assigned a process, it obtains exclusive rights to that process within the process table.
 This exclusivity also means that it is possible for an exector to dynamically add new children processes to the workflow DAG on the fly.
@@
 \subsubsection{Cron}
-The Colonies framework is based on a stateless architecture where each request is treated as an independent transaction and is processed without any information from prior requests.
+The Colonies framework is based on a stateless architecture where each request is treated as an independent transaction which is processed without any information from prior requests.
 Since there is no session data stored in memory, Colonies can easily switch between servers without disruptions or data losses.
@@
 It can for example be used to automate tasks such as routinely fetching data from external systems.
-To make Colonies robust and easy to implement it is essential that Cron workflows also are managed stateless.
+To make Colonies robust and easy to implement, it is essential that Cron workflows also are managed stateless.
 This can be achieved by introducing an additional database table to store information about Cron workflow, and let the elected Colonies server leader assume the responsibility of managing Cron workflows through a two-step process.
@@
 Note that this protocol is completely stateless as no session information is stored in memory between cron scans.
-In the event that the leader Colonies server experiences a crash, a new leader is simply selected and assumes the responsibility of managing Cron workflows according to the two-step protocol.
+In the event that the leader Colonies server experiences a crash, a new leader is selected and assumes the responsibility of managing Cron workflows according to the two-step protocol.
 
@@
 Colonies guarantees that all pack requests submitted to a generator are processed correctly, even if a Colonies server is restarted or fails.
-Without generators, third-party integration software would have to use an alternative method such as a message broker or a database to enqueue input data and periodically empty the queues to generate workflows themselves.
+Without this robustness guarantee, third-party integration software would have to use an alternative method such as a message broker or a database to enqueue input data and periodically empty the queues to generate workflows themselves.
 This adds complexity to the integration process and may require additional infrastructure components.
@@
 This ensures that the generation of workflows is not impacted by server failures, and the system can continue to operate smoothly even in case of failures.
-It can be worth pointing out that since the pack request only adds information to the generator table and does not manipulate any states, it can be processed by any Colonies server without synchronization precautious.
-This means that any Colonies server in the cluster can handle incoming pack requests, which increases the scalability and fault-tolerance of the system as a whole.
+It can be worth pointing out that since the \emph{pack} request only adds information to the generator table and does not manipulate any states, it can be processed by any Colonies server without synchronization precautious.
+This means that any Colonies server in the cluster can handle incoming \emph{pack} requests, which increases the scalability and fault-tolerance of the system as a whole.
 
@@
 The public key is openly available and uploaded to the Colonies server, whereas the private key is kept secret and is only used by the executor to sign API request messages.
-The Colony server can then verify the signature of incoming request messages and use the public keys to determine whether an executor is a member of the colony it is attempting to interact with.
+The Colony server can then verify the signature of incoming request messages and use the public keys to determine whether an executor is a member of a certain colony.
 
@@
 An advantage of using a stateless architecture for workflow management as proposed in the paper is the ability to monitor executor performance to detect any unusual or suspicious behaviors.
-Because all transactions are recorded in a database, it is possible to track the activity of individual executors and identify any unusual patterns or behaviors.
-For example, if an executor suddenly starts to consume a lot of resources, it could indicate a performance issue, or it could be a sign of malicious activity, such as a denial-of-service attack.
+Because the execution history is recorded in a database, it is possible to track the activity of individual executors and identify any unusual patterns.
+For example, if an executor suddenly starts to consume a lot of resources, it could indicate a performance issue or it could be a sign of malicious activity, such as a denial-of-service attack.
 Automatic monitoring can for example be implemented using anomaly detection algorithms and machine learning to grade the performance of each executor.
@@
 Furthermore, there are several software development kits (SDKs) available for various programming languages, including Golang\footnote{https://github.com/colonyos/colonies/tree/main/pkg/client}, Rust\footnote{https://github.com/colonyos/rust}, Julia\footnote{https://github.com/colonyos/Colonies.jl}, JavaScript\footnote{https://github.com/colonyos/colonies.js}, Python\footnote{https://github.com/colonyos/pycolonies}, and Haskell\footnote{https://github.com/colonyos/haskell}.
-As an introduction to Colonies, the following section gives a brief overview of the Colonies tools and how to use the Python SDK.
+As an introduction to Colonies, the following section gives a brief overview how to use the Python SDK.
 
@@
 The first step in implementing an executor in Python is to create a private key and identity, followed by registering the executor with a colony.
-Registering an executor can only be carried out by the colony owner with access to the colony private key.
-Code in Listening \ref{code:regexecutor} illustrates how to register an executor and which functions a specific executor is capable of running.
+As mentioned before, executor can registration can only be carried out by the colony owner using the colony private key.
+Code in Listening \ref{code:regexecutor} illustrates how to register an executor, including registering functions a specific executor is capable of running.
 In the example, an executor of type \emph{helloworld\_executor} is registered to run a function named \emph{helloworld}.
@@
 After registration, the executor must connect to the Colonies server and request process assignments of the type \emph{helloworld\_executor}.
-Note that executor properties, such as the executor type, are not explicitly specified, but rather derived implicitly by the Colonies server from the executor identity, which is derived from the signature generated by the executor's private key.
+Note that executor properties, such as the executor type, are not explicitly set in the \emph{assign} request, but rather derived implicitly by the Colonies server from the executor identity, which is derived from the signature generated by the executor's private key.
 This approach ensures that the executor can only be assigned processes that match its registered capabilities.
@@
 
-This \emph{assign} request is always initiated by the executor using an HTTP-based protocol, allowing executors to be deployed behind firewalls.
+This \emph{assign} request is always initiated by the executor, allowing executors to be deployed behind firewalls.
 When receiving the request, the Colonies server hangs the request until a matching process is found or a timer expires.
@@
 
-When assigned to a process, the executor interprets the metadata stored in the process data structure, including the function name and arguments, and performs some kind of computation such as preprocessing some data or training a neural network.
+When assigned to a process, the executor interprets the metadata stored in the process data structure, including the function name and arguments, and performs some kind of computation such as preprocessing data or training a neural network.
 Upon completion, the executor closes the process and sets the output.
@@
 In this model, the cloud service provider dynamically allocates and manages resources such as adding computing power or storage required by an application.
-Function-as-a-Service (FaaS) is a serverless computing model where developers provide a code snippet that should be executed, and the cloud provider takes care of scaling, configuration and deployment of that code snippet.
+Function-as-a-Service (FaaS) is a serverless computing model where developers provide code snippets that should be executed, and the cloud provider takes care of scaling, configuration and deployment of that code snippets.
 FaaS is often used in event-driven architectures where functions are triggered by specific events, thus eliminating the need to keep servers continuously running.
@@
 Colonies offer functionality to call functions in remote executors, a fundamental requirement to implement a FaaS platform.
-Colonies could also provide service discovery and security isolation to prevent non-authorized function calls.
+Colonies also provides mechanisms for service discovery and security isolation to prevent non-authorized function calls.
 In some cases, it could make sense to manually manage a colony by an administrator, and manually registering and deploying executors.
-Alternatively, a software solution could be developed to automate the administration and automatically deploy or undeploy executors.
+Alternatively, a software solution could be developed to automatically deploy or undeploy executors.
 Such an approach would be very similar to a Function-as-a-Service (FaaS) platform.
@@
 The advantage of Colonies for implementing a FaaS platform is that it introduces an extra level of separation and the possibility to develop a loosely coupled system.
-For example, if an executor is missing, the process is simply enqueued in the process table.
+For example, if an executor is missing, the process is simply enqueued until a free executor becomes available.
 A software agent could then monitor pending processes on the Colonies server, predict the system load, and automatically deploy executors to meet the load, resulting in a highly efficient and scalable system.
-In the future, the AI orchestrator developed in the Cognit project could also be used together with Colonies to reduce energy consumption or minimize network latency by efficiently utilizing edge servers.
+In the future, the AI orchestrator developed in the Cognit project could also be used together with Colonies to reduce energy consumption, or minimize network latency by utilizing edge servers.
 
-\subsection{HPC integration}
+\begin{figure}[h]
+	\centering
+    \includegraphics[scale=0.43]{hpc_overlay.png}
+	\caption{A Colonies-based submission system for HPC.}
+	\label{fig:hpc_overlay}
+\end{figure}
+
+\subsection{Edge computing}
+
+\begin{figure}[h]
+	\centering
+    \includegraphics[scale=0.43]{edge.png}
+	\caption{Using Colonies to control drones.}
+	\label{fig:dashboard2}
+\end{figure}
+
+\subsection{High-performance computing}
 \label{hpc_use_case}
@@
 Submitting a job typically requires users to develop job scripts and manually transfer data to the HPC environment before a job can be executed.
-Compared to cloud platforms, Slurm is less mature when it comes to APIs and users have to develop their own automation tools, which can make the usage more time-consuming and manual.
+Compared to cloud platforms, Slurm is less mature when it comes to API and automation support, which can make the usage more time-consuming and manual.
 
-\begin{figure}[h]
-	\centering
-    \includegraphics[scale=0.43]{hpc_workflow.png}
-	\caption{HPC workflow.}
-	\label{fig:hpc_workflow}
-\end{figure}
-
-Colonies could be used to fill the gap between cloud and HPC and create workflows that can automate processing of data in the cloud while also taking advantage of the vaste computational capacity offered by HPC.
+Colonies could be used to fill the gap between cloud and HPC and create workflows that can partly run in cloud while also taking advantage of the vaste computational capacity offered by HPC.
 For example, a company could develop a workflow that automatically copy training data to a HPC center, train a neural network, move the result back to the cloud, and finally re-deploy a Kubernetes service to use the new model.
+With Colonies, users can indirectly control HPC systems by submitting Colonies processes that are subsequently picked up by HPC executors, without requiring direct connection to the HPC system.
+Essentially, using Colonies as a cloud-based submission system for HPC connecting different HPC centers together.
+Figure \ref{fig:hpc_overlay} depicts such a submission system.
 
@@
 	\centering
-    \includegraphics[scale=0.43]{hpc_overlay.png}
-	\caption{A global grid-like master-control submission system for HPC.}
-	\label{fig:hpc_overlay}
+    \includegraphics[scale=0.43]{hpc_workflow.png}
+	\caption{HPC workflow.}
+	\label{fig:hpc_workflow}
 \end{figure}
 
-In general, integrating with various authentication services presents a major challenge when developing HPC workflows.
-For example, the Data Importer may need to access an S3 service to fetch data.
-The Container Importer must authenticate with a Docker registry to be able to pull a Docker container to the HPC system.
-In addition, it is not straightforward to connect to the HPC system.
-In many cases, users have to manually connect using SSH terminals, making it time-consuming and difficult to automate HPC workflows.
-
-With Colonies, users can indirectly control HPC systems by submitting Colonies processes that are subsequently picked up by HPC executors, without requiring direct connection to the HPC system.
-Essentially, using Colonies as a cloud-based submission system for HPC, connecting different HPC centers together in a grid-like overlay network.
-Figure \ref{fig:hpc_overlay} depicts such a submission system.
-
-Additionally, the security model provided by Colonies allows different executors to trust one another, thereby enabling the sharing of confidential information, such as S3 secrets.
-As credentials are stored in the assigned process record, there is no need to store them locally on the HPC system.
-As a result, HPC executors become more straightforward to configure and deploy, further simplifying creation of HPC workflows.
-
-\subsection{OpenEO example}
+\subsection{Earth Observation}
 OpenEO \cite{openeo} is an open-source project that aims at providing a standardized and interoperable solution for accessing, processing, and analyzing Earth observation (EO) data.
@@
 
-Colonies could offer a promising solution for implementing a decentralized and grid-like compute infrastructure, capable of executing Earth Observation (EO) workloads across diverse platforms.
-By encapsulating computational EO tasks into Colonies processes, interoperability among platforms can be further enhanced.
+Colonies can be used to implement a decentralized grid-based compute infrastructure for Earth Observation (EO) workloads across diverse platforms.
+By wrapping computational EO tasks into Colonies processes, interoperability among platforms can be further enhanced.
 For instance, it becomes possible to create a broker-based architecture that enables OpenEO applications to submit job specifications indirectly to a Colonies server.
-These specifications could then be picked up by executors connected to different backend providers, e.g.
-a Data and Information Access Services (DIAS) executor.
-Executors can also be used to leveraged to provision computational resources.
-For example, automatically provision AWS cloud resources or deploy workflows at HPC systems to handle high-demand EO workloads.
+These specifications could then be picked up by executors connected to different backend providers.
+Executors could also be used to automatically provision compute resources.
+For example, automatically setting up Dask\footnote{https://www.dask.org} clusters, or deploy workflows at HPC systems to handle high-demand EO workloads.
 
 \section{Discussion}
-This paper has presented an open-source framework called Colonies, which streamlines execution of computational workloads across platforms.
-Colonies enables creation of a \emph{grid-like distributed virtual computer}, called a \emph{Colony}, which can be seen as a collective computer machine consisting of distributed executor software running on the Internet.
+This paper has presented an open-source framework called Colonies, which is designed to streamline execution of computational workloads across platforms.
+Colonies enable the creation of a \emph{decentralized virtual computer system}, called a \emph{Colony}, which serves as an intermediary layer to indirectly control and integrate with other computer systems.
+Instead of directly connecting different systems togeather, Colonies allows machines or users to publish instructions and then assign the instructions to other machines.
+By chaining these instructions together, it becomes possible to execute workloads that can operate seamlessly across different platforms.
+
 Unlike many other systems where security sometimes is added later on, Colonies builds on the principle: \emph{security first}.
-
 Public key encryption allows administrators and users to maintain strict control over a colony, even when executors are dispersed throughout the Internet.
 Each colony is governed by a private key that serves as a central authority, analogous to \emph{one ring to rule them all}, making it possible to control a colony as if it was a single machine.
-The zero-trust principle, \emph{trust, but verify} enables development of immune system-like functionality where malfunctional executors can be excluded and replaced with fresh new ones, ensuring that the overall performance and reliability of the system is not compromised.
+The zero-trust principle, \emph{trust, but verify} also enables development of immune system-like functionality where malfunctional executors can be excluded and replaced with fresh new ones, ensuring that the overall performance and reliability of the system is not compromised.
 
-A significant portion of the paper has focused on how to implement robustness, which allows software components to fail without compromising the entire system.
-In Colonies, robustness is a unique feature that can be leveraged not only to build stable data pipelines, but also to efficiently handle IT operations such as continuous deployments and software upgrades.
-When used together with Kubernetes, Colonies can significantly simplify software updates simply by killing pods in a live system, and let Kubernetes replace them with new versions, assuming the new pods have been thoroughly tested and are compatible with the old version.
-This approach can be further enhanced by applying chaos engineering and randomly killing executors to ensure that the system is capable of gracefully tolerating failures.
-In theory, by controlling the rate at which executors are killed, it would be possible to dynamically scale the system up or down, somewhat similar to programmed cell death mechanisms used in biological systems \footnote{https://en.wikipedia.org/wiki/Programmed\_cell\_death}.
+A significant portion of the paper has focused on describing how to handle robustness and allow software components to fail without compromising the entire system.
+A critical design decision that makes Colonies robust and scalable is its stateless architecture.
+Colonies also makes sure that workflow execution run to completion, even if some executor failes.
+This makes it possible to implement robust data pipelines, but also to efficiently handle IT operations such as Continuous Deployments (CD) and software upgrades.
+Both security and robustness are crucial factors to consider when developing a workflow management system that is capable of functioning across platforms and handle constantly changing infrastructures.
 
-Security and robustness are crucial factors to consider when developing a workflow management system that is capable of functioning across various platforms and handling constantly changing infrastructures.
-Another important mechanism is the use of queues, which makes it possible to separate responsibilities between different system components and ensure that each component can operate independently.
+It has been mentioned in the paper several times that Colonies is designed to simplify integration with other systems, but what feature in Colonies makes integration easier?
+Microservices is a design principle used by Colonies to simplify integration.
+As executors are designed to be loosly-coupled and can be implemented in any language (supported by Colonies), the whole system becomes easier to develop and maintain.
+For example, different teams can take on responsibility to develop a certain executor, which can then be used as a part of complex workflows running across platforms.
+By designing each executor as a standalone microservice, it becomes easy to scale the system by simply deploying more executors.
+This also opens up an exciting possibility to delegate colony management responsibility to another software to implement a serverless computing platform capable of automatically provisioning and deploying executors whenever they are needed.
 
-The use of queues also enables a split architecture, where executors can be dynamically managed.
-In Colonies, it is possible to submit a workflow even if no executor is available to execute the workload.
-The processes associated with the workflow are simply enqueued until an executor capable of handling the workload is deployed.
-By designing each executor as a standalone microservice, it becomes easy to scale the system by simply deploying more executors.
-However, if too few executors are available, a queue of pending processes will begin to build up.
-This opens up an exciting possibility to delegate colony management responsibility to another software capable of automatically provisioning and deploying executors whenever they are needed, and then also scale down the system to conserve resources and save costs.
+The paper has also presented some potential use cases.
+In all of the use cases, Colonies serves as intermediary layer that breaks down complex heterogenous systems in to smaller, resilient and more scalable services, and then offer a unified interface for the services to interact with each other.
+For example, in HPC use case, Colonies can help to simplify HPC workflow managemewnt by breaking down complex perations into a set of functions that can be executed as part of a workflow.
 
-It has been mentioned in the paper more than once that Colonies is designed to simplify integration with other systems.
-What feature in Colonies is it that makes the integration easier?
-A core idea in the Colonies framework is to package meta information and instructions how to carry out a computation in a database record called a process, and then provide an API to manipulate the process record, e.g.
-assigning it to an executor.
-As Colonies is designed to be stateless and all data is stored in a database, Colonies provides full transparency and third-party can simply lookup the process to retrieve its status.
-As a result, third-party systems do not have to actively monitor submitted jobs, but can instead trust that the data will eventually be processed.
-This approach offloads a significant portion of error handling onto Colonies instead of the third-party software.
-
-Microservices is another design principle of the Colonies framework that simplifies integration.
-As executors are designed to be loosly-coupled and can be implemented in any language (supported by Colonies), the whole system becomes easier to develop and maintain.
-For example, different teams can take on responsibility to develop a certain executor, which can then be used as a part of a complex workflow running across platforms.
-
-As a final remark, it is worth highlighting once again the significant advantage of Colonies as a security solution.
-As a colony is a trusted community, an executor can share secrets with other colony members.
-By storing credentials, such as one-time passwords or authentication keys in process records, executors can just execute process in run-and-forget style and be able to access third-party infrastructures without the need to store and manage credentials locally.
-Additionally, by targeting specific executor types when submitting a function specification, a colony can be further segmented so that only a subset of the executors in a colony can access the credentials, further improving security.
+% The paper has presented several potential use cases where Colonies serves as an intermediary layer that decomposes complex heterogeneous systems into smaller, resilient, and scalable services, and then provide a unified interface for seamless interaction between these services.
+For instance, in the HPC use cases, Colonies can streamline the workflow management process by HPC operations into a collection of functions that can be seamlessly executed as part of a workflow.
 
@@
 \appendix
-\section{Appendix - Colonies operations}
+\section{Appendix - Colonies Operations}
 \begin{table}[h]

### 2023-04-13 paper_2023/paper.tex
second draft

@@
 To address these challenges, the paper presents an open-source framework called Colonies\footnote{https://github.com/colonyos/colonies}, which facilitates execution of computational workloads across a diverse range of platforms.
-Colonies enables development of a \emph{decentralized virtual computer system}, called a \emph{Colony}, which can effortlessly integrate with any third-party application or other workflow systems.
+Colonies enables development of a \emph{decentralized runtime environment}, called a \emph{Colony}, which can effortlessly integrate with any third-party application or other workflow systems.
 
@@
 
-This paper presents a framework called Colonies, specifically designed to implement a \emph{distributed virtual computer system} that can easily integrate with any third-party system.
+The primary objective of the Colonies framework is to offer uninterrupted access to computing resources, without interruptions or limitations, as data flows from one service to another.
+To establish such seamless compute continuums, the paper introduces Colonies, a \emph{decentralized runtime environment} designed to integrate with third-party systems to create a seamlessly integrated computing ecosystem.
 Colonies is built around a loosely-coupled microservices architecture that separates workflow definitions from implementation and deployment.
@@
 
-The advantage of Colonies for implementing a FaaS platform is that it introduces an extra level of separation and the possibility to develop a loosely coupled system.
+The advantage of using Colonies to implement a FaaS platform is that it introduces an extra level of separation and the possibility to develop a loosely coupled system.
 For example, if an executor is missing, the process is simply enqueued until a free executor becomes available.
@@
 
-\begin{figure}[h]
-	\centering
-    \includegraphics[scale=0.43]{hpc_overlay.png}
-	\caption{A Colonies-based submission system for HPC.}
-	\label{fig:hpc_overlay}
-\end{figure}
+\subsection{Edge computing}
+Edge computing is a distributed computing model that brings computing resources closer to the source where data is generated to improve efficiency or reduce latency.
+In edge computing, computing and processing is usually done on the \emph{edge} of a network, such as at a server connected to a 5G base station.
+One could for example imagine a drone video surveillance use case where parts of a video processing pipeline run on an edge server.
+This approach would make it possible to execute complex AI models that would otherwise be difficult to run on a drone with limited computing capabilities.
 
-\subsection{Edge computing}
+Setting up a video pipeline is a complex problem.
+First, the drone must be able to discover nearby edge servers, make sure the edge server has sufficient resources, deploy some software components at the edge server, then start a video server and finally connect to the video server.
+Orchestrating such a workflow requires configuration and provisioning of computational resources across platforms.
 
@@
     \includegraphics[scale=0.43]{edge.png}
-	\caption{Using Colonies to control drones.}
-	\label{fig:dashboard2}
+	\caption{Edge workflows, using Colonies to control drones.}
+	\label{fig:edge}
 \end{figure}
+
+The Colonies framework has the potential to streamline the setup of edge computing workflows by enabling devices and edge servers to be represented as distributed executors that can interact with other.
+Figure \ref{fig:edge} illustrates a setup of three executors.
+The \emph{Edge Executor} is responsible for deploying and setting up video processing software at an edge server, while the \emph{Drone Executor} is responsible for activating the video camera at the drone and establishing a connection to the video server hosted on the edge server.
+The \emph{Video Surveillance Executor} is responsible for identifying suitable edge servers and facilitating submission of workflows by users.
+Discovery of edge executors can be achieved by interacting with a Colonies server and searching the colony for available executors.
+
+Figure \ref{fig:edge} also depicts a Colonies workflow.
+The workflow guarantees that the video pipeline is set up in the right order.
+For example, the drone executor should only connect to the edge server after the video server is running.
+The workflow can also be used to pass information between the executors.
+After the \emph{Edge Executor} has set up a video server, it sends a \emph{close} request containing the IP address to the video server.
+The IP address will then be passed as input to the next process in the workflow, allowing the \emph{Drone Executor} to establish a connection to the edge server.
 
@@
 
+\begin{figure}[h]
+	\centering
+    \includegraphics[scale=0.43]{hpc_overlay.png}
+	\caption{A Colonies-based submission system for HPC.}
+	\label{fig:hpc_overlay}
+\end{figure}
+
 Colonies could be used to fill the gap between cloud and HPC and create workflows that can partly run in cloud while also taking advantage of the vaste computational capacity offered by HPC.
@@
 Several executors run at the HPC system, each with their own responsibilities.
-The Container Importer Executor is responsible for converting container images to a format that is supported by Slurm, e.g.
+The \emph{Container Importer Executor} is responsible for converting container images to a format that is supported by Slurm, e.g.
 Singularity containers\footnote{https://docs.sylabs.io}.
-The Data Importer Executors downloads data to the HPC file system.
-The Slurm Executor is responsible for submitting and overseeing Slurm jobs, while the Data Exporter executor is responsible for uploading result data to the cloud.
+The \emph{Data Importer Executor} downloads data to the HPC file system.
+The \emph{Slurm Executor} is responsible for submitting and overseeing Slurm jobs, while the \emph{Data Exporter} executor is responsible for uploading result data to the cloud.
 
@@
 This paper has presented an open-source framework called Colonies, which is designed to streamline execution of computational workloads across platforms.
-Colonies enable the creation of a \emph{decentralized virtual computer system}, called a \emph{Colony}, which serves as an intermediary layer to indirectly control and integrate with other computer systems.
-Instead of directly connecting different systems togeather, Colonies allows machines or users to publish instructions and then assign the instructions to other machines.
-By chaining these instructions together, it becomes possible to execute workloads that can operate seamlessly across different platforms.
+Colonies enable the creation of a \emph{decentralized runtime environment}, called a \emph{Colony}, which serves as an intermediary layer to indirectly control and integrate with other computer systems.
+Instead of directly connecting different systems together, Colonies allows services or users to publish instructions and then assign these instructions to other services.
+By chaining instructions together, it becomes possible to execute workloads that can operate seamlessly across different platforms.
 
@@
 
-A significant portion of the paper has focused on describing how to handle robustness and allow software components to fail without compromising the entire system.
+A significant portion of the paper has focused on describing how to handle robustness to allow software components to fail without compromising the entire system.
 A critical design decision that makes Colonies robust and scalable is its stateless architecture.
-Colonies also makes sure that workflow execution run to completion, even if some executor failes.
-This makes it possible to implement robust data pipelines, but also to efficiently handle IT operations such as Continuous Deployments (CD) and software upgrades.
-Both security and robustness are crucial factors to consider when developing a workflow management system that is capable of functioning across platforms and handle constantly changing infrastructures.
+Colonies also make sure that workflow execution runs to completion, even if some executors fail.
+This makes it possible to implement robust data pipelines, but also efficiently handle IT operations such as continuous deployments and software updates.
+Both security and robustness are crucial factors to consider when developing a workflow management system that is capable of functioning across platforms and handling constantly changing infrastructures.
 
@@
 Microservices is a design principle used by Colonies to simplify integration.
-As executors are designed to be loosly-coupled and can be implemented in any language (supported by Colonies), the whole system becomes easier to develop and maintain.
+As executors are designed to be loosely-coupled microservices, the entire system becomes easier to develop, test, and maintain.
 For example, different teams can take on responsibility to develop a certain executor, which can then be used as a part of complex workflows running across platforms.
-By designing each executor as a standalone microservice, it becomes easy to scale the system by simply deploying more executors.
-This also opens up an exciting possibility to delegate colony management responsibility to another software to implement a serverless computing platform capable of automatically provisioning and deploying executors whenever they are needed.
+By designing each executor as a standalone microservice, it also becomes easy to scale the system by simply deploying more executors.
+This also opens up for an intriguing possibility to delegate colony management responsibility to a software agent that can automatically provision and deploy executors on demand.
 
-The paper has also presented some potential use cases.
-In all of the use cases, Colonies serves as intermediary layer that breaks down complex heterogenous systems in to smaller, resilient and more scalable services, and then offer a unified interface for the services to interact with each other.
-For example, in HPC use case, Colonies can help to simplify HPC workflow managemewnt by breaking down complex perations into a set of functions that can be executed as part of a workflow.
+While microservices is a powerful design pattern, Colonies goes beyond the conventional microservice model and offers a decentralized microservice architecture that allows microservices to reside anywhere on the Internet.
+Also, by acting as an intermediary layer, Colonies provides a unified interface to facilitate interaction between these microservices, making it possible to create workflows that span multiple platforms.
+For instance, Colonies could be used to expose an entire HPC system as a microservice and then connected it to other microservices running in the cloud or IoT devices.
 
-% The paper has presented several potential use cases where Colonies serves as an intermediary layer that decomposes complex heterogeneous systems into smaller, resilient, and scalable services, and then provide a unified interface for seamless interaction between these services.
-For instance, in the HPC use cases, Colonies can streamline the workflow management process by HPC operations into a collection of functions that can be seamlessly executed as part of a workflow.
+In summary, the primary objective of Colonies is to offer uninterrupted access to computing resources, without interruptions or limitations, as data flows from one service to another.
+Achieving seamless compute continuum requires development of simple, secure, scalable, and robust solutions.
+Colonies serves as an initial step towards realizing this vision of a seamlessly integrated computing ecosystem.
 

### 2023-04-13 paper_2023/paper.tex
second draft2

@@
 
-The primary objective of the Colonies framework is to offer uninterrupted access to computing resources, without interruptions or limitations, as data flows from one service to another.
-To establish such seamless compute continuums, the paper introduces Colonies, a \emph{decentralized runtime environment} designed to integrate with third-party systems to create a seamlessly integrated computing ecosystem.
+The primary objective of the Colonies framework is to offer uninterrupted access to computing resources, without interruptions or limitations, allowing data to flow from one service to another.
+To establish such seamless compute continuums, the paper introduces Colonies, a \emph{decentralized runtime environment} designed to easily integrate with third-party systems to create a seamlessly integrated computing ecosystem.
 Colonies is built around a loosely-coupled microservices architecture that separates workflow definitions from implementation and deployment.
@@
 
-In summary, the primary objective of Colonies is to offer uninterrupted access to computing resources, without interruptions or limitations, as data flows from one service to another.
+In summary, the primary objective of Colonies is to offer uninterrupted access to computing resources, allowing data to flow from one service to another.
 Achieving seamless compute continuum requires development of simple, secure, scalable, and robust solutions.
-Colonies serves as an initial step towards realizing this vision of a seamlessly integrated computing ecosystem.
+Colonies serves as an initial step towards realizing the vision of a seamlessly integrated computing ecosystem.
 

### 2023-04-13 paper_2023/paper.tex
second draft3

@@
 Colonies is currently used commerically by a startup company called RockSigma AB\footnote{https://www.rocksigma.com} to implement a compute engine for analysing seismic activities in underground mines.
-This section will not dig into the details of this specific implementation; instead it will provide an outline of potential future use cases in which Colonies could make a considerable impact.
+The paper will not explain the details of this specific implementation; instead it will outline other potential use cases where Colonies could make a considerable impact.
 

### 2023-04-14 paper_2023/paper.tex
wip

@@
 
-The primary objective of the Colonies framework is to offer uninterrupted access to computing resources, without interruptions or limitations, allowing data to flow from one service to another.
+The long-term vision of the Colonies framework is to offer uninterrupted access to computing resources, without interruptions or limitations, allowing data to flow from one service to another.
 To establish such seamless compute continuums, the paper introduces Colonies, a \emph{decentralized runtime environment} designed to easily integrate with third-party systems to create a seamlessly integrated computing ecosystem.
@@
 
-\subsection{Earth Observation}
-OpenEO \cite{openeo} is an open-source project that aims at providing a standardized and interoperable solution for accessing, processing, and analyzing Earth observation (EO) data.
+% \subsection{Earth Observation}
+% OpenEO \cite{openeo} is an open-source project that aims at providing a standardized and interoperable solution for accessing, processing, and analyzing Earth observation (EO) data.
 Its primary objective is to facilitate seamless user interaction with diverse data sources and processing backends through a unified API.

### 2023-04-15 paper_2023/paper.tex
wip

@@
 To address these challenges, the paper presents an open-source framework called Colonies\footnote{https://github.com/colonyos/colonies}, which facilitates execution of computational workloads across a diverse range of platforms.
-Colonies enables development of a \emph{decentralized runtime environment}, called a \emph{Colony}, which can effortlessly integrate with any third-party application or other workflow systems.
+Colonies enables development of a \emph{distributed runtime environment}, called a \emph{Colony}, which can effortlessly integrate with any third-party application or other workflow systems.
 
@@
 
-The long-term vision of the Colonies framework is to offer uninterrupted access to computing resources, without interruptions or limitations, allowing data to flow from one service to another.
-To establish such seamless compute continuums, the paper introduces Colonies, a \emph{decentralized runtime environment} designed to easily integrate with third-party systems to create a seamlessly integrated computing ecosystem.
-Colonies is built around a loosely-coupled microservices architecture that separates workflow definitions from implementation and deployment.
-The primary objective is to establish a platform where workflows can be decomposed into independent microservices that can easily be integrated with other systems or applications running anywhere on the Internet.
-The remainder of the paper describes the Colonies framework and how it can be used to create robust and scalable AI systems across platforms.
+The long-term vision of the Colonies framework is to enable uninterrupted access to computing resources, allowing data to flow from one service to another establishing a seamlessly integrated computing ecosystem.
+To realize this vision, the paper presents Colonies, a \emph{distributed runtime environment} designed to easily integrate with third-party systems.
+The remainder of the paper describes the Colonies framework and how it can be used to create robust and scalable compute continuums.
 
@@
 
-Colonies can be used to implement a decentralized grid-based compute infrastructure for Earth Observation (EO) workloads across diverse platforms.
+% Colonies can be used to implement a decentralized grid-based compute infrastructure for Earth Observation (EO) workloads across diverse platforms.
 By wrapping computational EO tasks into Colonies processes, interoperability among platforms can be further enhanced.
@@
 This paper has presented an open-source framework called Colonies, which is designed to streamline execution of computational workloads across platforms.
-Colonies enable the creation of a \emph{decentralized runtime environment}, called a \emph{Colony}, which serves as an intermediary layer to indirectly control and integrate with other computer systems.
+Colonies enable the creation of a \emph{distributed runtime environment}, called a \emph{Colony}, which serves as an intermediary layer to indirectly control and integrate with other computer systems.
 Instead of directly connecting different systems together, Colonies allows services or users to publish instructions and then assign these instructions to other services.
@@
 
-While microservices is a powerful design pattern, Colonies goes beyond the conventional microservice model and offers a decentralized microservice architecture that allows microservices to reside anywhere on the Internet.
+While microservices is a powerful design pattern, Colonies goes beyond the conventional microservice model and offers a distributed microservice architecture that allows microservices to reside anywhere on the Internet.
 Also, by acting as an intermediary layer, Colonies provides a unified interface to facilitate interaction between these microservices, making it possible to create workflows that span multiple platforms.

### 2023-04-17 paper_2023/paper.tex
wip

@@
 To address these challenges, the paper presents an open-source framework called Colonies\footnote{https://github.com/colonyos/colonies}, which facilitates execution of computational workloads across a diverse range of platforms.
-Colonies enables development of a \emph{distributed runtime environment}, called a \emph{Colony}, which can effortlessly integrate with any third-party application or other workflow systems.
+Colonies provides a \emph{distributed runtime environment}, called a \emph{Colony}, which can effortlessly integrate with any third-party application or other workflow systems.
 
 Based on a stateless microservice architecture, complex workflows can be broken down into composable functions.
-These composable functions can be implemented in any programming language and be executed by distributed executors deployed across platforms, anywhere on the Internet such as IoT, cloud, edge, and supercomputers.
+These composable functions can be implemented in any programming language and be executed by distributed executors deployed across platforms, anywhere on the Internet such as IoT, cloud, edge, or a supercomputer.
 A zero-trust security protocol enables a collection of distributed executors to operate as a unified entity, establishing seamless compute continuums across platforms.
@@
 The long-term vision of the Colonies framework is to enable uninterrupted access to computing resources, allowing data to flow from one service to another establishing a seamlessly integrated computing ecosystem.
-To realize this vision, the paper presents Colonies, a \emph{distributed runtime environment} designed to easily integrate with third-party systems.
+To realize this vision, the paper presents \emph{Colonies}, a \emph{distributed runtime environment}.
 The remainder of the paper describes the Colonies framework and how it can be used to create robust and scalable compute continuums.
@@
 The \emph{Colonies servers} form the backbone of the framework, functioning as a centralized control system for task submission and assignment.
-The Colonies server acts as a job broker for executors, almost like an employment agency for people.
+The Colonies server serves as a job broker for executors, almost like an employment agency for people.
 All execution information and history are stored in a database maintained by the Colonies servers.
@@
 This paper has presented an open-source framework called Colonies, which is designed to streamline execution of computational workloads across platforms.
-Colonies enable the creation of a \emph{distributed runtime environment}, called a \emph{Colony}, which serves as an intermediary layer to indirectly control and integrate with other computer systems.
+Colonies provides a \emph{distributed runtime environment}, called a \emph{Colony}, which serves as an intermediary layer to indirectly control and integrate with other computer systems.
 Instead of directly connecting different systems together, Colonies allows services or users to publish instructions and then assign these instructions to other services.
@@
 
-While microservices is a powerful design pattern, Colonies goes beyond the conventional microservice model and offers a distributed microservice architecture that allows microservices to reside anywhere on the Internet.
-Also, by acting as an intermediary layer, Colonies provides a unified interface to facilitate interaction between these microservices, making it possible to create workflows that span multiple platforms.
-For instance, Colonies could be used to expose an entire HPC system as a microservice and then connected it to other microservices running in the cloud or IoT devices.
+While microservices is a powerful design pattern, Colonies goes beyond the conventional microservice model and offers a distributed microservice architecture allowing microservices to reside anywhere on the Internet.
+Also, by acting as an intermediary layer, Colonies provides a unified interface to facilitate interaction between these microservices, ultimately making it possible to create workflows that span multiple platforms.
+The paper has also presented into several use cases related to edge computing, cloud computing and high-performance computing, highlighting the benefits that such a framework could provide, in particular how to create loosely-coupled systems that offers both flexibility, robustness, and scalability.
 
-In summary, the primary objective of Colonies is to offer uninterrupted access to computing resources, allowing data to flow from one service to another.
+To sum up, the long-term vision of the Colonies framework is to offer uninterrupted access to computing resources, allowing data to flow from one service to another.
 Achieving seamless compute continuum requires development of simple, secure, scalable, and robust solutions.
-Colonies serves as an initial step towards realizing the vision of a seamlessly integrated computing ecosystem.
+Colonies serves as an initial step towards realizing this vision of a seamlessly integrated computing ecosystem.
 

### 2023-05-02 paper_2023/paper.tex
wip

@@
 \section{Discussion}
-This paper has presented an open-source framework called Colonies, which is designed to streamline execution of computational workloads across platforms.
-Colonies provides a \emph{distributed runtime environment}, called a \emph{Colony}, which serves as an intermediary layer to indirectly control and integrate with other computer systems.
+This paper has presented an open-source framework called Colonies designed to streamline execution of computational workloads across platforms.
+Colonies provides a \emph{distributed runtime environment}, called a \emph{Colony}, which serves as an intermediary layer to indirectly control other computer systems.
 Instead of directly connecting different systems together, Colonies allows services or users to publish instructions and then assign these instructions to other services.
@@
 
-Unlike many other systems where security sometimes is added later on, Colonies builds on the principle: \emph{security first}.
+Colonies builds on the principle: \emph{security first}.
 Public key encryption allows administrators and users to maintain strict control over a colony, even when executors are dispersed throughout the Internet.
@@
 
-It has been mentioned in the paper several times that Colonies is designed to simplify integration with other systems, but what feature in Colonies makes integration easier?
+It has been mentioned in the paper several times that Colonies is designed to simplify integration with other systems.
 Microservices is a design principle used by Colonies to simplify integration.

### 2023-05-03 paper_2023/paper.tex
wip

@@
 At the same time, development and operation of AI workloads has become increasingly challenging.
-One difficulty is the lack of portability, making it cumbersome to move from one platform to another.
+One difficulty is lack of portability, making it cumbersome to move from one platform to another.
 Creating and operating fully automated end-to-end workflows across devices, edge, and cloud platforms is even more challenging.
@@
 Based on a stateless microservice architecture, complex workflows can be broken down into composable functions.
-These composable functions can be implemented in any programming language and be executed by distributed executors deployed across platforms, anywhere on the Internet such as IoT, cloud, edge, or a supercomputer.
+These composable functions can be implemented in any programming language and be executed by distributed executors deployed anywhere on the Internet such as IoT, cloud, edge, or at a supercomputer.
 A zero-trust security protocol enables a collection of distributed executors to operate as a unified entity, establishing seamless compute continuums across platforms.
@@
 In many cases, there is a constant flow of data ingested into the system which must be managed even in case of failures.
-This becomes even more challenging when nodes or parts of the underlying infrastructure become unavailable due to maintenance such as software updates, hardware replacements and sometimes misconfiguration problems.
+This becomes even more challenging when nodes or parts of the underlying infrastructure become unavailable due to maintenance such as software updates, hardware replacements or sometimes misconfiguration problems.
 
@@
 Scaling the system means that the underlying infrastructure may change at any time, causing instability issues for running services or workflows.
-Therefore, it must be possible to detect failed computations and reprocess failed tasks part of a larger workflow.
+Therefore, it must be possible to detect failed computations and re-process failed tasks part of a workflow.
 Workflows must hence be designed to handle an ever-changing infrastructure, and if a failed computation cannot be restored gracefully, engineers must be able to quickly perform root cause analysis to manually recover the system.
@@
 Additionally, many compute clusters operate on-premises installations.
-Sometimes it is necessary to temporarily increase the capacity of on-prem clusters by combining resources from multiple providers, for example, adding cloud compute resources to handle peak loads or utilize HPC resources to quickly reprocess historical data.
+Sometimes it is necessary to temporarily increase the capacity of on-prem clusters by combining resources from multiple providers, for example, adding cloud compute resources to handle peak loads or utilize HPC resources to quickly re-process historical data.
 Developing hybrid workflows where some jobs run in the cloud and others run on HPC systems requires even more software development efforts \cite{wf_challenges} and is beyond the scope of many users, preventing them from utilizing powerful hardware.
@@
 The long-term vision of the Colonies framework is to enable uninterrupted access to computing resources, allowing data to flow from one service to another establishing a seamlessly integrated computing ecosystem.
-To realize this vision, the paper presents \emph{Colonies}, a \emph{distributed runtime environment}.
-The remainder of the paper describes the Colonies framework and how it can be used to create robust and scalable compute continuums.
+To realize this vision, the paper presents \emph{Colonies}, a \emph{distributed runtime environment} enabling compute continuums.
+The remainder of the paper describes the Colonies framework and how it can be used to create robust and scalable AI workflows.
 
@@
 Attempts are currently being made to implement serverless workflow management systems.
-For example, OpenWolf \cite{openwolf} is a serverless workflow engine designed to utilize the Function-as-a-Service (FaaS) paradigm for composing complex scientific workflows.
+For example, OpenWolf \cite{openwolf} is a serverless workflow engine designed to utilize the Function-as-a-Service (FaaS) paradigm to compose complex scientific workflows.
 It is based on OpenFaaS \cite{openfaas}, which allows functions to run on Kubernetes clusters.
@@
 This will be further discussed in Section \ref{faas}.
-However, in contrast to previous work, Colonies is designed to be platform independent and does therefore not require Kubernetes.
+However, in contrast to previous work, Colonies is designed to be platform independent and does hence not require container orchestration platforms like Kubernetes.
 By using a zero-trust security protocol, functions can be securely executed by distributed executors deployed anywhere on the Internet.
-It is important to point out that Colonies does not provide an infrastructure for function execution.
-Instead, Colonies primary role is to serve as a platform for coordinating function executions which are carried out by distributed executors.
 
@@
 The vision is to create a distributed architecture where microservices can reside anywhere on the Internet and still function as a cohesive unit to execute cross-platform workloads.
-o
 
@@
 A core concept of the Colonies framework is the notion of \emph{processes}.
-A process contains meta-information about computations executed by remote computer programs, referred to as \emph{distributed executors}.
-Specifically, a process contains a definition of a function to be executed as well as contextual information such as execution status, including the result of the computation.
+A process contains meta-information about computations executed by remote computer programs, referred to as \emph{executors}.
+Specifically, a process contains a definition of a function to be executed as well as contextual information such as execution status and the result of the computation.
 One can think of a process as a digital twin of a real computing process.
-It is essential to note that a process does not necessarily have to be an operating system process, but can represent any type of computation, e.g.
+It is essential to point out that a process does not necessarily have to be an operating system process, but can represent any type of computation, e.g.
 a remote procedure call executed by any kind of software.
@@
 
-\emph{Distributed executors} are responsible for executing processes assigned by the Colonies Servers.
-Several distributed executors form a \emph{Colony}, which is a collection of executors operating as a cohesive unit where each executor is responsible for executing specific types of processes.
-One could view a Colony as an organization or a society of distributed computer softwares acting as a single virtual computer system.
+\emph{Executors} are responsible for executing processes assigned by the Colonies Servers.
+Several executors form a \emph{Colony}, which is a collection of executors operating as a cohesive unit where each executor is responsible for executing specific types of processes.
+One could view a Colony as an organization (like a company) of distributed computer softwares acting as a single virtual computer system.
 To interact with other executors, executors must prove their Colony membership using a cryptographic protocol that follows the zero-trust security principle: \emph{never trust, always verify}.
@@
 \item \textbf{Single responsibility:} Each executor is only accountable for executing specific functions. This makes the system easier to understand, develop, test and maintain.  
-\item \textbf{Loosely coupled:} Executors are designed with minimal dependencies on other executors, enabling various software development teams to work independently.
+\item \textbf{Loosely coupled:} Executors are designed with minimal dependencies on other executors, enabling different software development teams to work independently.
 For instance, a data engineering team may handle the implementation of pre-processing functions, while a data science team oversees machine learning functions, and another team manages visualization or customer integration etc.
 
-\item \textbf{Scalability:} Executors can be deployed independently, which enables horizontal scaling. This allows for better resource utilization, parallelism, and improved performance.
+\item \textbf{Scalability:} Executors can be deployed independently and be scaled horizontally. This enables efficient resource utilization, parallelism, and improved performance.
 
@@
 Executors' isolation from one another contributes to a more resilient and fault-tolerant system.
-If an executor crashes during execution, the process is automatically reassigned to another executor.
+If an executor crashes during execution, the process is automatically re-assigned to another executor.
 
@@
 
-\item \textbf{Technology agnostic:} Executors can be implemented in any programming language, facilitating seamless integration with other applications and systems.
-
-\item \textbf{Decentralized governance:} As the Colonies framework is technology-agnostic, different software development teams can make independent technology and design decisions when developing executors, promoting greater flexibility and adaptability.
+\item \textbf{Decentralized governance:} As executors are technology-agnostic, different software development teams can make independent technology and design decisions when developing new executors, promoting greater flexibility and adaptability.
 For example, some executors may be implemented in Rust, while others may use Python to leverage state-of-the-art machine learning frameworks.
@@
 
-Although microservices have attractive properties and can simplify both development and deployment of executors, they also introduce additional complexity that must be managed by the Colonies framework.
+Although microservices have attractive software engineering properties, they also introduce additional complexity that must be managed by the Colonies framework.
 To implement a workflow management framework supporting distributed microservices, the following challenges must be addressed:
@@
 
+\item \textbf{Load balancing:} The Colonies framework must manage load balancing among executors to optimize the overall performance.
+
 \item \textbf{Fault tolerance and recovery:} In the event of executor failures, the Colonies framework must be able to re-assign processes to other executors, manage executor restarts, or trigger recovery mechanisms to maintain system reliability and resilience.
-In addition, the Colonies framework must be able to handle restarts or crashes of Colonies servers, thereby preventing any internal states in Colonies from becoming corrupted.
-
-\item \textbf{Load balancing:} The Colonies framework must manage load balancing among executors to optimize performance and avoid overloading any single executor.
+In addition, the Colonies framework must be able to handle re-starts or crashes of Colonies servers and prevent any internal states from becoming corrupted.
 
@@
 \item \textbf{Workflow orchestration:} The Colonies framework must coordinate and orchestrate complex multi-step workflows executed by several executors, sometimes in parallel where executors run on different platforms.
-Colonies must define the sequence in which processes should be executed, manage dependencies among processes, and ensure that workflows execute successfully to completion.
+This includes defining the sequence in which processes should be executed, manage dependencies among processes, and ensure that workflows execute successfully to completion.
 
@@
 
-The Colonies framework uses a combination of technologies to address the aforementioned challenges.
+As will be discussed, the Colonies framework uses a combination of technologies to address the aforementioned challenges.
 A fundamental design principle is statelessness, which makes Colonies simpler, more reliable and scalable, and easier to implement.
@@
 The remainder of this section describes how the Colonies framework is implemented.
-The next discusses the role of queues to support batch jobs and realtime processing, but also queues roles in implementing loosely coupled systems.
+The next discusses the role of queues to support batch jobs and realtime processing, but also discusses queues roles in implementing loosely coupled systems.
 
@@
 For example, HTTP APIs can be used to abstract away implementation detail and provide a clear and simple interface for interacting with a particular service.
-However, HTTP APIs alone are insufficient for handling dynamic environments where components frequently fail or the underlying infrastructure is constantly changing.
+However, APIs alone are insufficient to handle dynamic environments where components frequently fail or the underlying infrastructure is constantly changing.
 To address such environments, an alternative mechanism is needed.
@@
 Queues can be implemented in different ways.
-While message brokers are a common solution, Colonies adopt an alternative strategy and leverage a standard database and query it for processes to assign to different executors.
+While message brokers are a common solution, Colonies adopt an alternative strategy and is based on a standard database.
 One key advantage of this approach is that it enables fine-grained process assignments, making it possible to assign specific processes to particular executors.
-For instance, an executor of the browser type can be limited to only executing processes capable of running in web browsers.
+For instance, an executor can be limited to only execute processes capable of running in web browsers.
 This level of granularity cannot easily be implemented using message brokers which generally do not offer introspection of queues, or provide the ability to pull specific messages out of the queue.
 Generally, the only way to retrieve a specific message is to pull all messages from the queue, obtain the message, and then place all remaining messages back into the queue in the same order.
-In contrast, a database can function as a queue and a query can match any columns making it possible to assign specific executors to specific processes.
+In contrast, a database can function as a queue and a query can match any columns making it possible to assign an executors to specific processes.
 
@@
 Rather, it is the responsibility of the executors to connect to the Colonies server.
-This enables executors to be deployed anywhere on the Internet, behind firewalls, commercial telco networks, or even in web browser-based applications.
+This enables executors to be deployed anywhere on the Internet, behind firewalls, commercial telco networks, or even in web browsers.
 
@@
 When registering, executors have to specify to the Colonies server which functions they are capable of executing.
-The Colonies server then makes sure that the conditions part of the function specification matches the capability of an executor.
+The Colonies server then makes sure that function specifications match the capability of executors.
 
@@
 It is worth noting that synchronization is not necessary for other requests.
-For example, as the \emph{submit} request only adds new entries to the process table, there are no race conditions.
+For example, as the \emph{submit} request only adds new entries to the process table, hence there are no race conditions.
 The \emph{close} request sets the output of the function innovation and updates the process state to either successful or failed in the process table.
@@
 
-By using the Raft protocol, Colonies can redirect incoming \emph{assign} requests to the leading Colonies server, thereby ensuring that only one Colonies server replica handles assign requests.
+By using the Raft protocol, Colonies can redirect incoming \emph{assign} requests to the Colonies server replica elected as the leader, thereby ensuring that only one Colonies server replica handles assign requests.
 A new leader is automatically elected in the event that a Colonies server replica fails.
 Figure \ref{fig:ha_deployment} shows an overview of a high-availability Colonies deployment.
-In addition, the Raft protocol makes it possible to seamlessly update Colonies software by making it possible to upgrade each Colonies server replica individually.
-Consequently, making Colonies servers well-suited for Kubernetes deployments.
 
@@
 In Colonies, workflows are represented as directed acyclic graphs (DAGs) where nodes represent processes and edges represent dependencies and data flow between processes.
-Similar to processes, workflows are managed completely stateless.
+Like processes, workflows are managed completely stateless.
 When a workflow is submitted, all processes part of the workflow are submitted and added to the process table similar to how ordinary processes are handled.
-To control the order processes can be assigned, Colonies sets a flag, \emph{wait for parents} to prevent processes to be assigned to an executor before their dependencies have been satisfied.
+To control the order processes are assigned, Colonies sets a flag, \emph{wait for parents} to prevent processes to be assigned to an executor before their dependencies have been satisfied.
 Note that processes may run in parallel if a sufficient number of executors are available, assuming their parent processes have finished.
@@
 
-To manage data flow between processes that are part of a workflow, it is necessary to store the input and output values resulting from function invocations in a database.
+To manage data flow between processes that are part of a workflow, it is necessary to store the input and output values from function executions in the database.
 An example of such a database is presented in Table \ref{inouttable}.
@@
 It can for example be used to automate tasks such as routinely fetching data from external systems.
-To make Colonies robust and easy to implement, it is essential that Cron workflows also are managed stateless.
-This can be achieved by introducing an additional database table to store information about Cron workflow, and let the elected Colonies server leader assume the responsibility of managing Cron workflows through a two-step process.
+To make Colonies robust and easy to implement, it is essential that Cron workflows also are managed statelessly.
+This can be achieved by introducing an additional database table to store information about Cron workflow, and let the elected Colonies server replica leader assume the responsibility of managing Cron workflows through a two-step process.
 
-Firstly, the leading Colonies server calculates a future deadline, indicating when a specific Cron workflow is scheduled to be executed.
+Firstly, the Colonies server replica leader calculates a future deadline, indicating when a specific Cron workflow is scheduled to be executed.
 The leader server then conducts periodic scans of the Cron table.
@@
 Note that this protocol is completely stateless as no session information is stored in memory between cron scans.
-In the event that the leader Colonies server experiences a crash, a new leader is selected and assumes the responsibility of managing Cron workflows according to the two-step protocol.
+In the event that the leader Colonies server crashes, a new leader is selected and assumes the responsibility of managing Cron workflows according to this two-step protocol.
 
@@
 
-Generators can be implemented using a similar method to how Cron workflows are implemented.
+Generators is implemented using a similar method like how Cron workflows are implemented.
 The elected leader Colonies server takes on the responsibility for scanning a special generator table to decide if a workflow should be triggered.
@@
 
-This \emph{assign} request is always initiated by the executor, allowing executors to be deployed behind firewalls.
+This \emph{assign} request is always initiated by the executors, allowing executors to be deployed behind firewalls.
 When receiving the request, the Colonies server hangs the request until a matching process is found or a timer expires.
@@
 In this model, the cloud service provider dynamically allocates and manages resources such as adding computing power or storage required by an application.
-Function-as-a-Service (FaaS) is a serverless computing model where developers provide code snippets that should be executed, and the cloud provider takes care of scaling, configuration and deployment of that code snippets.
+Function-as-a-Service (FaaS) is a serverless computing model where developers provide code snippets that should be executed, and the cloud provider takes care of scaling and deployment of that code snippets.
 FaaS is often used in event-driven architectures where functions are triggered by specific events, thus eliminating the need to keep servers continuously running.
@@
 
-Colonies offer functionality to call functions in remote executors, a fundamental requirement to implement a FaaS platform.
+Colonies offer functionality to call remote functions, a fundamental requirement to implement a FaaS platform.
 Colonies also provides mechanisms for service discovery and security isolation to prevent non-authorized function calls.
-In some cases, it could make sense to manually manage a colony by an administrator, and manually registering and deploying executors.
-Alternatively, a software solution could be developed to automatically deploy or undeploy executors.
-Such an approach would be very similar to a Function-as-a-Service (FaaS) platform.
+A software solution could be developed to automatically deploy and undeploy executors.
+Such an approach would be very similar to a Function-as-a-Service (FaaS) solution.
 
 Figure \ref{fig:faas} depicts a Colonies-based FaaS platform.
-Similar to AWS Lambda or Google Cloud Functions, developers would need to upload code snippets, which would then be packaged and published as Docker containers.
+Similar to AWS Lambda or Google Cloud Functions, developers need to upload executors source code, which would then be automatically packaged and published as Docker containers.
 To invoke a function, a function specification needs to be submitted to the Colonies server indicating which function to call.
@@
 The advantage of using Colonies to implement a FaaS platform is that it introduces an extra level of separation and the possibility to develop a loosely coupled system.
-For example, if an executor is missing, the process is simply enqueued until a free executor becomes available.
+For example, if an executor is missing, the process is simply enqueued until a auitable executor becomes available.
 A software agent could then monitor pending processes on the Colonies server, predict the system load, and automatically deploy executors to meet the load, resulting in a highly efficient and scalable system.
-In the future, the AI orchestrator developed in the Cognit project could also be used together with Colonies to reduce energy consumption, or minimize network latency by utilizing edge servers.
+In the future, the AI orchestrator developed in the Cognit project could also be used together with Colonies to reduce energy consumption or minimize network latency.
 
@@
 Edge computing is a distributed computing model that brings computing resources closer to the source where data is generated to improve efficiency or reduce latency.
-In edge computing, computing and processing is usually done on the \emph{edge} of a network, such as at a server connected to a 5G base station.
-One could for example imagine a drone video surveillance use case where parts of a video processing pipeline run on an edge server.
+In edge computing, computing and processing is usually done on the \emph{edge} of a network.
+One could for example imagine a drone video surveillance use case where parts of a video processing pipeline run on an edge server at 5G base station.
 This approach would make it possible to execute complex AI models that would otherwise be difficult to run on a drone with limited computing capabilities.
 
-Setting up a video pipeline is a complex problem.
+Setting up a video pipeline utilizing edge resources is a complex problem.
 First, the drone must be able to discover nearby edge servers, make sure the edge server has sufficient resources, deploy some software components at the edge server, then start a video server and finally connect to the video server.
@@
 
-The Colonies framework has the potential to streamline the setup of edge computing workflows by enabling devices and edge servers to be represented as distributed executors that can interact with other.
+The Colonies framework has the potential to streamline the setup of edge computing workflows by enabling devices and edge servers to be represented as executors that can interact with other.
 Figure \ref{fig:edge} illustrates a setup of three executors.
@@
 Submitting a job typically requires users to develop job scripts and manually transfer data to the HPC environment before a job can be executed.
-Compared to cloud platforms, Slurm is less mature when it comes to API and automation support, which can make the usage more time-consuming and manual.
+Compared to cloud platforms, Slurm is less mature when it comes to API and automation support, which can make the use of HPC environments more time-consuming.
 
@@
 With Colonies, users can indirectly control HPC systems by submitting Colonies processes that are subsequently picked up by HPC executors, without requiring direct connection to the HPC system.
-Essentially, using Colonies as a cloud-based submission system for HPC connecting different HPC centers together.
+Essentially, using Colonies as a cloud-based submission system for HPC to connect several HPC centers together.
 Figure \ref{fig:hpc_overlay} depicts such a submission system.
@@
 The \emph{Container Importer Executor} is responsible for converting container images to a format that is supported by Slurm, e.g.
-Singularity containers\footnote{https://docs.sylabs.io}.
+converting Docker containers to Singularity containers\footnote{https://docs.sylabs.io}.
 The \emph{Data Importer Executor} downloads data to the HPC file system.
@@
 \section{Discussion}
-This paper has presented an open-source framework called Colonies designed to streamline execution of computational workloads across platforms.
-Colonies provides a \emph{distributed runtime environment}, called a \emph{Colony}, which serves as an intermediary layer to indirectly control other computer systems.
-Instead of directly connecting different systems together, Colonies allows services or users to publish instructions and then assign these instructions to other services.
+This paper has presented an open-source framework called Colonies providing a \emph{distributed runtime environment} designed to streamline execution of computational workloads across platforms.
+Instead of directly connecting different systems together, Colonies allows special kind of microservices so-called executors to publish instructions which are subsequently assigned to other executors.
 By chaining instructions together, it becomes possible to execute workloads that can operate seamlessly across different platforms.
@@
 
-To sum up, the long-term vision of the Colonies framework is to offer uninterrupted access to computing resources, allowing data to flow from one service to another.
+To sum up, the long-term vision of the Colonies framework is to offer uninterrupted access to computing resources, allowing data to flow from one service to another seamlessly.
 Achieving seamless compute continuum requires development of simple, secure, scalable, and robust solutions.
-Colonies serves as an initial step towards realizing this vision of a seamlessly integrated computing ecosystem.
+Colonies is as an initial step towards realizing this vision of a seamlessly integrated computing ecosystem.
 

### 2023-06-02 paper_2023/paper.tex
improved related work section

@@
 \section{Related work}
+Distributed runtimes enable execution of software applications on many interconnected computing resources.
+It provides mechanisms and services to distribute, manage, and coordinate the execution of tasks.
+Apache Spark \cite{spark} and Apache Flink \cite{flink} are well known examples of distributed runtimes.
+
+Apache Spark is specifically designed for distributed data processing and analytics.
+It operates on distributed data sets, performing transformations and computations on those datasets.
+It offers a rich set of APIs and libraries for data processing.
+Apache Flink, on the other hand, is a distributed stream processing framework to process continuous streams of data in real-time.
+Flink is based on a dataflow programming model and is commonly used in real-time data analytics applications.
+In contrast, Colonies focuses primarily on workflow management.
+It provides a distributed runtime environment that facilitates task execution and coordination across platforms.
+Unlike Spark and Flink, which primarily operate on distributed data sets or data streams, Colonies emphasizes execution and coordination of computational tasks across platforms.
+By enabling the execution of composable functions, Colonies aims to tackle challenges related to portability, automation, scalability, and security in heterogeneous computing environments.
+
 Workflow management has been extensively studied in both academic and industrial settings with numerous approaches \cite{scafe, synapse, service_wfs, schmitt2022workflow, GarciaRepresa1740746, Ouyang2010, NIKOLOV2021100440, workflow_in_bigdata} proposed to address the challenges in this field.
 Recently, Apache Airflow \cite{apache_airflow} has become a popular open-source workflow management system for handling data engineering pipelines.
-Like Colonies, Apache Airflow enables developers to create custom operators and executors that can be integrated with various systems.
+Similar to Colonies, Apache Airflow enables developers to create custom operators and executors that can be integrated with various systems.
 Additionally, Apache Airflow offers an HTTP API that makes it possible to develop Software Development Kits (SDKs) in various programming languages.
 However, Apache Airflow must be integrated with a message broker, such as RabbitMQ \cite{rabbitmq} or Kafka \cite{apache_kafka} to implement task queues, resulting in a more complex architecture than Colonies.
-Furthermore, Colonies is based on a distributed microservice architecture that makes it more suitable for DevOps software development.
-As Colonies is loosely coupled, executors can be implemented and dynamically deployed on the Internet without reconfiguring the workflow engine.
+Moreover, Airflow is primarily designed to be used within one single platform, e.g.
+a Hadoop platform.
 
@@
 It is based on OpenFaaS \cite{openfaas}, which allows functions to run on Kubernetes clusters.
-
 Colonies can also be used to implement serverless workflow management systems.
 This will be further discussed in Section \ref{faas}.
-However, in contrast to previous work, Colonies is designed to be platform independent and does hence not require container orchestration platforms like Kubernetes.
+However, in contrast to previous work, Colonies is designed to be platform independent and does not require a container orchestration platform like Kubernetes.
 By using a zero-trust security protocol, functions can be securely executed by distributed executors deployed anywhere on the Internet.
 
-Currently, microservices are primarily used to implement large-scale web applications or Internet applications requiring high-availability.
+Colonies is based on a distributed microservice architecture, which makes it suitable for DevOps software development.
+Currently, microservices are primarily used to implement large-scale and high-availability applications.
 It has not yet become a prevalent design principle for workload management or implementation of HPC applications.
@@
 The Colonies framework is founded upon a grid computing model with the primary purpose of assigning tasks to distributed executors.
-To the best of the authors' knowledge, no previous work has integrated a microservice-oriented architecture with a grid computing model to serve as an integration point for coordinating artificial intelligence workloads across a diverse range of platforms.
-Additionally, Colonies is designed to function as a ledger, offering complete transparency and execution history, which is essential for implementing zero-trust security.
+To the best of the authors' knowledge, no previous work has integrated a microservice-oriented architecture with a grid computing model to serve as an integration point for coordinating workloads.
+Additionally, Colonies is also designed to function as a ledger, offering complete transparency and execution history, which is essential for implementing zero-trust security.
 
